### 2025-04-19 demo/demo.qmd
init

---
title: European or American? A re-examination of the transatlantic influence over the IMF
subtitle: GV499 Dissertation
header-includes: |
  \AtBeginDocument{\title[GV499 Dissertation]{European or American? A re-examination of the transatlantic influence over the IMF}}
date: last-modified
author: Dianyi Yang
format:
  beamer:
    theme: Madrid
bibliography: references.bib
csl: extra/apa.csl
project:
  execute-dir: "D:/OneDrive - London School of Economics/Desktop/GV499/IMF"
---

## Background

-   The International Monetary Fund (IMF) and the World Bank were established toward the end of World War II as part of the Bretton Woods agreement in 1944 [@jensen2004].

-   Believed US dominance in the beginning,

-   "Gentlemen's agreement": the IMF managing director being a European, and the World Bank president being an American [@Keating2024; @Weiss2023].

## Research Question

Who has more influence over the IMF, the US or Europe?

-   US?

    -   HQ in the US

    -   US veto power

    -   US hegemony

    -   US deputy

    -   Case Study: Mexico

-   Europe?

    -   European managing director

    -   Case Study: Euro Crisis

-   Difficulties in comparison

    -   Consensus-based governance

    -   closed-door negotiations

## Empirical Strategy

-   Data source/empirical inspiration: @lipscy2018 (1980-2010)

    -   but their focus is the effect on self-insurance

    -   did not separate Europe from the US - 'Western influence'

-   Proxies for influence (merged through PCA)

    -   IMF direct: nationals employed as economists, IMF quota.

    -   US: UN affinity score with US, trade volume with US, US bank lending

    -   European: UN affinity score with European powers, trade volume with European powers, European bank lending

## Empirical Strategy

-   Dependent variables

    -   IMF loan to GDP ratio

    -   IMF participation rate

    -   IMF loan approval

    -   IMF conditions

-   Independent variables

    -   IMF influence (used by @lipscy2018 and separated by me into EU, US and IMF)

    -   GDP

    -   Growth

    -   Reserves

    -   OECD

-   Hypotheses

    -   $H_0$: No difference in the influence of US and Europe over the IMF.

    -   $H_1$: Either the US or EU has more influence over the IMF.

    -   Test of equality of the US and EU coefficients.

## Results from [@lipscy2018]

![](images/clipboard-2567109775.png)

## Preliminary Results

```{r setup}
knitr::opts_knit$set(root.dir = 'D:/OneDrive - London School of Economics/Desktop/GV499/IMF')
need <- c('tidyverse','haven','AER','fixest','modelsummary',
          'corrplot', 'kableExtra','car','parameters') # list packages needed
have <- need %in% rownames(installed.packages()) # checks packages you have
if(any(!have)) install.packages(need[!have]) # install missing packages
invisible(lapply(need, library, character.only=T))
```

```{r results}
load('save/regtable.RData')

#t-tests
regtable$`Probit: IMF loan approval`$df.residual <- degrees_of_freedom(regtable$`Probit: IMF loan approval`)

glance_custom.tobit <- function(x, ...) {
  data.frame(
    'equality'=paste0('[','$p=',round(lht(x, test='F', 'us=eu')[2,4],3),'$]'),
    'vcov.type' = 'by: Country'
  )
}

glance_custom.fixest <- function(x, ...) {
  data.frame(
    'equality'=paste0('[','$p=',round(lht(x, test='F', 'us=eu')[2,4],3),'$]'),
    'vcov.type' = 'by: Country'
  )
}

coefmap <- c('us' = 'US Influence',
             'eu' = 'EU Influence',
             'imf' = 'IMF Influence',
             'lnrgdpnew' = 'GDP',
             'lnrgdpnewsq' = 'GDP$^2$',
             'rgdpchnew' = 'GDPpc',
             'rgdpchnewsquare' = "GDPpc$^2$",
             'growth1new' = 'GDPpc growth',
             'reserv1' = 'Reserves',
             'oecd1' = 'OECD',
              '(Intercept)' = '(Intercept)')
gof_map <- list(list("raw" = "equality", "clean" = "US=EU", "fmt" = NULL),
                list("raw" = "nobs", "clean" = "$N$", "fmt" = 0),
                list('raw' = 'vcov.type', 'clean' = 'Std.Errors', 'fmt' = NULL))
msummary(regtable, coef_map=coefmap, gof_map=gof_map, stars=T, escape=F) %>% 
  kable_styling(latex_options="scale_down")
```

## References {.unnumbered}

::: {#refs}
:::

### 2025-04-19 manuscript/dissertation.qmd
init

---
title: European or American? A re-examination of the transatlantic influence over the IMF
subtitle: GV499 Dissertation
date: last-modified
authors:
  - name: Dianyi Yang
    email: d.yang17@lse.ac.uk
    orcid: 0009-0004-4652-3429
    correspondence: true

format:
  #native: default
  #docx: default
  orcid-pdf:
    keep-tex: true
    number-sections: true
bibliography: references.bib
csl: extra/apa.csl
abstract: This is an abstract.
keywords: IMF, US hegemony, Europe
---

\newpage

# Introduction

The International Monetary Fund (IMF) and the World Bank[^1] were established toward the end of World War II as part of the Bretton Woods agreement in 1944 [@jensen2004]. While the World Bank specialised in long-term project financing, the IMF was initially designed for addressing the short-term balance-of-payment crises which would otherwise undermine the fixed exchange rate arrangements under the Bretton Woods system [@jensen2004].

[^1]: Familar name for the International Bank for Reconstruction and Development (IBRD), the lending arm of the later created World Bank Group.

One renowned fact about the IMF and its sister institution, the World Bank, is the long-standing "gentlemen's agreement" of the IMF managing director being a European, and the World Bank president being an American [@Keating2024; @Weiss2023], despite no such provisions in either institution's articles of agreements [@Fund2020; @WorldBank2021].

Such arrangements inspire a straightforward question that is yet difficult to answer: who has more influence over the IMF, the US or Europe? Whereas the US dominance in the World Bank is undisputed [@clark2020; @wade2002; @woods2003], the case of the IMF is much more complicated. As the [Literature Review](#sec-litreview) section demonstrates, there has been abundant evidence of both European and American influence on the IMF, but short of a qualitative or quantitative comparison between the two.

Qualitative attempts at this question may have been thwarted by the closed-door and consensus-based style of governance of the IMF, as epitomised by the "gentlemen's agreement". On the one hand, the frequent use of conventions and consensus renders any analyses on the formal structure of the IMF (e.g. veto power) insufficient on its own. On the other hand, the closed-door negotiations imply that we often observe the outcome, but not the process of the bargaining. While specific case studies (e.g. Mexico and European debt crisis, see @lipscy2018 and @hodson2014) may have shed light on the influence of the US or Europe, they rarely allow for a transatlantic comparison and their representativeness can be questioned.

This dissertation therefore attempts to quantitatively compare the influence of the US and Europe on the IMF.

The structure of this dissertation is as follows: @sec-litreview provides a review over the literature on the influence of the US and Europe over the IMF.

# Literature Review {#sec-litreview}

-   IMF chief European [@Kahler2001]

# References {.unnumbered}

::: {#refs}
:::

### 2025-04-19 manuscript/dissertation.tex
init

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  abstract=true]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[top=30mm,left=30mm,right=30mm,heightrounded]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

% TODO: Add custom LaTeX header directives here
\usepackage{float, booktabs, siunitx, orcidlink}
\setkomafont{disposition}{\bfseries}
\providecommand{\keywords}[1]
{
  \small	
  \textbf
{\textit{Keywords---}} #1
}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={European or American? A re-examination of the transatlantic influence over the IMF},
  pdfkeywords={IMF, US hegemony, Europe},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{European or American? A re-examination of the transatlantic
influence over the IMF}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{GV499 Dissertation}
\author{Dianyi Yang~\orcidlink{0009-0004-4652-3429}\textsuperscript{}}
\date{2024-01-06}

\begin{document}
\maketitle
\begin{abstract}
This is an abstract.
\end{abstract}
\begin{center}
\keywords{IMF, US hegemony, Europe}
\end{center}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, breakable, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, sharp corners]}{\end{tcolorbox}}\fi

\newpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The International Monetary Fund (IMF) and the World Bank\footnote{Familar
  name for the International Bank for Reconstruction and Development
  (IBRD), the lending arm of the later created World Bank Group.} were
established toward the end of World War II as part of the Bretton Woods
agreement in 1944 (\protect\hyperlink{ref-jensen2004}{Jensen, 2004}).
While the World Bank specialised in long-term project financing, the IMF
was initially designed for addressing the short-term balance-of-payment
crises which would otherwise undermine the fixed exchange rate
arrangements under the Bretton Woods system
(\protect\hyperlink{ref-jensen2004}{Jensen, 2004}).

One renowned fact about the IMF and its sister institution, the World
Bank, is the long-standing ``gentlemen's agreement'' of the IMF managing
director being a European, and the World Bank president being an
American (\protect\hyperlink{ref-Keating2024}{Keating, 2024};
\protect\hyperlink{ref-Weiss2023}{Weiss, 2023}), despite no such
provisions in either institution's articles of agreements
(\protect\hyperlink{ref-Fund2020}{IMF, 2020};
\protect\hyperlink{ref-WorldBank2021}{World Bank, 2021}).

Such arrangements inspire a straightforward question that is yet
difficult to answer: who has more influence over the IMF, the US or
Europe? Whereas the US dominance in the World Bank is undisputed
(\protect\hyperlink{ref-clark2020}{Clark \& Dolan, 2020};
\protect\hyperlink{ref-wade2002}{Wade, 2002};
\protect\hyperlink{ref-woods2003}{Woods, 2003}), the case of the IMF is
much more complicated. As the
\protect\hyperlink{sec-litreview}{Literature Review} section
demonstrates, there has been abundant evidence of both European and
American influence on the IMF, but short of a qualitative or
quantitative comparison between the two.

Qualitative attempts at this question may have been thwarted by the
closed-door and consensus-based style of governance of the IMF, as
epitomised by the ``gentlemen's agreement''. On the one hand, the
frequent use of conventions and consensus renders any analyses on the
formal structure of the IMF (e.g.~veto power) insufficient on its own.
On the other hand, the closed-door negotiations imply that we often
observe the outcome, but not the process of the bargaining. While
specific case studies (e.g.~Mexico and European debt crisis, see Lipscy
\& Lee (\protect\hyperlink{ref-lipscy2018}{2018}) and Hodson
(\protect\hyperlink{ref-hodson2014}{2014})) may have shed light on the
influence of the US or Europe, they rarely allow for a transatlantic
comparison and their representativeness can be questioned.

This dissertation therefore attempts to quantitatively compare the
influence of the US and Europe on the IMF.

The structure of this dissertation is as follows:
Section~\ref{sec-litreview} provides a review over the literature on the
influence of the US and Europe over the IMF.

\hypertarget{sec-litreview}{%
\section{Literature Review}\label{sec-litreview}}

\begin{itemize}
\tightlist
\item
  IMF chief European (\protect\hyperlink{ref-Kahler2001}{Kahler, 2001})
\end{itemize}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-clark2020}{}}%
Clark, R., \& Dolan, L. R. (2020). Pleasing the Principal: U.S.
Influence in World Bank Policymaking. \emph{American Journal of
Political Science}, \emph{65}(1), 36--51.
\url{https://doi.org/10.1111/ajps.12531}

\leavevmode\vadjust pre{\hypertarget{ref-hodson2014}{}}%
Hodson, D. (2014). The IMF as a de facto institution of the EU: A
multiple supervisor approach. \emph{Review of International Political
Economy}, \emph{22}(3), 570--598.
\url{https://doi.org/10.1080/09692290.2014.956136}

\leavevmode\vadjust pre{\hypertarget{ref-Fund2020}{}}%
IMF. (2020). \emph{{Articles} of {Agreement}}. International Monetary
Fund.

\leavevmode\vadjust pre{\hypertarget{ref-jensen2004}{}}%
Jensen, N. M. (2004). Crisis, Conditions, and Capital. \emph{Journal of
Conflict Resolution}, \emph{48}(2), 194--210.
\url{https://doi.org/10.1177/0022002703262860}

\leavevmode\vadjust pre{\hypertarget{ref-Kahler2001}{}}%
Kahler, M. (2001). \emph{Leadership selection in the major
multilaterals}. Institute for International Economics.

\leavevmode\vadjust pre{\hypertarget{ref-Keating2024}{}}%
Keating, J. E. (2024). Why {Is} the {IMF} {Chief} {Always} a {European}?
In \emph{Foreign Policy}.
\url{https://foreignpolicy.com/2011/05/18/why-is-the-imf-chief-always-a-european/}

\leavevmode\vadjust pre{\hypertarget{ref-lipscy2018}{}}%
Lipscy, P. Y., \& Lee, H. N.-K. (2018). The IMF As a Biased Global
Insurance Mechanism: Asymmetrical Moral Hazard, Reserve Accumulation,
and Financial Crises. \emph{International Organization}, \emph{73}(1),
35--64. \url{https://doi.org/10.1017/s0020818318000371}

\leavevmode\vadjust pre{\hypertarget{ref-wade2002}{}}%
Wade, R. H. (2002). US hegemony and the World Bank: the fight over
people and ideas. \emph{Review of International Political Economy},
\emph{9}(2), 215--243. \url{https://doi.org/10.1080/09692290110126092}

\leavevmode\vadjust pre{\hypertarget{ref-Weiss2023}{}}%
Weiss, M. A. (2023). Selecting the world bank president. In
\emph{Congressional Research Service}.
\url{https://sgp.fas.org/crs/row/R42463.pdf}

\leavevmode\vadjust pre{\hypertarget{ref-woods2003}{}}%
Woods, N. (2003). \emph{The united states and the international
financial institutions: Power and influence within the world bank and
the IMF} (pp. 92--114). Oxford University PressOxford.
\url{https://doi.org/10.1093/0199261431.003.0005}

\leavevmode\vadjust pre{\hypertarget{ref-WorldBank2021}{}}%
World Bank. (2021). \emph{Articles of agreement}.
\url{https://www.worldbank.org/en/about/articles-of-agreement}

\end{CSLReferences}

\end{document}

### 2025-04-19 code/PCA.R

@@ -1,0 +1,121 @@
+rm(list=ls())
+need <- c('tidyverse','haven','AER','fixest','modelsummary',
+          'corrplot',"ggfortify") # list packages needed
+have <- need %in% rownames(installed.packages()) # checks packages you have
+if(any(!have)) install.packages(need[!have]) # install missing packages
+invisible(lapply(need, library, character.only=T))
+
+setwd(gsub('/code','',dirname(rstudioapi::getSourceEditorContext()$path)))
+data <- read_dta("data/panel_data.dta")
+
+#PCA
+PCA <- data |>
+  select(shstaffl,shquotal,frusanew,
+         frdfgnew,ustradenew,eutradenew,
+         usq4,avgukfrgmq4) |>
+  na.omit() |>
+  prcomp(scale=T)
+summary(PCA)
+
+plot(PCA, type="l")
+autoplot(PCA,loadings = TRUE, loadings.label = TRUE,)
+
+#3 components
+IMF <- a |>
+  select(shstaffl,shquotal) |>
+  na.omit() |>
+  prcomp(scale=T)
+summary(IMF)
+
+US <- a |>
+  select(frusanew, ustradenew, usq4) |>
+  na.omit() |>
+  prcomp(scale=T)
+summary(US)
+
+EU <- a |>
+  select(frdfgnew, eutradenew, avgukfrgmq4) |>
+  na.omit() |>
+  prcomp(scale=T)
+summary(EU)
+
+#regression
+data$imf <- -predict(IMF, data)[,1]
+data$us <- predict(US, data)[,1]
+data$eu <- predict(EU, data)[,1]
+data$pca <- -predict(PCA, data)[,1]
+
+#plot correlation
+corr <- data |>
+  select('imf','us','eu','pca') |>
+  cor(use='pairwise.complete.obs')
+
+corrplot(corr, method='color')
+
+ #tobit regression - IMF loan to GDP ratio
+loan <- tobit(imfloannew100 ~ us + eu + imf + lnrgdpnew + lnrgdpnewsq + rgdpchnew + 
+              rgdpchnewsquare + growth1new + reserv1 + oecd1 + year1980 + year1985 + 
+              year1990 + year1995 + year2000, data=data, left=0, right=Inf, 
+              cluster=shcode)
+
+summary(loan)
+
+ #tobit regression - IMF participation rate
+part <- tobit(imf_p ~ us + eu + imf + lnrgdpnew + lnrgdpnewsq + rgdpchnew + 
[…]

### 2025-04-22 manuscript/dissertation.qmd
Subtitle-fix (#1)
* comment out subtitle

* update.gitignore

* comment out subtitle in dissertation

@@
 title: European or American? A re-examination of the transatlantic influence over the IMF
-subtitle: GV499 Dissertation
+# subtitle: GV499 Dissertation
 date: last-modified

### 2025-04-22 demo/demo.qmd
comment out subtitle
fix demo paths (#3)

@@
 title: European or American? A re-examination of the transatlantic influence over the IMF
-subtitle: GV499 Dissertation
+# subtitle: GV499 Dissertation
 header-includes: |
@@
 ```{r setup}
-knitr::opts_knit$set(root.dir = 'D:/OneDrive - London School of Economics/Desktop/GV499/IMF')
 need <- c('tidyverse','haven','AER','fixest','modelsummary',
@@
 ```{r results}
-load('save/regtable.RData')
+load('../save/regtable.RData')
 
@@
                 list('raw' = 'vcov.type', 'clean' = 'Std.Errors', 'fmt' = NULL))
-msummary(regtable, coef_map=coefmap, gof_map=gof_map, stars=T, escape=F) %>% 
+msummary(regtable, coef_map=coefmap, gof_map=gof_map, stars=T, escape=F, output = "kableExtra") %>% 
   kable_styling(latex_options="scale_down")

### 2025-04-22 demo/demo.qmd
demo: disable package installation

@@
           'corrplot', 'kableExtra','car','parameters') # list packages needed
-have <- need %in% rownames(installed.packages()) # checks packages you have
-if(any(!have)) install.packages(need[!have]) # install missing packages
+# have <- need %in% rownames(installed.packages()) # checks packages you have
+# if(any(!have)) install.packages(need[!have]) # install missing packages
 invisible(lapply(need, library, character.only=T))

### 2025-04-22 demo/demo.qmd
reduce dependencies

--- 2025-04-22 demo/demo.qmd
@@
 ```{r setup}
-need <- c('tidyverse','haven','AER','fixest','modelsummary',
-          'corrplot', 'kableExtra','car','parameters') # list packages needed
-have <- need %in% rownames(installed.packages()) # checks packages you have
-if(any(!have)) install.packages(need[!have]) # install missing packages
-invisible(lapply(need, library, character.only=T))
+need <- c('AER','car','kableExtra','modelsummary','parameters') # list packages needed
+# have <- need %in% rownames(installed.packages()) # checks packages you have
+# if(any(!have)) install.packages(need[!have]) # install missing packages
+invisible(lapply(need, library, character.only = T))
 ```
@@
                 list('raw' = 'vcov.type', 'clean' = 'Std.Errors', 'fmt' = NULL))
-msummary(regtable, coef_map=coefmap, gof_map=gof_map, stars=T, escape=F, output = "kableExtra") %>% 
+msummary(regtable, coef_map=coefmap, gof_map=gof_map, stars=T, escape=F, output = "kableExtra") |>
   kable_styling(latex_options="scale_down")

### 2025-04-22 demo/demo.qmd
Fix yaml

@@
 csl: extra/apa.csl
-project:
-  execute-dir: "D:/OneDrive - London School of Economics/Desktop/GV499/IMF"
 ---

### 2025-04-22 demo/demo.qmd
use project dir as wd
Change subtitle
add name placeholders

@@
 title: European or American? A re-examination of the transatlantic influence over the IMF
-# subtitle: GV499 Dissertation
 header-includes: |
-  \AtBeginDocument{\title[GV499 Dissertation]{European or American? A re-examination of the transatlantic influence over the IMF}}
+  \AtBeginDocument{\title[European or American]{European or American? A re-examination of the transatlantic influence over the IMF}}
 date: last-modified
-author: Dianyi Yang
+author:
+- name: Dianyi Yang
+# - name: Tong Hu // name placeholders subject to change
+# - name: Zhenyi Chen
 format:
@@
 ```{r results}
-load('../save/regtable.RData')
+load('save/regtable.RData')
 

### 2025-04-22 demo/demo.qmd
demo: reformat

@@
 
-    -   IMF loan to GDP ratio
+    -   IMF loan to GDP ratio, IMF participation rate
 
-    -   IMF participation rate
-
-    -   IMF loan approval
-
-    -   IMF conditions
+    -   IMF loan approval, IMF conditions
 
@@
 
-    -   GDP
+    -   GDP, Growth
 
-    -   Growth
-
-    -   Reserves
-
-    -   OECD
+    -   Reserves, OECD
 
@@
 ```{r setup}
-need <- c('AER','car','kableExtra','modelsummary','parameters') # list packages needed
+need <- c('AER', 'car', 'kableExtra', 'modelsummary', 'parameters') # list packages needed
 # have <- need %in% rownames(installed.packages()) # checks packages you have
@@
 #t-tests
-regtable$`Probit: IMF loan approval`$df.residual <- degrees_of_freedom(regtable$`Probit: IMF loan approval`)
+regtable$`Probit: IMF loan approval`$df.residual <- degrees_of_freedom(
+  regtable$`Probit: IMF loan approval`
+)
 
@@
   data.frame(
-    'equality'=paste0('[','$p=',round(lht(x, test='F', 'us=eu')[2,4],3),'$]'),
+    'equality' = paste0(
+      '[',
+      '$p=',
+      round(lht(x, test = 'F', 'us=eu')[2, 4], 3),
+      '$]'
+    ),
     'vcov.type' = 'by: Country'
@@
 
-coefmap <- c('us' = 'US Influence',
+coefmap <- c(
+  'us' = 'US Influence',
   'eu' = 'EU Influence',
@@
   'oecd1' = 'OECD',
-              '(Intercept)' = '(Intercept)')
-gof_map <- list(list("raw" = "equality", "clean" = "US=EU", "fmt" = NULL),
+  '(Intercept)' = '(Intercept)'
+)
+gof_map <- list(
+  list("raw" = "equality", "clean" = "US=EU", "fmt" = NULL),
   list("raw" = "nobs", "clean" = "$N$", "fmt" = 0),
-                list('raw' = 'vcov.type', 'clean' = 'Std.Errors', 'fmt' = NULL))
-msummary(regtable, coef_map=coefmap, gof_map=gof_map, stars=T, escape=F, output = "kableExtra") |>
-  kable_styling(latex_options="scale_down")
+  list('raw' = 'vcov.type', 'clean' = 'Std.Errors', 'fmt' = NULL)
+)
+msummary(
+  regtable,
+  coef_map = coefmap,
+  gof_map = gof_map,
+  stars = T,
+  escape = F,
+  output = "kableExtra"
+) |>
+  kable_styling(latex_options = "scale_down")
 ```

### 2025-04-22 demo/demo.qmd
Overhaul demo (#4)
* demo: disable package installation

* reduce dependencies

* Fix yaml

* use project dir as wd

* Change subtitle

* add name placeholders

* demo: reformat

* demo: change theme colour

* demo: update pdf:

@@
     theme: Madrid
+    include-in-header: 
+      - text: |
+          \definecolor{myblue}{RGB}{0, 113, 187}
+          \colorlet{beamer@blendedblue}{myblue}
 bibliography: references.bib

### 2025-04-22 manuscript/dissertation.tex
share csl and bib

@@
 % Make \paragraph and \subparagraph free-standing
+\makeatletter
 \ifx\paragraph\undefined\else
   \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+  \renewcommand{\paragraph}{
+    \@ifstar
+      \xxxParagraphStar
+      \xxxParagraphNoStar
+  }
+  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
+  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
 \fi
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}{
+    \@ifstar
+      \xxxSubParagraphStar
+      \xxxSubParagraphNoStar
+  }
+  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
+  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
+\makeatother
 
@@
 \makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\newsavebox\pandoc@box
+\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
+  \sbox\pandoc@box{#1}%
+  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
+  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
+  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
+  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
+  \else\usebox{\pandoc@box}%
+  \fi%
+}
+% Set default figure placement to htbp
+\def\fps@figure{htbp}
 \makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
+% definitions for citeproc citations
+\NewDocumentCommand\citeproctext{}{}
+\NewDocumentCommand\citeproc{mm}{%
+  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
 \makeatletter
-\def\fps@figure{htbp}
+ % allow citations to break across lines
+ \let\@cite@ofmt\@firstofone
+ % avoid brackets around text for \cite:
+ \def\@biblabel#1{}
+ \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
 \makeatother
@@
 \setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
+\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
+ {\begin{list}{}{%
+  \setlength{\itemindent}{0pt}
+  \setlength{\leftmargin}{0pt}
+  \setlength{\parsep}{0pt}
   % turn on hanging indent if param 1 is 1
   \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
+   \setlength{\leftmargin}{\cslhangindent}
+   \setlength{\itemindent}{-1\cslhangindent}
   \fi
   % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
- }%
- {}
+  \setlength{\itemsep}{#2\baselineskip}}}
+ {\end{list}}
 \usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
@@
 }
-\makeatletter
-\makeatother
-\makeatletter
-\makeatother
 \makeatletter
@@
 \makeatletter
+\makeatother
+\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
@@
 \makeatother
-\makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
-\makeatother
-\makeatletter
-\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
-\makeatother
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+
+\usepackage{bookmark}
+
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 influence over the IMF}
-\usepackage{etoolbox}
-\makeatletter
-\providecommand{\subtitle}[1]{% add subtitle to \maketitle
-  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
-}
-\makeatother
-\subtitle{GV499 Dissertation}
 \author{Dianyi Yang~\orcidlink{0009-0004-4652-3429}\textsuperscript{}}
-\date{2024-01-06}
+\date{2025-04-22}
 
@@
 \end{center}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, breakable, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}
 
@@
 established toward the end of World War II as part of the Bretton Woods
-agreement in 1944 (\protect\hyperlink{ref-jensen2004}{Jensen, 2004}).
-While the World Bank specialised in long-term project financing, the IMF
-was initially designed for addressing the short-term balance-of-payment
+agreement in 1944 (\citeproc{ref-jensen2004}{Jensen, 2004}). While the
+World Bank specialised in long-term project financing, the IMF was
+initially designed for addressing the short-term balance-of-payment
 crises which would otherwise undermine the fixed exchange rate
 arrangements under the Bretton Woods system
-(\protect\hyperlink{ref-jensen2004}{Jensen, 2004}).
+(\citeproc{ref-jensen2004}{Jensen, 2004}).
 
@@
 director being a European, and the World Bank president being an
-American (\protect\hyperlink{ref-Keating2024}{Keating, 2024};
-\protect\hyperlink{ref-Weiss2023}{Weiss, 2023}), despite no such
-provisions in either institution's articles of agreements
-(\protect\hyperlink{ref-Fund2020}{IMF, 2020};
-\protect\hyperlink{ref-WorldBank2021}{World Bank, 2021}).
+American (\citeproc{ref-Keating2024}{Keating, 2024};
+\citeproc{ref-Weiss2023}{Weiss, 2023}), despite no such provisions in
+either institution's articles of agreements
+(\citeproc{ref-Fund2020}{IMF, 2020}; \citeproc{ref-WorldBank2021}{World
+Bank, 2021}).
 
@@
 Europe? Whereas the US dominance in the World Bank is undisputed
-(\protect\hyperlink{ref-clark2020}{Clark \& Dolan, 2020};
-\protect\hyperlink{ref-wade2002}{Wade, 2002};
-\protect\hyperlink{ref-woods2003}{Woods, 2003}), the case of the IMF is
-much more complicated. As the
-\protect\hyperlink{sec-litreview}{Literature Review} section
-demonstrates, there has been abundant evidence of both European and
-American influence on the IMF, but short of a qualitative or
-quantitative comparison between the two.
+(\citeproc{ref-clark2020}{Clark \& Dolan, 2020};
+\citeproc{ref-wade2002}{Wade, 2002}; \citeproc{ref-woods2003}{Woods,
+2003}), the case of the IMF is much more complicated. As the
+\hyperref[sec-litreview]{Literature Review} section demonstrates, there
+has been abundant evidence of both European and American influence on
+the IMF, but short of a qualitative or quantitative comparison between
+the two.
 
@@
 specific case studies (e.g.~Mexico and European debt crisis, see Lipscy
-\& Lee (\protect\hyperlink{ref-lipscy2018}{2018}) and Hodson
-(\protect\hyperlink{ref-hodson2014}{2014})) may have shed light on the
-influence of the US or Europe, they rarely allow for a transatlantic
-comparison and their representativeness can be questioned.
+\& Lee (\citeproc{ref-lipscy2018}{2018}) and Hodson
+(\citeproc{ref-hodson2014}{2014})) may have shed light on the influence
+of the US or Europe, they rarely allow for a transatlantic comparison
+and their representativeness can be questioned.
 
@@
 
-\hypertarget{sec-litreview}{%
-\section{Literature Review}\label{sec-litreview}}
+\section{Literature Review}\label{sec-litreview}
 
@@
 \item
-  IMF chief European (\protect\hyperlink{ref-Kahler2001}{Kahler, 2001})
+  IMF chief European (\citeproc{ref-Kahler2001}{Kahler, 2001})
 \end{itemize}
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
+\section*{References}\label{references}
 \addcontentsline{toc}{section}{References}
 
-\hypertarget{refs}{}
+\phantomsection\label{refs}
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-clark2020}{}}%
+\bibitem[\citeproctext]{ref-clark2020}
 Clark, R., \& Dolan, L. R. (2020). Pleasing the Principal: U.S.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-hodson2014}{}}%
+\bibitem[\citeproctext]{ref-hodson2014}
 Hodson, D. (2014). The IMF as a de facto institution of the EU: A
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Fund2020}{}}%
+\bibitem[\citeproctext]{ref-Fund2020}
 IMF. (2020). \emph{{Articles} of {Agreement}}. International Monetary
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-jensen2004}{}}%
+\bibitem[\citeproctext]{ref-jensen2004}
 Jensen, N. M. (2004). Crisis, Conditions, and Capital. \emph{Journal of
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Kahler2001}{}}%
+\bibitem[\citeproctext]{ref-Kahler2001}
 Kahler, M. (2001). \emph{Leadership selection in the major
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Keating2024}{}}%
+\bibitem[\citeproctext]{ref-Keating2024}
 Keating, J. E. (2024). Why {Is} the {IMF} {Chief} {Always} a {European}?
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-lipscy2018}{}}%
+\bibitem[\citeproctext]{ref-lipscy2018}
 Lipscy, P. Y., \& Lee, H. N.-K. (2018). The IMF As a Biased Global
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-wade2002}{}}%
+\bibitem[\citeproctext]{ref-wade2002}
 Wade, R. H. (2002). US hegemony and the World Bank: the fight over
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Weiss2023}{}}%
+\bibitem[\citeproctext]{ref-Weiss2023}
 Weiss, M. A. (2023). Selecting the world bank president. In
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-woods2003}{}}%
+\bibitem[\citeproctext]{ref-woods2003}
 Woods, N. (2003). \emph{The united states and the international
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-WorldBank2021}{}}%
+\bibitem[\citeproctext]{ref-WorldBank2021}
 World Bank. (2021). \emph{Articles of agreement}.

### 2025-04-22 manuscript/dissertation.qmd
share csl and bib (#5)

--- 2025-04-19 manuscript/dissertation.qmd
@@
     correspondence: true
-
 format:
@@
     number-sections: true
-bibliography: references.bib
-csl: extra/apa.csl
 abstract: This is an abstract.

### 2025-04-22 code/PCA.R
Add info and refactor dependencies

@@ -1,9 +1,26 @@
-rm(list=ls())
-need <- c('tidyverse','haven','AER','fixest','modelsummary',
-          'corrplot',"ggfortify") # list packages needed
+######## INFO ########
+# Authors: Dianyi Yang
+# R Script
+# Purpose: This script runs PCA and regression analysis for the demo.
+# Inputs: data/panel_data.dta
+# Outputs: save/regtable.RData
+
+
+# SETUP ----------------------------------------------------------------------------------------------------------------
+rm(list = ls()) # Clear workspace
+
+need <- c(
+  'tidyverse',
+  'haven',
+  'AER',
+  'fixest',
+  'modelsummary',
+  'corrplot',
+  "ggfortify"
+) # list packages needed
 have <- need %in% rownames(installed.packages()) # checks packages you have
-if(any(!have)) install.packages(need[!have]) # install missing packages
-invisible(lapply(need, library, character.only=T))
+if (any(!have)) install.packages(need[!have]) # install missing packages
+invisible(lapply(need, library, character.only = T))
 
 setwd(gsub('/code','',dirname(rstudioapi::getSourceEditorContext()$path)))
 data <- read_dta("data/panel_data.dta")

### 2025-04-22 code/PCA.R
remove setwd

@@ -22,8 +22,7 @@ have <- need %in% rownames(installed.packages()) # checks packages you have
 if (any(!have)) install.packages(need[!have]) # install missing packages
 invisible(lapply(need, library, character.only = T))
 
-setwd(gsub('/code','',dirname(rstudioapi::getSourceEditorContext()$path)))
-data <- read_dta("data/panel_data.dta")
+data <- read_dta("data/panel_data.dta") # Load data
 
 #PCA
 PCA <- data |>

### 2025-04-22 code/PCA.R
Reformat regressions

@@ -80,33 +80,103 @@ corr <- data |>
 
 corrplot(corr, method = 'color')
 
- #tobit regression - IMF loan to GDP ratio
-loan <- tobit(imfloannew100 ~ us + eu + imf + lnrgdpnew + lnrgdpnewsq + rgdpchnew + 
-              rgdpchnewsquare + growth1new + reserv1 + oecd1 + year1980 + year1985 + 
-              year1990 + year1995 + year2000, data=data, left=0, right=Inf, 
-              cluster=shcode)
-
+## tobit regression - IMF loan to GDP ratio
+loan <- tobit(
+  imfloannew100 ~
+    us +
+      eu +
+      imf +
+      lnrgdpnew +
+      lnrgdpnewsq +
+      rgdpchnew +
+      rgdpchnewsquare +
+      growth1new +
+      reserv1 +
+      oecd1 +
+      year1980 +
+      year1985 +
+      year1990 +
+      year1995 +
+      year2000,
+  data = data,
+  left = 0,
+  right = Inf,
+  cluster = shcode
+)
 summary(loan)
 
- #tobit regression - IMF participation rate
-part <- tobit(imf_p ~ us + eu + imf + lnrgdpnew + lnrgdpnewsq + rgdpchnew + 
-              rgdpchnewsquare + growth1new + reserv1 + oecd1 + year1980 + year1985 + 
-              year1990 + year1995 + year2000, data=data, left=0, right=1, 
-              cluster=shcode)
+## Tobit regression - IMF participation rate ----------------------------------------------------------------------------------------------------------------
+part <- tobit(
+  imf_p ~
+    us +
+      eu +
+      imf +
+      lnrgdpnew +
+      lnrgdpnewsq +
+      rgdpchnew +
+      rgdpchnewsquare +
+      growth1new +
+      reserv1 +
+      oecd1 +
+      year1980 +
+      year1985 +
+      year1990 +
+      year1995 +
+      year2000,
+  data = data,
+  left = 0,
+  right = 1,
+  cluster = shcode
+)
 summary(part)
 
- #probit regression - IMF loan approval
-approval <- feglm(imf5a ~ us + eu + imf + lnrgdpnew + lnrgdpnewsq + rgdpchnew + 
-                  rgdpchnewsquare + growth1new + reserv1 + oecd1 + year1980 + year1985 + 
-                  year1990 + year1995 + year2000, data=data, family=binomial(link="probit"),
-                  cluster='shcode')
[…]

### 2025-04-22 code/PCA.R
Refactor model saving

@@ -179,16 +179,15 @@ condition <- tobit(
 )
 summary(condition)
 
-#t-tests
-
-
-#regression table
-regtable <- list('Tobit: IMF loan to GDP ratio' = loan, 
+# Save the models --------------------------------------------------------------------------------------------------------------------------------------------
+regModels <- list(
+  'Tobit: IMF loan to GDP ratio' = loan,
                  'Tobit: IMF participation rate' = part, 
                  'Probit: IMF loan approval' = approval, 
-                 'Tobit: number of IMF conditions' = condition)
-save(regtable, data, file='save/regtable.RData')
-msummary(regtable, stars=T)
+  'Tobit: number of IMF conditions' = condition
+)
+save(regModels, data, file = 'save/regMable.RData')
+msummary(regModels, stars = T)
 
 #ftest
 #tobit regression - IMF loan to GDP ratio

### 2025-04-22 code/PCA.R
save model to regtable to regModels

@@ -186,7 +186,7 @@ regModels <- list(
                  'Probit: IMF loan approval' = approval, 
   'Tobit: number of IMF conditions' = condition
 )
-save(regModels, data, file = 'save/regMable.RData')
+save(regModels, data, file = 'save/regModels.RData')
 msummary(regModels, stars = T)
 
 #ftest

### 2025-04-22 code/PCA.R
Reformat auxiliary f-tests

@@ -189,30 +189,76 @@ regModels <- list(
 save(regModels, data, file = 'save/regModels.RData')
 msummary(regModels, stars = T)
 
-#ftest
-#tobit regression - IMF loan to GDP ratio
-loan_f <- tobit(imfloannew100 ~ shstaffl+shquotal+frusanew+frdfgnew+ustradenew+
-                  eutradenew+usq4+avgukfrgmq4 + lnrgdpnew + lnrgdpnewsq + rgdpchnew + 
-                rgdpchnewsquare + growth1new + reserv1 + oecd1 + year1980 + year1985 + 
-                year1990 + year1995 + year2000, data=data, left=0, right=Inf, 
-              cluster=shcode)
-  #significance of US
+# Auxiliary F-tests -------------------------------------------------------------------------------------------------------------------------------------------
+# You don't have to run this to replicate the demo.
+
+## Tobit regression - IMF loan to GDP ratio -------------------------------------------------------------------------------------------------------------------
+loan_f <- tobit(
+  imfloannew100 ~
+    shstaffl +
+      shquotal +
+      frusanew +
+      frdfgnew +
+      ustradenew +
+      eutradenew +
+      usq4 +
+      avgukfrgmq4 +
+      lnrgdpnew +
+      lnrgdpnewsq +
+      rgdpchnew +
+      rgdpchnewsquare +
+      growth1new +
+      reserv1 +
+      oecd1 +
+      year1980 +
+      year1985 +
+      year1990 +
+      year1995 +
+      year2000,
+  data = data,
+  left = 0,
+  right = Inf,
+  cluster = shcode
+)
+### significance of US ----
   matrix = c('frusanew = 0', 'ustradenew = 0', 'usq4 = 0')
-  lht(loan_f, test='F', matrix)
-  #significance of EU
+lht(loan_f, test = 'F', matrix)
+### significance of EU ----
   matrix = c('frdfgnew = 0', 'eutradenew = 0', 'avgukfrgmq4 = 0')
-  lht(loan_f, test='F', matrix)
+lht(loan_f, test = 'F', matrix)
   
-#tobit Total number of IMF conditions
-  condition_f <- tobit(tc ~ shstaffl+shquotal+frusanew+frdfgnew+ustradenew+
-                         eutradenew+usq4+avgukfrgmq4 + lnrgdpnew + lnrgdpnewsq + rgdpchnew + 
[…]

### 2025-04-22 code/PCA.R
Reformat whitespaces

@@ -68,7 +68,7 @@ plot(EU, type = "l")
 # Run regression ---------------------------------------------------------------------------------------------------------------------------------------------------------------
 
 ## Create primary component variables for regression ---------------------------------------------------------------------------------------------------------------------------
-data$imf <- -predict(IMF, data)[, 1]
+data$imf <- -predict(IMF, data)[, 1] 
 data$us <- predict(US, data)[, 1]
 data$eu <- predict(EU, data)[, 1]
 data$pca <- -predict(PCA, data)[, 1]
@@ -182,8 +182,8 @@ summary(condition)
 # Save the models --------------------------------------------------------------------------------------------------------------------------------------------
 regModels <- list(
   'Tobit: IMF loan to GDP ratio' = loan,
-                 'Tobit: IMF participation rate' = part, 
-                 'Probit: IMF loan approval' = approval, 
+  'Tobit: IMF participation rate' = part,
+  'Probit: IMF loan approval' = approval,
   'Tobit: number of IMF conditions' = condition
 )
 save(regModels, data, file = 'save/regModels.RData')
@@ -221,12 +221,12 @@ loan_f <- tobit(
   cluster = shcode
 )
 ### significance of US ----
-  matrix = c('frusanew = 0', 'ustradenew = 0', 'usq4 = 0')
+matrix = c('frusanew = 0', 'ustradenew = 0', 'usq4 = 0')
 lht(loan_f, test = 'F', matrix)
 ### significance of EU ----
-  matrix = c('frdfgnew = 0', 'eutradenew = 0', 'avgukfrgmq4 = 0')
+matrix = c('frdfgnew = 0', 'eutradenew = 0', 'avgukfrgmq4 = 0')
 lht(loan_f, test = 'F', matrix)
-  
+
 ## Tobit Total number of IMF conditions -----------------------------------------------------------------------------------------------------------------------
 condition_f <- tobit(
   tc ~
@@ -256,9 +256,8 @@ condition_f <- tobit(
   cluster = shcode
 )
 ### significance of US ----
-  matrix = c('frusanew = 0', 'ustradenew = 0', 'usq4 = 0')
+matrix = c('frusanew = 0', 'ustradenew = 0', 'usq4 = 0')
 lht(condition_f, test = 'F', matrix)
 ### significance of EU ----
[…]

### 2025-04-22 code/PCA.R
Annotate msummary

@@ -187,7 +187,7 @@ regModels <- list(
   'Tobit: number of IMF conditions' = condition
 )
 save(regModels, data, file = 'save/regModels.RData')
-msummary(regModels, stars = T)
+msummary(regModels, stars = T) # Rough regression table with all variables and stars
 
 # Auxiliary F-tests -------------------------------------------------------------------------------------------------------------------------------------------
 # You don't have to run this to replicate the demo.

### 2025-04-22 code/PCA.R
Refactored PCA R Code (#8)
* Add info and refactor dependencies

* remove setwd

* Annotate Lipscy & Lee's original PCA

* Reformat 3 components and prediction

* add plots for three components

* annotate pca variables creation and corrplot

* Reformat regressions

* Refactor model saving

* save model to regtable to regModels

* Reformat auxiliary f-tests

* Reformat whitespaces

* Annotate msummary

@@ -1,121 +1,263 @@
-rm(list=ls())
-need <- c('tidyverse','haven','AER','fixest','modelsummary',
-          'corrplot',"ggfortify") # list packages needed
+######## INFO ########
+# Authors: Dianyi Yang
+# R Script
+# Purpose: This script runs PCA and regression analysis for the demo.
+# Inputs: data/panel_data.dta
+# Outputs: save/regtable.RData
+
+
+# SETUP ----------------------------------------------------------------------------------------------------------------
+rm(list = ls()) # Clear workspace
+
+need <- c(
+  'tidyverse',
+  'haven',
+  'AER',
+  'fixest',
+  'modelsummary',
+  'corrplot',
+  "ggfortify"
+) # list packages needed
 have <- need %in% rownames(installed.packages()) # checks packages you have
-if(any(!have)) install.packages(need[!have]) # install missing packages
-invisible(lapply(need, library, character.only=T))
+if (any(!have)) install.packages(need[!have]) # install missing packages
+invisible(lapply(need, library, character.only = T))
 
-setwd(gsub('/code','',dirname(rstudioapi::getSourceEditorContext()$path)))
-data <- read_dta("data/panel_data.dta")
+data <- read_dta("data/panel_data.dta") # Load data
 
-#PCA
+# Replicate Lipscy & Lee's (2018) original PCA analysis in R ------------------------------------------------------------------------------------------------------------------
 PCA <- data |>
-  select(shstaffl,shquotal,frusanew,
-         frdfgnew,ustradenew,eutradenew,
-         usq4,avgukfrgmq4) |>
+  select(
+    shstaffl,
+    shquotal,
+    frusanew,
+    frdfgnew,
+    ustradenew,
+    eutradenew,
+    usq4,
+    avgukfrgmq4
+  ) |>
   na.omit() |>
-  prcomp(scale=T)
-summary(PCA)
+  prcomp(scale = T)
+summary(PCA) # show primary components
 
-plot(PCA, type="l")
-autoplot(PCA,loadings = TRUE, loadings.label = TRUE,)
+plot(PCA, type = "l") # Scree plot - "elbow" shape indicates good representation by the first component 
+autoplot(PCA, loadings = TRUE, loadings.label = TRUE) # shows components of the first two components
 
-#3 components
-IMF <- a |>
-  select(shstaffl,shquotal) |>
[…]

### 2025-04-22 code/PCA.R
Add load point

@@ -187,6 +187,8 @@ regModels <- list(
   'Tobit: number of IMF conditions' = condition
 )
 save(regModels, data, file = 'save/regModels.RData')
+
+# load('save/regModels.RData') # Load the models again if needed
 msummary(regModels, stars = T) # Rough regression table with all variables and stars
 
 # Auxiliary F-tests -------------------------------------------------------------------------------------------------------------------------------------------

### 2025-04-22 code/PCA.R
Rename output

@@ -3,7 +3,7 @@
 # R Script
 # Purpose: This script runs PCA and regression analysis for the demo.
 # Inputs: data/panel_data.dta
-# Outputs: save/regtable.RData
+# Outputs: save/regModels.RData
 
 
 # SETUP ----------------------------------------------------------------------------------------------------------------

### 2025-04-22 code/PCA.R
Reformat PCA.R

@@ -5,7 +5,6 @@
 # Inputs: data/panel_data.dta
 # Outputs: save/regModels.RData
 
-
 # SETUP ----------------------------------------------------------------------------------------------------------------
 rm(list = ls()) # Clear workspace
 
@@ -68,7 +67,7 @@ plot(EU, type = "l")
 # Run regression ---------------------------------------------------------------------------------------------------------------------------------------------------------------
 
 ## Create primary component variables for regression ---------------------------------------------------------------------------------------------------------------------------
-data$imf <- -predict(IMF, data)[, 1] 
+data$imf <- -predict(IMF, data)[, 1]
 data$us <- predict(US, data)[, 1]
 data$eu <- predict(EU, data)[, 1]
 data$pca <- -predict(PCA, data)[, 1]
@@ -179,7 +178,9 @@ condition <- tobit(
 )
 summary(condition)
 
-# Save the models --------------------------------------------------------------------------------------------------------------------------------------------
+# Table generation -------------------------------------------------------------------------------------------------------------------------------------------
+
+## Save the models --------------------------------------------------------------------------------------------------------------------------------------------
 regModels <- list(
   'Tobit: IMF loan to GDP ratio' = loan,
   'Tobit: IMF participation rate' = part,

### 2025-04-22 code/PCA.R
add KE as a dependency

@@ -13,6 +13,7 @@ need <- c(
   'haven',
   'AER',
   'fixest',
+  "kableExtra",
   'modelsummary',
   'corrplot',
   "ggfortify"

### 2025-04-22 code/PCA.R
Add parameters as a dependency

@@ -16,7 +16,8 @@ need <- c(
   "kableExtra",
   'modelsummary',
   'corrplot',
-  "ggfortify"
+  "ggfortify",
+  "parameters"
 ) # list packages needed
 have <- need %in% rownames(installed.packages()) # checks packages you have
 if (any(!have)) install.packages(need[!have]) # install missing packages

### 2025-04-22 code/PCA.R
whitespace change

@@ -41,7 +41,7 @@ PCA <- data |>
   prcomp(scale = T)
 summary(PCA) # show primary components
 
-plot(PCA, type = "l") # Scree plot - "elbow" shape indicates good representation by the first component 
+plot(PCA, type = "l") # Scree plot - "elbow" shape indicates good representation by the first component
 autoplot(PCA, loadings = TRUE, loadings.label = TRUE) # shows components of the first two components
 
 # Break down PCA into IMF, US, and EU components ------------------------------------------------------------------------------------------------------------------------------

### 2025-04-22 code/PCA.R
copy and paste result table making from demo

@@ -194,6 +194,67 @@ save(regModels, data, file = 'save/regModels.RData')
 # load('save/regModels.RData') # Load the models again if needed
 msummary(regModels, stars = T) # Rough regression table with all variables and stars
 
+## t-tests generation -----------------------------------------------------------------------------------------------------------------------------------------
+
+regModels$`Probit: IMF loan approval`$df.residual <- degrees_of_freedom(
+  regModels$`Probit: IMF loan approval`
+)
+
+glance_custom.tobit <- function(x, ...) {
+  data.frame(
+    'equality' = paste0(
+      '[',
+      '$p=',
+      round(lht(x, test = 'F', 'us=eu')[2, 4], 3),
+      '$]'
+    ),
+    'vcov.type' = 'by: Country'
+  )
+}
+
+glance_custom.fixest <- function(x, ...) {
+  data.frame(
+    'equality' = paste0(
+      '[',
+      '$p=',
+      round(lht(x, test = 'F', 'us=eu')[2, 4], 3),
+      '$]'
+    ),
+    'vcov.type' = 'by: Country'
+  )
+}
+
+## Variable renaming -------------------------------------------------------------------------------------------------------------------------------------------
+
+coefmap <- c(
+  'us' = 'US Influence',
+  'eu' = 'EU Influence',
+  'imf' = 'IMF Influence',
+  'lnrgdpnew' = 'GDP',
+  'lnrgdpnewsq' = 'GDP$^2$',
+  'rgdpchnew' = 'GDPpc',
+  'rgdpchnewsquare' = "GDPpc$^2$",
+  'growth1new' = 'GDPpc growth',
+  'reserv1' = 'Reserves',
+  'oecd1' = 'OECD',
+  '(Intercept)' = '(Intercept)'
+)
+gof_map <- list(
+  list("raw" = "equality", "clean" = "US=EU", "fmt" = NULL),
+  list("raw" = "nobs", "clean" = "$N$", "fmt" = 0),
+  list('raw' = 'vcov.type', 'clean' = 'Std.Errors', 'fmt' = NULL)
+)
+
+(resultsTable <- msummary(
+  regModels,
+  coef_map = coefmap,
+  gof_map = gof_map,
+  stars = T,
+  escape = F,
+  output = "kableExtra"
+) |>
+  kable_styling(latex_options = "scale_down"))
+
 # Auxiliary F-tests -------------------------------------------------------------------------------------------------------------------------------------------
 # You don't have to run this to replicate the demo.
 

### 2025-04-22 demo/demo.qmd
share csl and bib
Make table in PCA.R (with tinytable)

@@
           \colorlet{beamer@blendedblue}{myblue}
-bibliography: references.bib
-csl: extra/apa.csl
 ---
@@
 ```{r results}
-load('save/regtable.RData')
-
-#t-tests
-regtable$`Probit: IMF loan approval`$df.residual <- degrees_of_freedom(
-  regtable$`Probit: IMF loan approval`
-)
-
-glance_custom.tobit <- function(x, ...) {
-  data.frame(
-    'equality' = paste0(
-      '[',
-      '$p=',
-      round(lht(x, test = 'F', 'us=eu')[2, 4], 3),
-      '$]'
-    ),
-    'vcov.type' = 'by: Country'
-  )
-}
-
-glance_custom.fixest <- function(x, ...) {
-  data.frame(
-    'equality' = paste0(
-      '[',
-      '$p=',
-      round(lht(x, test = 'F', 'us=eu')[2, 4], 3),
-      '$]'
-    ),
-    'vcov.type' = 'by: Country'
-  )
-}
-
-coefmap <- c(
-  'us' = 'US Influence',
-  'eu' = 'EU Influence',
-  'imf' = 'IMF Influence',
-  'lnrgdpnew' = 'GDP',
-  'lnrgdpnewsq' = 'GDP$^2$',
-  'rgdpchnew' = 'GDPpc',
-  'rgdpchnewsquare' = "GDPpc$^2$",
-  'growth1new' = 'GDPpc growth',
-  'reserv1' = 'Reserves',
-  'oecd1' = 'OECD',
-  '(Intercept)' = '(Intercept)'
-)
-gof_map <- list(
-  list("raw" = "equality", "clean" = "US=EU", "fmt" = NULL),
-  list("raw" = "nobs", "clean" = "$N$", "fmt" = 0),
-  list('raw' = 'vcov.type', 'clean' = 'Std.Errors', 'fmt' = NULL)
-)
-msummary(
-  regtable,
-  coef_map = coefmap,
-  gof_map = gof_map,
-  stars = T,
-  escape = F,
-  output = "kableExtra"
-) |>
-  kable_styling(latex_options = "scale_down")
+load('save/regTable.RData')
+resultsTable
 ```

### 2025-04-22 code/PCA.R

@@ -245,15 +245,22 @@ gof_map <- list(
   list('raw' = 'vcov.type', 'clean' = 'Std.Errors', 'fmt' = NULL)
 )
 
-(resultsTable <- msummary(
-  regModels,
-  coef_map = coefmap,
-  gof_map = gof_map,
-  stars = T,
-  escape = F,
-  output = "kableExtra"
-) |>
-  kable_styling(latex_options = "scale_down"))
+getTable <- function(output = "html") {
+  msummary(
+    regModels,
+    coef_map = coefmap,
+    gof_map = gof_map,
+    stars = T,
+    escape = F,
+    output = output
+  )
+}
+
+getTable() # for previewing the table in IDE
+
+resultsTable <- getTable("latex") |> tinytable::theme_tt("resize", width = 0.95) # LaTeX table or slide/paper presentation
+
+save(resultsTable, file = 'save/regTable.RData') # Save the table
 
 # Auxiliary F-tests -------------------------------------------------------------------------------------------------------------------------------------------
 # You don't have to run this to replicate the demo.

### 2025-04-22 demo/demo.qmd
no packages are needed for demo

@@
 
-```{r setup}
-need <- c('AER', 'car', 'kableExtra', 'modelsummary', 'parameters') # list packages needed
-# have <- need %in% rownames(installed.packages()) # checks packages you have
-# if(any(!have)) install.packages(need[!have]) # install missing packages
-invisible(lapply(need, library, character.only = T))
-```
-
 ```{r results}

### 2025-04-22 code/PCA.R
Move table generation from demo to PCA.R (#9)
* Add load point

* Rename output

* Reformat PCA.R

* add KE as a dependency

* Add parameters as a dependency

* whitespace change

* copy and paste result table making from demo

* Make table in PCA.R (with tinytable)

* no packages are needed for demo

@@ -3,8 +3,7 @@
 # R Script
 # Purpose: This script runs PCA and regression analysis for the demo.
 # Inputs: data/panel_data.dta
-# Outputs: save/regtable.RData
-
+# Outputs: save/regModels.RData
 
 # SETUP ----------------------------------------------------------------------------------------------------------------
 rm(list = ls()) # Clear workspace
@@ -14,9 +13,11 @@ need <- c(
   'haven',
   'AER',
   'fixest',
+  "kableExtra",
   'modelsummary',
   'corrplot',
-  "ggfortify"
+  "ggfortify",
+  "parameters"
 ) # list packages needed
 have <- need %in% rownames(installed.packages()) # checks packages you have
 if (any(!have)) install.packages(need[!have]) # install missing packages
@@ -40,7 +41,7 @@ PCA <- data |>
   prcomp(scale = T)
 summary(PCA) # show primary components
 
-plot(PCA, type = "l") # Scree plot - "elbow" shape indicates good representation by the first component 
+plot(PCA, type = "l") # Scree plot - "elbow" shape indicates good representation by the first component
 autoplot(PCA, loadings = TRUE, loadings.label = TRUE) # shows components of the first two components
 
 # Break down PCA into IMF, US, and EU components ------------------------------------------------------------------------------------------------------------------------------
@@ -68,7 +69,7 @@ plot(EU, type = "l")
 # Run regression ---------------------------------------------------------------------------------------------------------------------------------------------------------------
 
 ## Create primary component variables for regression ---------------------------------------------------------------------------------------------------------------------------
-data$imf <- -predict(IMF, data)[, 1] 
+data$imf <- -predict(IMF, data)[, 1]
 data$us <- predict(US, data)[, 1]
 data$eu <- predict(EU, data)[, 1]
 data$pca <- -predict(PCA, data)[, 1]
@@ -179,7 +180,9 @@ condition <- tobit(
 )
 summary(condition)
 
-# Save the models --------------------------------------------------------------------------------------------------------------------------------------------
+# Table generation -------------------------------------------------------------------------------------------------------------------------------------------
+
[…]

### 2025-05-18 code/PCA.R
change “t-test" comment section to f-test

@@ -195,7 +195,7 @@ save(regModels, data, file = 'save/regModels.RData')
 # load('save/regModels.RData') # Load the models again if needed
 msummary(regModels, stars = T) # Rough regression table with all variables and stars
 
-## t-tests generation -----------------------------------------------------------------------------------------------------------------------------------------
+## F-tests generation -----------------------------------------------------------------------------------------------------------------------------------------
 
 regModels$`Probit: IMF loan approval`$df.residual <- degrees_of_freedom(
   regModels$`Probit: IMF loan approval`

### 2025-05-18 manuscript/manuscript.qmd
Rename dissertation.qmd to manuscript.qmd

--- 2025-04-22 manuscript/dissertation.qmd
@@
 title: European or American? A re-examination of the transatlantic influence over the IMF
-# subtitle: GV499 Dissertation
 date: last-modified
@@
 
-[^1]: Familar name for the International Bank for Reconstruction and Development (IBRD), the lending arm of the later created World Bank Group.
+[^1]: Familiar name for the International Bank for Reconstruction and Development (IBRD), the lending arm of the later created World Bank Group.
 
@@
 
-This dissertation therefore attempts to quantitatively compare the influence of the US and Europe on the IMF.
+This paper therefore attempts to quantitatively compare the influence of the US and Europe on the IMF.
 
-The structure of this dissertation is as follows: @sec-litreview provides a review over the literature on the influence of the US and Europe over the IMF.
+The structure of this paper is as follows: @sec-litreview provides a review over the literature on the influence of the US and Europe over the IMF.
 
@@
 
+# Empirical Analysis
+
+# Case Study
+
+# Conclusion
+
 # References {.unnumbered}

### 2025-05-21 manuscript/manuscript.qmd
add execute options
Add results table to manuscript (#11)
plus move table sizing to qmd from PCA.R

* add execute options

* move resizing to demo

* add unresized regtable.rdata

* Add results table to manuscript

@@
 keywords: IMF, US hegemony, Europe
+execute:
+  echo: false
+  warning: false
+  error: false
+  message: false
 ---
@@
 
+```{r results}
+#| label: tbl-results
+#| tbl-cap: Comparative Effects of US and EU Influence on IMF Lending Outcomes (1980–2010)
+#| tbl-pos: H
+load('save/regTable.RData')
+resultsTable |> tinytable::theme_tt("resize", width = 1)
+```
+
 # Case Study

### 2025-05-21 code/PCA.R

@@ -259,7 +259,7 @@ getTable <- function(output = "html") {
 
 getTable() # for previewing the table in IDE
 
-resultsTable <- getTable("latex") |> tinytable::theme_tt("resize", width = 0.95) # LaTeX table or slide/paper presentation
+resultsTable <- getTable("latex")
 
 save(resultsTable, file = 'save/regTable.RData') # Save the table
 

### 2025-05-21 manuscript/manuscript.qmd
add hypotheses

@@
     number-sections: true
+    include-in-header:
+      - text: |
+          \newcommand{\centerstart}{\begin{center}}
+          \newcommand{\centerend}{\end{center}}
+          \usepackage{amsthm, thm-restate}
+          \newtheorem{hyp}{Hypothesis}
+          \newtheorem{subhyp}{Hypothesis}[hyp]
+          \renewcommand{\thesubhyp}{\thehyp\alph{subhyp}}
 abstract: This is an abstract.
@@
 
+\begin{restatable}[$H_{1}$]{hyp}{hypmain}
+\label{hyp-main}
+The United States and Europe exert different levels of influence on IMF lending decisions.
+\end{restatable}
+
+```{=latex}
+\setcounter{hyp}{2}
+\begin{restatable}[$H_{2a}$]{subhyp}{hypus}
+\label{hyp-us}
+The United States exerts a statistically significant influence on IMF lending decisions, controlling for European influence.
+\end{restatable}
+\vspace{-1em} % remove vertical space
+\begin{restatable}[$H_{2b}$]{subhyp}{hypeu}
+\label{hyp-eu}
+Europe exerts a statistically significant influence on IMF lending decisions, controlling for United States influence.
+\end{restatable}
+```
+
 ```{r results}

### 2025-05-21 manuscript/manuscript.qmd
add hypothese justification

@@
 
+Since the [literature review](#sec-litreview) has outlined arguments for both US and European influence over IMF decision-making, this section turns to a quantitative comparison of their respective impacts.
+Accordingly, we formulate our main hypothesis:
+
 \begin{restatable}[$H_{1}$]{hyp}{hypmain}
@@
 \end{restatable}
+
+Moreover, prior literature has either combined the United States and Europe under the umbrella of “Western” influence on IMF lending decisions [@lipscy2018], or has examined the influence of one without accounting for the other [@oatley2004].
+To address this gap, we reassess the influence of the United States and Europe separately, while controlling for the presence of the other.
+Accordingly, we formulate the following two hypotheses:
 

### 2025-05-21 manuscript/manuscript.qmd
fix citations after zotero key update

@@
 
-One renowned fact about the IMF and its sister institution, the World Bank, is the long-standing "gentlemen's agreement" of the IMF managing director being a European, and the World Bank president being an American [@Keating2024; @Weiss2023], despite no such provisions in either institution's articles of agreements [@Fund2020; @WorldBank2021].
+One renowned fact about the IMF and its sister institution, the World Bank, is the long-standing "gentlemen's agreement" of the IMF managing director being a European, and the World Bank president being an American [@keating2024; @weiss2023], despite no such provisions in either institution's articles of agreements [@fund2020; @WorldBank2021].
 
@@
 
--   IMF chief European [@Kahler2001]
+-   IMF chief European [@kahler2001]
 

### 2025-05-21 manuscript/manuscript.qmd
Add subheadings to empirical analysis
Add data header

@@
 
+## Hypotheses
+
 Since the [literature review](#sec-litreview) has outlined arguments for both US and European influence over IMF decision-making, this section turns to a quantitative comparison of their respective impacts.
@@
 
-Moreover, prior literature has either combined the United States and Europe under the umbrella of “Western” influence on IMF lending decisions [@lipscy2018], or has examined the influence of one without accounting for the other [@oatley2004].
+Moreover, prior literature has either combined the United States and Europe under the umbrella of “Western” influence on IMF lending decisions [@lipscy2018], or has examined the influence of one without accounting for the other [@oatley2004; @thacker1999].
 To address this gap, we reassess the influence of the United States and Europe separately, while controlling for the presence of the other.
@@
 ```
+
+## Data
+
+## Results
 

### 2025-05-21 code/PCA.R
add density plot

@@ -67,13 +67,23 @@ EU <- data |>
 summary(EU)
 plot(EU, type = "l")
 
-# Run regression ---------------------------------------------------------------------------------------------------------------------------------------------------------------
-
 ## Create primary component variables for regression ---------------------------------------------------------------------------------------------------------------------------
 data$imf <- -predict(IMF, data)[, 1]
 data$us <- predict(US, data)[, 1]
 data$eu <- predict(EU, data)[, 1]
 data$pca <- -predict(PCA, data)[, 1]
+
+## draw a kernel density plot for US and EU PC1 components  -----------------------------------------------------------------------------------------------------------------------------------
+data |> 
+  select(us, eu) |>
+  na.omit() |>
+  pivot_longer(cols = everything(), names_to = 'region', values_to = 'value') |>
+  ggplot(aes(x = value, fill = region)) +
+  geom_density(alpha = 0.5) +
+  labs(title = "Kernel Density Plot of US and EU variables",
+       x = "PC1",
+       y = "Density") +
+  theme_minimal()
 
 ## plot correlation between our variables with the original one. ---------------------------------------------------------------------------------------------------------------
 corr <- data |>
@@ -81,6 +91,8 @@ corr <- data |>
   cor(use = 'pairwise.complete.obs')
 
 corrplot(corr, method = 'color')
+
+# Run regression ---------------------------------------------------------------------------------------------------------------------------------------------------------------
 
 ## tobit regression - IMF loan to GDP ratio
 loan <- tobit(

### 2025-05-21 code/PCA.R
add scaling of PCA variables

@@ -68,9 +68,9 @@ summary(EU)
 plot(EU, type = "l")
 
 ## Create primary component variables for regression ---------------------------------------------------------------------------------------------------------------------------
-data$imf <- -predict(IMF, data)[, 1]
-data$us <- predict(US, data)[, 1]
-data$eu <- predict(EU, data)[, 1]
+data$imf <- -predict(IMF, data)[, 1] |> scale()
+data$us <- predict(US, data)[, 1] |> scale()
+data$eu <- predict(EU, data)[, 1] |> scale()
 data$pca <- -predict(PCA, data)[, 1]
 
 ## draw a kernel density plot for US and EU PC1 components  -----------------------------------------------------------------------------------------------------------------------------------

### 2025-05-21 code/PCA.R
Standardise PC1 (#18)
* add density plot

* add scaling of PCA variables

@@ -67,13 +67,23 @@ EU <- data |>
 summary(EU)
 plot(EU, type = "l")
 
-# Run regression ---------------------------------------------------------------------------------------------------------------------------------------------------------------
-
 ## Create primary component variables for regression ---------------------------------------------------------------------------------------------------------------------------
-data$imf <- -predict(IMF, data)[, 1]
-data$us <- predict(US, data)[, 1]
-data$eu <- predict(EU, data)[, 1]
+data$imf <- -predict(IMF, data)[, 1] |> scale()
+data$us <- predict(US, data)[, 1] |> scale()
+data$eu <- predict(EU, data)[, 1] |> scale()
 data$pca <- -predict(PCA, data)[, 1]
+
+## draw a kernel density plot for US and EU PC1 components  -----------------------------------------------------------------------------------------------------------------------------------
+data |> 
+  select(us, eu) |>
+  na.omit() |>
+  pivot_longer(cols = everything(), names_to = 'region', values_to = 'value') |>
+  ggplot(aes(x = value, fill = region)) +
+  geom_density(alpha = 0.5) +
+  labs(title = "Kernel Density Plot of US and EU variables",
+       x = "PC1",
+       y = "Density") +
+  theme_minimal()
 
 ## plot correlation between our variables with the original one. ---------------------------------------------------------------------------------------------------------------
 corr <- data |>
@@ -81,6 +91,8 @@ corr <- data |>
   cor(use = 'pairwise.complete.obs')
 
 corrplot(corr, method = 'color')
+
+# Run regression ---------------------------------------------------------------------------------------------------------------------------------------------------------------
 
 ## tobit regression - IMF loan to GDP ratio
 loan <- tobit(

### 2025-05-21 code/PCA.R
fix table presentation

@@ -213,35 +213,49 @@ regModels$`Probit: IMF loan approval`$df.residual <- degrees_of_freedom(
   regModels$`Probit: IMF loan approval`
 )
 
+get_stars <- function(p){
+  if (p < 0.001) {
+    return('***')
+  } else if (p < 0.01) {
+    return('**')
+  } else if (p < 0.05) {
+    return('*')
+  } else if (p < 0.1) {
+    return('+')
+  } else {
+    return('')
+  }
+}
+
 glance_custom.tobit <- function(x, ...) {
+  p <- lht(x, test = 'F', 'us=eu')[2, 4]
   data.frame(
     'equality' = paste0(
       '[',
-      '$p=',
-      round(lht(x, test = 'F', 'us=eu')[2, 4], 3),
-      '$]'
-    ),
-    'vcov.type' = 'by: Country'
+      round(p, 3),
+      get_stars(p),
+      ']'
+    )
   )
 }
 
 glance_custom.fixest <- function(x, ...) {
+  p <- lht(x, test = 'F', 'us=eu')[2, 4]
   data.frame(
     'equality' = paste0(
       '[',
-      '$p=',
-      round(lht(x, test = 'F', 'us=eu')[2, 4], 3),
-      '$]'
-    ),
-    'vcov.type' = 'by: Country'
+      round(p, 3),
+      get_stars(p),
+      ']'
+    )
   )
 }
 
 ## Variable renaming -------------------------------------------------------------------------------------------------------------------------------------------
 
 coefmap <- c(
-  'us' = 'US Influence',
-  'eu' = 'EU Influence',
+  'us' = 'USA Influence',
+  'eu' = 'EUP Influence',
   'imf' = 'IMF Influence',
   'lnrgdpnew' = 'GDP',
   'lnrgdpnewsq' = 'GDP$^2$',
@@ -253,9 +267,9 @@ coefmap <- c(
   '(Intercept)' = '(Intercept)'
 )
 gof_map <- list(
-  list("raw" = "equality", "clean" = "US=EU", "fmt" = NULL),
-  list("raw" = "nobs", "clean" = "$N$", "fmt" = 0),
-  list('raw' = 'vcov.type', 'clean' = 'Std.Errors', 'fmt' = NULL)
+  list("raw" = "equality", "clean" = "USA=EUP", "fmt" = NULL),
[…]

### 2025-05-21 demo/demo.qmd
move resizing to demo
fix demo citations

@@
     theme: Madrid
+    link-citations: true
     include-in-header: 
@@
 
--   "Gentlemen's agreement": the IMF managing director being a European, and the World Bank president being an American [@Keating2024; @Weiss2023].
+-   "Gentlemen's agreement": the IMF managing director being a European, and the World Bank president being an American [@keating2024; @weiss2023].
 
@@
 load('save/regTable.RData')
-resultsTable
+resultsTable |> tinytable::theme_tt("resize", width = 0.95)
 ```

### 2025-05-21 manuscript/manuscript.qmd
Add thacker to hypotheses
start data description

@@
 
+```{r}
+dta <- haven::read_dta('data/panel_data.dta') 
+```
+
+This paper uses the dataset compiled by @lipscy2018, which extends the @barro2005 dataset through 2010.
+It provides panel data for 130 countries in five-year intervals from 1980 to 2010[^2], including information on IMF borrowing and participation outcomes, economic indicators, and measures of each country's relationship with the United States and major European powers.
+
+[^2]: e.g. 1980-1985, 1985-1990.
+
 ## Results

### 2025-05-21 manuscript/manuscript.qmd
Start describing dependent variables
formatting: remove trailing line
finish dv (short of citations)

@@
 
+## Dependent Variables
+
+In line with @lipscy2018 and @barro2005, we use the following four measures of IMF lending outcomes as our dependent variables:
+
+-   IMF loan to GDP ratio
+-   IMF participation rate
+-   IMF loan to GDP approval
+-   Number of IMF conditions
+
+These variables follow the definitions used in @lipscy2018.
+The **loan-to-GDP ratio** captures the average size of IMF lending relative to a country's GDP over each five-year interval.
+The **participation rate** measures the share of months within each interval during which a country was subject to an IMF programme.
+The **approval indicator** is a binary variable coded 1 if a new IMF programme was initiated during the interval, and 0 otherwise.
+Finally, the **number of conditions** reflects the total number of performance criteria, benchmarks, and prior actions required for disbursement under IMF programmes.
+
 ## Results

### 2025-05-21 manuscript/manuscript.qmd
describe data and DVs (#19)
* Add data header

* start data description

* Start describing dependent variables

* formatting: remove trailing line

* finish dv (short of citations)

* add references to DV

--- 2025-05-21 manuscript/manuscript.qmd
@@
 
+## Dependent Variables
+
+In line with @lipscy2018 and @barro2005, we use the following four measures of IMF lending outcomes as our dependent variables:
+
+-   IMF loan to GDP ratio
+-   IMF participation rate
+-   IMF loan to GDP approval
+-   Number of IMF conditions
+
+These variables follow the definitions used in @lipscy2018.
+The *loan-to-GDP ratio* captures the average size of IMF lending relative to a country's GDP over each five-year interval.
+The *participation rate* measures the share of months within each interval during which a country was subject to an IMF programme.
+The *approval indicator* is a binary variable coded 1 if a new IMF programme was initiated during the interval, and 0 otherwise [see also @barro2005].
+Finally, the *number of conditions* reflects the total number of performance criteria, benchmarks, and prior actions required for disbursement under IMF programmes [see also @copelovitch2010; @dreher2015].
+
 ## Results

### 2025-05-21 code/PCA.R
separate staff and quota

@@ -99,7 +99,8 @@ loan <- tobit(
   imfloannew100 ~
     us +
       eu +
-      imf +
+      shstaffl +
+      shquotal +
       lnrgdpnew +
       lnrgdpnewsq +
       rgdpchnew +
@@ -124,7 +125,8 @@ part <- tobit(
   imf_p ~
     us +
       eu +
-      imf +
+      shstaffl +
+      shquotal +
       lnrgdpnew +
       lnrgdpnewsq +
       rgdpchnew +
@@ -149,7 +151,8 @@ approval <- feglm(
   imf5a ~
     us +
       eu +
-      imf +
+      shstaffl +
+      shquotal +
       lnrgdpnew +
       lnrgdpnewsq +
       rgdpchnew +
@@ -173,7 +176,8 @@ condition <- tobit(
   tc ~
     us +
       eu +
-      imf +
+      shstaffl +
+      shquotal +
       lnrgdpnew +
       lnrgdpnewsq +
       rgdpchnew +
@@ -256,7 +260,8 @@ glance_custom.fixest <- function(x, ...) {
 coefmap <- c(
   'us' = 'USA Influence',
   'eu' = 'EUP Influence',
-  'imf' = 'IMF Influence',
+  'shstaffl' = 'IMF Staff',
+  'shquotal' = 'IMF Quota',
   'lnrgdpnew' = 'GDP',
   'lnrgdpnewsq' = 'GDP$^2$',
   'rgdpchnew' = 'GDPpc',
@@ -323,6 +328,7 @@ loan_f <- tobit(
   right = Inf,
   cluster = shcode
 )
+
 ### significance of US ----
 matrix = c('frusanew = 0', 'ustradenew = 0', 'usq4 = 0')
 lht(loan_f, test = 'F', matrix)

### 2025-05-22 code/PCA.R
show period FE

@@ -251,7 +251,8 @@ glance_custom.fixest <- function(x, ...) {
       round(p, 3),
       get_stars(p),
       ']'
-    )
+    ),
+    'periodFE' = if_else("year2000" %in% names(x$coefficients), "Yes",  "")
   )
 }
 
@@ -273,6 +274,7 @@ coefmap <- c(
 )
 gof_map <- list(
   list("raw" = "equality", "clean" = "USA=EUP", "fmt" = NULL),
+  list("raw" = "periodFE", "clean" = "Period FE", "fmt" = NULL),
   list("raw" = "nobs", "clean" = "$N$", "fmt" = 0)
   # list('raw' = 'vcov.type', 'clean' = 'Std.Errors', 'fmt' = NULL)
 )

### 2025-05-22 code/PCA.R
ad missing period fe line

@@ -239,7 +239,8 @@ glance_custom.tobit <- function(x, ...) {
       round(p, 3),
       get_stars(p),
       ']'
-    )
+    ),
+    'periodFE' = if_else("year2000" %in% names(x$coefficients), "Yes",  "")
   )
 }
 

### 2025-05-22 code/PCA.R
comment out/remove IMF related PCA

@@ -45,13 +45,13 @@ summary(PCA) # show primary components
 plot(PCA, type = "l") # Scree plot - "elbow" shape indicates good representation by the first component
 autoplot(PCA, loadings = TRUE, loadings.label = TRUE) # shows components of the first two components
 
-# Break down PCA into IMF, US, and EU components ------------------------------------------------------------------------------------------------------------------------------
-IMF <- data |>
-  select(shstaffl, shquotal) |>
-  na.omit() |>
-  prcomp(scale = T)
-summary(IMF)
-plot(IMF, type = "l")
+# Break down PCA into US, and EU components ------------------------------------------------------------------------------------------------------------------------------
+# IMF <- data |>
+#   select(shstaffl, shquotal) |>
+#   na.omit() |>
+#   prcomp(scale = T)
+# summary(IMF)
+# plot(IMF, type = "l")
 
 US <- data |>
   select(frusanew, ustradenew, usq4) |>
@@ -68,7 +68,7 @@ summary(EU)
 plot(EU, type = "l")
 
 ## Create primary component variables for regression ---------------------------------------------------------------------------------------------------------------------------
-data$imf <- -predict(IMF, data)[, 1] |> scale()
+# data$imf <- -predict(IMF, data)[, 1] |> scale()
 data$us <- predict(US, data)[, 1] |> scale()
 data$eu <- predict(EU, data)[, 1] |> scale()
 data$pca <- -predict(PCA, data)[, 1]
@@ -87,7 +87,7 @@ data |>
 
 ## plot correlation between our variables with the original one. ---------------------------------------------------------------------------------------------------------------
 corr <- data |>
-  select('imf', 'us', 'eu', 'pca') |>
+  select('us', 'eu', 'pca') |>
   cor(use = 'pairwise.complete.obs')
 
 corrplot(corr, method = 'color')

### 2025-05-22 code/PCA.R
Move regression type to bottom

@@ -201,10 +201,10 @@ summary(condition)
 
 ## Save the models --------------------------------------------------------------------------------------------------------------------------------------------
 regModels <- list(
-  'Tobit: IMF loan to GDP ratio' = loan,
-  'Tobit: IMF participation rate' = part,
-  'Probit: IMF loan approval' = approval,
-  'Tobit: number of IMF conditions' = condition
+  'IMF loan to GDP ratio' = loan,
+  'IMF participation rate' = part,
+  'IMF loan approval' = approval,
+  'Number of IMF conditions' = condition
 )
 save(regModels, data, file = 'save/regModels.RData')
 
@@ -240,7 +240,8 @@ glance_custom.tobit <- function(x, ...) {
       get_stars(p),
       ']'
     ),
-    'periodFE' = if_else("year2000" %in% names(x$coefficients), "Yes",  "")
+    'periodFE' = if_else("year2000" %in% names(x$coefficients), "Yes",  ""),
+    'regression' = 'Tobit'
   )
 }
 
@@ -253,7 +254,8 @@ glance_custom.fixest <- function(x, ...) {
       get_stars(p),
       ']'
     ),
-    'periodFE' = if_else("year2000" %in% names(x$coefficients), "Yes",  "")
+    'periodFE' = if_else("year2000" %in% names(x$coefficients), "Yes",  ""),
+    'regression' = 'Probit'
   )
 }
 
@@ -276,6 +278,7 @@ coefmap <- c(
 gof_map <- list(
   list("raw" = "equality", "clean" = "USA=EUP", "fmt" = NULL),
   list("raw" = "periodFE", "clean" = "Period FE", "fmt" = NULL),
+  list("raw" = "regression", "clean" = "Regression", "fmt" = NULL),
   list("raw" = "nobs", "clean" = "$N$", "fmt" = 0)
   # list('raw' = 'vcov.type', 'clean' = 'Std.Errors', 'fmt' = NULL)
 )

### 2025-05-22 code/PCA.R
Separate staff and quota and adjust gof (#21)
* separate staff and quota

* show period FE

* adjust demo width

* ad missing period fe line

* comment out/remove IMF related PCA

* Move regression type to bottom

* Adjust demo width for new form

@@ -45,13 +45,13 @@ summary(PCA) # show primary components
 plot(PCA, type = "l") # Scree plot - "elbow" shape indicates good representation by the first component
 autoplot(PCA, loadings = TRUE, loadings.label = TRUE) # shows components of the first two components
 
-# Break down PCA into IMF, US, and EU components ------------------------------------------------------------------------------------------------------------------------------
-IMF <- data |>
-  select(shstaffl, shquotal) |>
-  na.omit() |>
-  prcomp(scale = T)
-summary(IMF)
-plot(IMF, type = "l")
+# Break down PCA into US, and EU components ------------------------------------------------------------------------------------------------------------------------------
+# IMF <- data |>
+#   select(shstaffl, shquotal) |>
+#   na.omit() |>
+#   prcomp(scale = T)
+# summary(IMF)
+# plot(IMF, type = "l")
 
 US <- data |>
   select(frusanew, ustradenew, usq4) |>
@@ -68,7 +68,7 @@ summary(EU)
 plot(EU, type = "l")
 
 ## Create primary component variables for regression ---------------------------------------------------------------------------------------------------------------------------
-data$imf <- -predict(IMF, data)[, 1] |> scale()
+# data$imf <- -predict(IMF, data)[, 1] |> scale()
 data$us <- predict(US, data)[, 1] |> scale()
 data$eu <- predict(EU, data)[, 1] |> scale()
 data$pca <- -predict(PCA, data)[, 1]
@@ -87,7 +87,7 @@ data |>
 
 ## plot correlation between our variables with the original one. ---------------------------------------------------------------------------------------------------------------
 corr <- data |>
-  select('imf', 'us', 'eu', 'pca') |>
+  select('us', 'eu', 'pca') |>
   cor(use = 'pairwise.complete.obs')
 
 corrplot(corr, method = 'color')
@@ -99,7 +99,8 @@ loan <- tobit(
   imfloannew100 ~
     us +
       eu +
-      imf +
+      shstaffl +
+      shquotal +
       lnrgdpnew +
       lnrgdpnewsq +
       rgdpchnew +
@@ -124,7 +125,8 @@ part <- tobit(
   imf_p ~
[…]

### 2025-05-22 manuscript/manuscript.qmd
extra lit review
American influence
describe PCA sources

@@
 
+## Independent Variables
+
+To assess the influence of the United States and European powers on IMF lending decisions, we construct two key independent variables that capture each actor's *propensity* to exert influence in favour of a country's borrowing from the IMF at a given time.
+These propensities are based on widely used proxies from the existing literature, which reflect the degree of a country’s political and economic alignment with the US or Europe [@lipscy2018].
+This approach assumes that both the US and European powers are more likely to advocate for lending decisions that benefit their close partners.
+
+Specifically, we construct our independent variables using three indicators commonly interpreted as channels of political and economic affinity, similar to @lipscy2018:
+
+-   UN General Assembly voting similarity[^3]
+-   Bilateral trade intensity[^4]
+-   Banking sector exposure[^5]
+
+[^3]: Originallly used in @thacker1999.
+
+[^4]: From @barro2005.
+
+[^5]: See @broz2006; @lipscy2017; @oatley2004.
+
 ## Results

### 2025-05-22 manuscript/manuscript.qmd
explain PCA

@@
 
+However, as noted by @lipscy2018 and @belsley2004, including multiple proxies for the same latent construct in a single regression model leads to multicollinearity, which reduces statistical power and inflates standard errors, as the variables are highly correlated in our data.
+To address this issue, we apply principal component analysis (PCA) to reduce the dimensionality of the data [@dunteman2008].
+Specifically, we extract the first principal component — a linear combination of the original variables that captures the greatest share of variance — to serve as a single measure of influence.
+
+Our approach differs from @lipscy2018 in that we do not pool the US and European variables together.
+Instead, we construct two separate measures: US influence and European influence.
+The PCA is conducted separately for each actor, using the same three indicators described above.
+The resulting first principal components are then normalised to have a mean of zero and a standard deviation of one, allowing for easier interpretation and comparability of coefficients in subsequent analyses.
+
 ## Results

### 2025-05-22 manuscript/manuscript.qmd
add correlation matrices

@@
           \renewcommand{\thesubhyp}{\thehyp\alph{subhyp}}
+          \usepackage{chngcntr}
+          \usepackage{apptools}
+          \AtAppendix{\counterwithin{figure}{section}}
 abstract: This is an abstract.
@@
 ```{r}
+#| label: fig-data
 dta <- haven::read_dta('data/panel_data.dta') 
@@
 
-However, as noted by @lipscy2018 and @belsley2004, including multiple proxies for the same latent construct in a single regression model leads to multicollinearity, which reduces statistical power and inflates standard errors, as the variables are highly correlated in our data.
+However, as noted by @lipscy2018 and @belsley2004, including multiple proxies for the same latent construct in a single regression model leads to multicollinearity, which reduces statistical power and inflates standard errors, as the variables are highly correlated in our data[^6].
 To address this issue, we apply principal component analysis (PCA) to reduce the dimensionality of the data [@dunteman2008].
 Specifically, we extract the first principal component — a linear combination of the original variables that captures the greatest share of variance — to serve as a single measure of influence.
+
+[^6]: See Appendix -@sec-corr.
 
@@
 
-```{r results}
+```{r}
 #| label: tbl-results
@@
 :::
+
+\appendix
+
+\newpage
+
+# Correlation Matrices {#sec-corr .appendix}
+
+```{r}
+#| label: fig-corr-eu
+#| fig-cap: Correlation Matrix of European Influence Variables
+#| fig-pos: H
+#| fig-height: 5
+load('save/corr.RData')
+corrplot::corrplot(corr_EU, method = 'color', addCoef.col = 'darkgrey', tl.col = "black")
+```
+
+```{r}
+#| label: fig-corr-us
+#| fig-cap: Correlation Matrix of US Influence Variables
+#| fig-pos: H
+#| fig-height: 5
+corrplot::corrplot(corr_US, method = 'color', addCoef.col = 'darkgrey', tl.col = "black")
+```

### 2025-05-22 code/PCA.R

@@ -26,6 +26,23 @@ invisible(lapply(need, library, character.only = T))
 
 data <- read_dta("data/panel_data.dta") # Load data
 
+# Correlation matrices --------------------------------------------------------------------------------------------------------------------------
+corr_US <- data |>
+  select('frusanew', 'ustradenew', 'usq4') |>
+  rename(Voting = 'frusanew', Trade = 'ustradenew', Bank = 'usq4') |>
+  cor(use = 'pairwise.complete.obs')
+
+corrplot(corr_US, method = 'color', addCoef.col = 'darkgrey', tl.col = "black")
+
+
+corr_EU <- data |>
+  select('frdfgnew', 'eutradenew', 'avgukfrgmq4') |>
+  rename(Voting = 'frdfgnew', Trade = 'eutradenew', Bank = 'avgukfrgmq4') |>
+  cor(use = 'pairwise.complete.obs')
+corrplot(corr_EU, method = 'color', addCoef.col = 'darkgrey', tl.col = "black")
+
+save(corr_US, corr_EU, file = "save/corr.RData")
+
 # Replicate Lipscy & Lee's (2018) original PCA analysis in R ------------------------------------------------------------------------------------------------------------------
 PCA <- data |>
   select(

### 2025-05-22 manuscript/manuscript.qmd
add scree plots

@@
 To address this issue, we apply principal component analysis (PCA) to reduce the dimensionality of the data [@dunteman2008].
-Specifically, we extract the first principal component — a linear combination of the original variables that captures the greatest share of variance — to serve as a single measure of influence.
+Specifically, we extract the first principal component — a linear combination of the original variables that captures the greatest share of variance — to serve as a single measure of influence[^7].
 
-[^6]: See Appendix -@sec-corr.
+```{r percentage_variance}
+load('save/PCA.Rdata')
+US_PC1_percentage <- US$sdev[1]^2 / sum(US$sdev^2) * 100
+EU_PC1_percentage <- EU$sdev[1]^2 / sum(EU$sdev^2) * 100
+```
+
+[^6]: See Appendix [-@sec-corr].
+
+[^7]: The first principal components capture the majority of the variance (`r format(EU_PC1_percentage, digits = 1)`% for Europe and `r format(US_PC1_percentage, digits = 1)`% for the United States).
+See Appendix [-@sec-scree] for scree plots and additional details.
 
@@
 ```
+
+# Scree Plots {#sec-scree .appendix}
+
+```{r}
+#| label: fig-scree-eu
+#| fig-cap: Scree Plot of European Influence PCA
+#| fig-pos: H
+load('save/scree.Rdata')
+scree_EU
+```
+
+```{r}
+#| label: fig-scree-us
+#| fig-cap: Scree Plot of US Influence PCA
+#| fig-pos: H
+scree_US
+```

### 2025-05-22 code/PCA.R

@@ -75,14 +75,63 @@ US <- data |>
   na.omit() |>
   prcomp(scale = T)
 summary(US)
-plot(US, type = "l")
 
 EU <- data |>
   select(frdfgnew, eutradenew, avgukfrgmq4) |>
   na.omit() |>
   prcomp(scale = T)
 summary(EU)
-plot(EU, type = "l")
+
+save(US, EU, file = "save/PCA.RData") # Save PCA results
+
+## scree plots -------------------------------------------------------------------------------------------------------------------------------------
+(scree_US <- data.frame(
+  var = US$sdev^2 / sum(US$sdev^2),
+  dim = 1:length(US$sdev)
+) |>
+  ggplot(aes(x = dim, y = var)) +
+  geom_line() +
+  geom_point() +
+  geom_text(
+    aes(label = scales::percent(var, accuracy = 0.1)),
+    vjust = -1,
+    hjust = 0.1,
+    size = 3
+  ) +
+  labs(
+    x = "Dimensions",
+    y = "Percentage of explained variances"
+  ) +
+  scale_x_continuous(breaks = 1:3, limits = c(1, 3.1)) +
+  scale_y_continuous(labels = scales::percent,
+                     limits = c(0, 0.55)) +
+  theme_bw()+
+  theme(panel.grid.minor.x = element_blank()))
+
+(scree_EU <- data.frame(
+  var = EU$sdev^2 / sum(EU$sdev^2),
+  dim = 1:length(EU$sdev)
+) |>
+  ggplot(aes(x = dim, y = var)) +
+  geom_line() +
+  geom_point() +
+  geom_text(
+    aes(label = scales::percent(var, accuracy = 0.1)),
+    vjust = -1,
+    hjust = 0.1,
+    size = 3
+  ) +
+  labs(
+    x = "Dimensions",
+    y = "Percentage of explained variances"
+  ) +
+  scale_x_continuous(breaks = 1:3, limits = c(1, 3.1)) +
+  scale_y_continuous(labels = scales::percent,
+                     limits = c(0, 0.65)) +
+  theme_bw()+
+  theme(panel.grid.minor.x = element_blank()))
+
+save(scree_US, scree_EU, file = "save/scree.RData") # Save scree plot results
 
 ## Create primary component variables for regression ---------------------------------------------------------------------------------------------------------------------------
 # data$imf <- -predict(IMF, data)[, 1] |> scale()

### 2025-05-22 manuscript/manuscript.qmd
Add covariates

@@
 
-However, as noted by @lipscy2018 and @belsley2004, including multiple proxies for the same latent construct in a single regression model leads to multicollinearity, which reduces statistical power and inflates standard errors, as the variables are highly correlated in our data[^6].
-To address this issue, we apply principal component analysis (PCA) to reduce the dimensionality of the data [@dunteman2008].
-Specifically, we extract the first principal component — a linear combination of the original variables that captures the greatest share of variance — to serve as a single measure of influence[^7].
-
 ```{r percentage_variance}
@@
 ```
+
+However, as noted by @lipscy2018 and @belsley2004, including multiple proxies for the same latent construct in a single regression model leads to multicollinearity, which reduces statistical power and inflates standard errors, as the variables are highly correlated in our data[^6].
+To address this issue, we apply principal component analysis (PCA) to reduce the dimensionality of the data [@dunteman2008].
+Specifically, we extract the first principal component — a linear combination of the original variables that captures the greatest share of variance — to serve as a single measure of influence[^7].
 
@@
 The resulting first principal components are then normalised to have a mean of zero and a standard deviation of one, allowing for easier interpretation and comparability of coefficients in subsequent analyses.
+
+## Covariates
+
+The existing literature has identified several factors beyond geopolitical alignment that influence IMF lending decisions or a country's demand for IMF assistance.
+We include these variables as covariates to account for potential confounding effects.
+Notably, countries with larger IMF quotas (and thus greater voting power), as well as those with nationals among IMF staff, are expected to receive more favourable treatment from the Fund [@barro2005].
+
+We also include economic performance indicators such as GDP, per capita GDP, and lagged GDP growth as determinants of lending and borrowing behaviour.
+International reserves as a share of imports are used to capture a country’s ability to finance its balance of payments without IMF support.
+Finally, a binary indicator for OECD membership is included to proxy for institutional and economic differences among countries.
+
+The selection of covariates, including their polynomial transformations[^8], closely follows @lipscy2018 and @barro2005 to ensure comparability with previous findings.
+
+[^8]: Quadratic terms are included for GDP and per capita GDP to capture potential non-linear effects. See [Model Specification] for details.
+
+## Model Specification {#sec-model}
 

### 2025-05-22 manuscript/manuscript.qmd
add UK, France, Germany footnote

@@
 These propensities are based on widely used proxies from the existing literature, which reflect the degree of a country’s political and economic alignment with the US or Europe [@lipscy2018].
-This approach assumes that both the US and European powers are more likely to advocate for lending decisions that benefit their close partners.
+This approach assumes that both the US and European powers[^3] are more likely to advocate for lending decisions that benefit their close partners.
+
+[^3]: For operationalisation, we focus only on ties to the three largest Western European powers—namely, the United Kingdom, France, and Germany—and use their average values.
+This approach reflects the assumption that smaller European countries are unlikely to exert meaningful influence over IMF decisions on behalf of their partners.
 
@@
 
--   UN General Assembly voting similarity[^3]
--   Bilateral trade intensity[^4]
--   Banking sector exposure[^5]
+-   UN General Assembly voting similarity[^4]
+-   Bilateral trade intensity[^5]
+-   Banking sector exposure[^6]
 
-[^3]: Originallly used in @thacker1999.
+[^4]: Originallly used in @thacker1999.
 
-[^4]: From @barro2005.
+[^5]: From @barro2005.
 
-[^5]: See @broz2006; @lipscy2017; @oatley2004.
+[^6]: See @broz2006; @lipscy2017; @oatley2004.
 
@@
 
-However, as noted by @lipscy2018 and @belsley2004, including multiple proxies for the same latent construct in a single regression model leads to multicollinearity, which reduces statistical power and inflates standard errors, as the variables are highly correlated in our data[^6].
+However, as noted by @lipscy2018 and @belsley2004, including multiple proxies for the same latent construct in a single regression model leads to multicollinearity, which reduces statistical power and inflates standard errors, as the variables are highly correlated in our data[^7].
 To address this issue, we apply principal component analysis (PCA) to reduce the dimensionality of the data [@dunteman2008].
-Specifically, we extract the first principal component — a linear combination of the original variables that captures the greatest share of variance — to serve as a single measure of influence[^7].
+Specifically, we extract the first principal component — a linear combination of the original variables that captures the greatest share of variance — to serve as a single measure of influence[^8].
 
-[^6]: See Appendix [-@sec-corr].
+[^7]: See Appendix [-@sec-corr].
 
-[^7]: The first principal components capture the majority of the variance (`r format(EU_PC1_percentage, digits = 1)`% for Europe and `r format(US_PC1_percentage, digits = 1)`% for the United States).
+[^8]: The first principal components capture the majority of the variance (`r format(EU_PC1_percentage, digits = 1)`% for Europe and `r format(US_PC1_percentage, digits = 1)`% for the United States).
 See Appendix [-@sec-scree] for scree plots and additional details.
@@
 
-The selection of covariates, including their polynomial transformations[^8], closely follows @lipscy2018 and @barro2005 to ensure comparability with previous findings.
+The selection of covariates, including their polynomial transformations[^9], closely follows @lipscy2018 and @barro2005 to ensure comparability with previous findings.
 
-[^8]: Quadratic terms are included for GDP and per capita GDP to capture potential non-linear effects. See [Model Specification] for details.
+[^9]: Quadratic terms are included for GDP and per capita GDP to capture potential non-linear effects. See [Model Specification](#sec-model) for details.
 

### 2025-05-22 code/PCA.R
Add independent variables and covariates (#22)
* describe PCA sources

* explain PCA

* ignore manuscript_files

* add correlation matrices

* add scree plots

* Add covariates

* add UK, France, Germany footnote

@@ -26,6 +26,23 @@ invisible(lapply(need, library, character.only = T))
 
 data <- read_dta("data/panel_data.dta") # Load data
 
+# Correlation matrices --------------------------------------------------------------------------------------------------------------------------
+corr_US <- data |>
+  select('frusanew', 'ustradenew', 'usq4') |>
+  rename(Voting = 'frusanew', Trade = 'ustradenew', Bank = 'usq4') |>
+  cor(use = 'pairwise.complete.obs')
+
+corrplot(corr_US, method = 'color', addCoef.col = 'darkgrey', tl.col = "black")
+
+
+corr_EU <- data |>
+  select('frdfgnew', 'eutradenew', 'avgukfrgmq4') |>
+  rename(Voting = 'frdfgnew', Trade = 'eutradenew', Bank = 'avgukfrgmq4') |>
+  cor(use = 'pairwise.complete.obs')
+corrplot(corr_EU, method = 'color', addCoef.col = 'darkgrey', tl.col = "black")
+
+save(corr_US, corr_EU, file = "save/corr.RData")
+
 # Replicate Lipscy & Lee's (2018) original PCA analysis in R ------------------------------------------------------------------------------------------------------------------
 PCA <- data |>
   select(
@@ -58,14 +75,63 @@ US <- data |>
   na.omit() |>
   prcomp(scale = T)
 summary(US)
-plot(US, type = "l")
 
 EU <- data |>
   select(frdfgnew, eutradenew, avgukfrgmq4) |>
   na.omit() |>
   prcomp(scale = T)
 summary(EU)
-plot(EU, type = "l")
+
+save(US, EU, file = "save/PCA.RData") # Save PCA results
+
+## scree plots -------------------------------------------------------------------------------------------------------------------------------------
+(scree_US <- data.frame(
+  var = US$sdev^2 / sum(US$sdev^2),
+  dim = 1:length(US$sdev)
+) |>
+  ggplot(aes(x = dim, y = var)) +
+  geom_line() +
+  geom_point() +
+  geom_text(
+    aes(label = scales::percent(var, accuracy = 0.1)),
+    vjust = -1,
+    hjust = 0.1,
+    size = 3
+  ) +
+  labs(
+    x = "Dimensions",
+    y = "Percentage of explained variances"
+  ) +
+  scale_x_continuous(breaks = 1:3, limits = c(1, 3.1)) +
+  scale_y_continuous(labels = scales::percent,
[…]

### 2025-05-22 manuscript/manuscript.qmd
Add functional forms

@@
 
+\begin{align*}
+Y^*_{it} &= \alpha + \beta_1 \text{USA}_{it} + \beta_2 \text{EUP}_{it} + \gamma^\top \mathbf{P}_{it} + \delta^\top \mathbf{Z}_{it} + \lambda_t + \epsilon_{it} \\
+Y_{it} &= \begin{cases}
+          0~\text{if}~ Y_{it}^* \leq 0 \\
+          Y_{it}^* ~\text{if}~ 0 < Y_{it}^* < c\\
+          c ~\text{if}~ Y_{it}^* \geq c \\
+          \end{cases}
+\end{align*}
+
+$$
+\mathbf{P}_{it} = \left( \ln(\text{GDP}_{it}), \ln(\text{GDP}_{it})^2, \ln(\text{PCGDP}_{it}), \ln(\text{PCGDP}_{it})^2 \right)^\top
+$$
+
+$$
+\Pr(Y_{it}=1)=\Phi(\alpha + \beta_1 \text{USA}_{it} + \beta_2 \text{EUP}_{it} + \gamma^\top \mathbf{P}_{it} + \delta^\top \mathbf{Z}_{it} + \lambda_t)
+$$
+
 ## Results

### 2025-05-22 manuscript/manuscript.qmd
change probit specification
introduce tobit

@@
 
+One notable feature of IMF loan outcomes is that they are *left-censored* at zero—that is, loan amounts, participation rates, and the number of conditions are often zero, but cannot take negative values.
+In such cases, ordinary least squares (OLS) estimation yields biased and inconsistent results.
+The Tobit model addresses this issue by explicitly modeling the latent continuous variable underlying the censored observed outcomes [@tobin1958], and is particularly well-suited to our setting [@amemiya1984].
+
 \begin{align*}
@@
 $$
-\Pr(Y_{it}=1)=\Phi(\alpha + \beta_1 \text{USA}_{it} + \beta_2 \text{EUP}_{it} + \gamma^\top \mathbf{P}_{it} + \delta^\top \mathbf{Z}_{it} + \lambda_t)
+Y_{it} = 
+\begin{cases}
+1 & \text{if } Y^*_{it} > 0 \\
+0 & \text{if } Y^*_{it} \leq 0
+\end{cases}
 $$

### 2025-05-22 manuscript/manuscript.qmd
finish tobit

@@
 In such cases, ordinary least squares (OLS) estimation yields biased and inconsistent results.
-The Tobit model addresses this issue by explicitly modeling the latent continuous variable underlying the censored observed outcomes [@tobin1958], and is particularly well-suited to our setting [@amemiya1984].
+The Tobit model addresses this issue by explicitly modelling the latent continuous variable ($Y_{it}^*$) underlying the censored observed outcomes ($Y_{it}$) [@tobin1958], and is particularly well-suited to our setting [@amemiya1984].
+
+For each non-binary dependent variable $Y_{it}$, we estimate the following Tobit model:
 
@@
 
-$$
-\mathbf{P}_{it} = \left( \ln(\text{GDP}_{it}), \ln(\text{GDP}_{it})^2, \ln(\text{PCGDP}_{it}), \ln(\text{PCGDP}_{it})^2 \right)^\top
-$$
+where:
+
+-   $Y_{it}$ is the observed outcome for country $i$ at time $t$, which may be the IMF loan-to-GDP ratio, IMF participation rate, or the number of IMF conditions.
+-   $Y^*_{it}$ is the latent continuous variable underlying the observed outcome.
+-   $c$ is the upper censoring point, set to 1 for participation rate and $\infty$ otherwise.
+-   $\text{USA}_{it}$ and $\text{EUP}_{it}$ represent the propensities for country $i$ to benefit from US and European influence at time $t$, respectively, extracted via principal component analysis.
+-   $\mathbf{P}_{it} = \left( \ln(\text{GDP}_{it}), \ln(\text{GDP}_{it})^2, \text{GDPpc}_{it}, \text{GDPpc}_{it}^2 \right)^\top$ includes logged GDP and per capita GDP, along with their squared terms.
+-   $\mathbf{Z}_{it}$ is a vector of additional covariates, including per capita GDP growth, reserves as a share of imports, and OECD membership.
+-   $\lambda_t$ denotes five-year period fixed effects.
+-   $\epsilon_{it}$ is the error term.
 

### 2025-05-22 manuscript/manuscript.qmd
probit done

@@
 -   $\lambda_t$ denotes five-year period fixed effects.
--   $\epsilon_{it}$ is the error term.
+-   $\epsilon_{it}$ is the error term, assumed to be independently and identically distributed as $\mathcal{N}(0, \sigma^2)$.
+
+For the IMF loan approval variable, which is binary ($Y_{it} \in \{0, 1\}$), we estimate a Probit model.
+This allows us to conveniently reuse the same latent variable specification $Y^*_{it}$ as in the Tobit model.
+The key differences are that the error term is now assumed to follow a standard normal distribution $\mathcal{N}(0, 1)$, and that the latent outcome maps to a binary observed outcome as follows[^10]:
+
+[^10]: Alternatively, the Probit model can be expressed in terms of probabilities as $\Pr(Y_{it}=1)=\Phi(\alpha + \beta_1 \text{USA}_{it} + \beta_2 \text{EUP}_{it} + \gamma^\top \mathbf{P}_{it} + \delta^\top \mathbf{Z}_{it} + \lambda_t)$, where $\Phi(\cdot)$ is the standard normal cumulative distribution function.
 

### 2025-05-22 manuscript/manuscript.qmd
Add model specification (#23)
* Add functional forms

* change probit specification

* introduce tobit

* finish tobit

* probit done

* Add incidental parameter problem footnote

--- 2025-05-22 manuscript/manuscript.qmd
@@
 -   $\lambda_t$ denotes five-year period fixed effects.
--   $\epsilon_{it}$ is the error term.
+-   $\epsilon_{it}$ is the error term, assumed to be independently and identically distributed as $\mathcal{N}(0, \sigma^2)$[^10].
+
+[^10]: Country fixed effects are not included because the number of time periods ($T$) is small relative to the number of countries ($N$).
+In such cases, including country fixed effects would lead to inconsistent estimates of the disturbance variance $\sigma$ [@greene2004].
+
+For the IMF loan approval variable, which is binary ($Y_{it} \in \{0, 1\}$), we estimate a Probit model.
+This allows us to conveniently reuse the same latent variable specification $Y^*_{it}$ as in the Tobit model.
+The key differences are that the error term is now assumed to follow a standard normal distribution $\mathcal{N}(0, 1)$, and that the latent outcome maps to a binary observed outcome as follows[^11]:
+
+[^11]: Alternatively, the Probit model can be expressed in terms of probabilities as $\Pr(Y_{it}=1)=\Phi(\alpha + \beta_1 \text{USA}_{it} + \beta_2 \text{EUP}_{it} + \gamma^\top \mathbf{P}_{it} + \delta^\top \mathbf{Z}_{it} + \lambda_t)$, where $\Phi(\cdot)$ is the standard normal cumulative distribution function.
 

### 2025-05-22 manuscript/manuscript.qmd
Interpret pro-EUP results

@@
 
+We now present the results from estimating the Tobit and Probit models described above, focusing on the effects of US and European influence on IMF lending outcomes.
+As shown in @tbl-results, stronger ties with Europe are significantly associated with a higher IMF loan-to-GDP ratio, a greater participation rate, and an increased likelihood of loan approval.
+In contrast, similar ties with the United States do not yield statistically significant effects on these outcomes.
+The differences between American and European influence on the first two outcomes are statistically significant at the 5% level[^12].
+
+[^12]: Based on an $F$-test for the null hypothesis that USA=EUP.
+The associated $p$-values are 0.021 and 0.014, respectively.
+For loan approval, the difference is nearly significant at the 10% level ($p=0.122$).
+
 ```{r}

### 2025-05-22 manuscript/manuscript.qmd
Add results (rough) (#24)
* Interpret pro-EUP results

* interpret conditions and link to neoliberalism

@@
 
+The only favourable treatment associated with American influence, perhaps surprisingly, is a reduction in the number of IMF conditions.
+In contrast, ties with Europe are not associated with such benefits.
+This finding echoes the argument by @lutz2014, who suggest that European actors have historically adhered more strictly to economic orthodoxy—most visibly during the Eurozone crisis—despite such policies traditionally being associated with the United States under the so-called “Washington Consensus.” Although the European adoption of neoliberal economic doctrines became especially prominent after the crisis, it is notable that similar patterns are already evident in the pre-2010 data examined here.
+
 [^12]: Based on an $F$-test for the null hypothesis that USA=EUP.

### 2025-05-27 demo/demo.qmd
adjust demo width
Adjust demo width for new form
change hypotheses

@@
 
-    -   $H_0$: No difference in the influence of US and Europe over the IMF.
-
-    -   $H_1$: Either the US or EU has more influence over the IMF.
+    -   $H_1$: Either the US or Europe has more influence over the IMF.
 
         -   Test of equality of the US and EU coefficients.
+
+    -   $H_2$: The US/Europe exerts a statistically significant influence on IMF lending decisions, controlling for European/American influence.
 
@@
 load('save/regTable.RData')
-resultsTable |> tinytable::theme_tt("resize", width = 0.95)
+resultsTable |> tinytable::theme_tt("resize", width = 0.81)
 ```

### 2025-05-27 demo/demo.qmd
Add interpretation

@@
 
+## Interpretation of Results
+
+-   $H_1$: The evidence supports the hypothesis that Europe exerts greater influence over the IMF than the United States, particularly in determining loan amounts and participation rates.
+
+-   $H_2$: The results offer a more nuanced view.
+
+    -   European influence is statistically significant in shaping the loan-to-GDP ratio, participation rate, and loan approval.
+
+    -   In contrast, US influence is **only** significant in reducing the number of IMF conditions.
+
+    -   This finding echoes @lutz2014, who argue that European actors have historically adhered more strictly to the principles of the Washington Consensus than their US counterparts.
+
 ## References {.unnumbered}

### 2025-07-11 code/PCA.R
bugfix: probit model name

@@ -279,8 +279,8 @@ msummary(regModels, stars = T) # Rough regression table with all variables and s
 
 ## F-tests generation -----------------------------------------------------------------------------------------------------------------------------------------
 
-regModels$`Probit: IMF loan approval`$df.residual <- degrees_of_freedom(
-  regModels$`Probit: IMF loan approval`
+regModels$`IMF loan approval`$df.residual <- degrees_of_freedom(
+  regModels$`IMF loan approval`
 )
 
 get_stars <- function(p){

### 2025-07-11 code/PCA.R
switch to Chisq test for probit in code

@@ -312,7 +312,7 @@ glance_custom.tobit <- function(x, ...) {
 }
 
 glance_custom.fixest <- function(x, ...) {
-  p <- lht(x, test = 'F', 'us=eu')[2, 4]
+  p <- lht(x, test = 'Chisq', 'us=eu')[2, 4]
   data.frame(
     'equality' = paste0(
       '[',

### 2025-07-11 manuscript/manuscript.qmd
Switch to Chi-squared test for Probit (#26)
* bugfix: probit model name

* switch to Chisq test for probit in code

* store new object

* Update footnote

@@
 
-[^12]: Based on an $F$-test for the null hypothesis that USA=EUP.
-The associated $p$-values are 0.021 and 0.014, respectively.
-For loan approval, the difference is nearly significant at the 10% level ($p=0.122$).
+[^12]: Based on tests of equality for the null hypothesis that $\text{USA} = \text{EUP}$.
+$F$-tests are used for the linear models; for the Probit model, the test corresponds to an asymptotic Chi-squared statistic, which is appropriate given the large sample size ($N = 522$).
+The $p$-values for the first two models are 0.021 and 0.014, respectively.
+For loan approval, the difference is nearly significant at the 10% level ($p = 0.122$).
 

### 2025-07-11 code/PCA.R

@@ -279,8 +279,8 @@ msummary(regModels, stars = T) # Rough regression table with all variables and s
 
 ## F-tests generation -----------------------------------------------------------------------------------------------------------------------------------------
 
-regModels$`Probit: IMF loan approval`$df.residual <- degrees_of_freedom(
-  regModels$`Probit: IMF loan approval`
+regModels$`IMF loan approval`$df.residual <- degrees_of_freedom(
+  regModels$`IMF loan approval`
 )
 
 get_stars <- function(p){
@@ -312,7 +312,7 @@ glance_custom.tobit <- function(x, ...) {
 }
 
 glance_custom.fixest <- function(x, ...) {
-  p <- lht(x, test = 'F', 'us=eu')[2, 4]
+  p <- lht(x, test = 'Chisq', 'us=eu')[2, 4]
   data.frame(
     'equality' = paste0(
       '[',

### 2025-07-11 code/continent.R
add continent to rds

@@ -1,0 +1,40 @@
+######## INFO ########
+# Authors: Dianyi Yang
+# R Script
+# Purpose: This script runs PCA and regression analysis for the demo.
+# Inputs:  data/panel_data_raw.dta
+#          data/continents.json
+# Outputs: data/panel_data.rds
+
+# SETUP ----------------------------------------------------------------------------------------------------------------
+rm(list = ls()) # Clear workspace
+
+need <- c(
+  'tidyverse',
+  'jsonlite'
+) # list packages needed
+have <- need %in% rownames(installed.packages()) # checks packages you have
+if (any(!have)) install.packages(need[!have]) # install missing packages
+invisible(lapply(need, library, character.only = T))
+
+# Read data ----------------------------------------------------------------------------------------------------------------
+raw_data <- read_dta("data/panel_data_raw.dta")
+continent_map <- fromJSON("data/continents.json")
+
+# Check coverage of countries in the map -----------------------------------------------------------------------------------------------
+continent_df <- data.frame(
+  country = names(continent_map),
+  continent = unname(unlist(continent_map))
+)
+
+country_list <- unique(raw_data$country)
+country_list[!(country_list %in% continent_df$country)]
+  ## character(0) # No missing countries
+
+# Merge data with continent map -----------------------------------------------------------------------------------------------
+data <- raw_data |> 
+  left_join(continent_df, by = "country") |> 
+  relocate(continent, .after = isocode)
+
+# Save data ------------------------------------------------------------------------------------------------------------------
+write_rds(data, "data/panel_data.rds")

### 2025-07-11 code/continent.R
add description

@@ -1,7 +1,7 @@
 ######## INFO ########
 # Authors: Dianyi Yang
 # R Script
-# Purpose: This script runs PCA and regression analysis for the demo.
+# Purpose: This script adds continent information to the panel data.
 # Inputs:  data/panel_data_raw.dta
 #          data/continents.json
 # Outputs: data/panel_data.rds

### 2025-07-11 code/PCA.R
Add continent info (#27)
* GPT-categorized

* rename data to raw

* add continent to rds

* PCA.R: switch to using new data with continent info

* add description

@@ -2,7 +2,7 @@
 # Authors: Dianyi Yang
 # R Script
 # Purpose: This script runs PCA and regression analysis for the demo.
-# Inputs:  data/panel_data.dta
+# Inputs:  data/panel_data.rds
 # Outputs: save/regModels.RData
 #          save/regTable.RData
 
@@ -24,7 +24,7 @@ have <- need %in% rownames(installed.packages()) # checks packages you have
 if (any(!have)) install.packages(need[!have]) # install missing packages
 invisible(lapply(need, library, character.only = T))
 
-data <- read_dta("data/panel_data.dta") # Load data
+data <- read_rds("data/panel_data.rds") # Load data
 
 # Correlation matrices --------------------------------------------------------------------------------------------------------------------------
 corr_US <- data |>

### 2025-07-11 code/continent.R

@@ -1,0 +1,40 @@
+######## INFO ########
+# Authors: Dianyi Yang
+# R Script
+# Purpose: This script adds continent information to the panel data.
+# Inputs:  data/panel_data_raw.dta
+#          data/continents.json
+# Outputs: data/panel_data.rds
+
+# SETUP ----------------------------------------------------------------------------------------------------------------
+rm(list = ls()) # Clear workspace
+
+need <- c(
+  'tidyverse',
+  'jsonlite'
+) # list packages needed
+have <- need %in% rownames(installed.packages()) # checks packages you have
+if (any(!have)) install.packages(need[!have]) # install missing packages
+invisible(lapply(need, library, character.only = T))
+
+# Read data ----------------------------------------------------------------------------------------------------------------
+raw_data <- read_dta("data/panel_data_raw.dta")
+continent_map <- fromJSON("data/continents.json")
+
+# Check coverage of countries in the map -----------------------------------------------------------------------------------------------
+continent_df <- data.frame(
+  country = names(continent_map),
+  continent = unname(unlist(continent_map))
+)
+
+country_list <- unique(raw_data$country)
+country_list[!(country_list %in% continent_df$country)]
+  ## character(0) # No missing countries
+
+# Merge data with continent map -----------------------------------------------------------------------------------------------
+data <- raw_data |> 
+  left_join(continent_df, by = "country") |> 
+  relocate(continent, .after = isocode)
+
+# Save data ------------------------------------------------------------------------------------------------------------------
+write_rds(data, "data/panel_data.rds")

### 2025-07-12 code/PCA.R
PCA: add missing divider

@@ -166,11 +166,11 @@ corrplot(corr, method = 'color')
 
 # Run regression ---------------------------------------------------------------------------------------------------------------------------------------------------------------
 
-## tobit regression - IMF loan to GDP ratio
+## tobit regression - IMF loan to GDP ratio ----------------------------------------------------------------------------------------------------
 loan <- tobit(
   imfloannew100 ~
     us +
-      eu +
+      eu*is_european +
       shstaffl +
       shquotal +
       lnrgdpnew +
@@ -196,7 +196,7 @@ summary(loan)
 part <- tobit(
   imf_p ~
     us +
-      eu +
+      eu*is_european +
       shstaffl +
       shquotal +
       lnrgdpnew +
@@ -222,7 +222,7 @@ summary(part)
 approval <- feglm(
   imf5a ~
     us +
-      eu +
+      eu*is_european +
       shstaffl +
       shquotal +
       lnrgdpnew +
@@ -247,7 +247,7 @@ summary(approval)
 condition <- tobit(
   tc ~
     us +
-      eu +
+      eu*is_european +
       shstaffl +
       shquotal +
       lnrgdpnew +

### 2025-07-12 code/PCA.R
PCA: reuse variables

@@ -166,25 +166,21 @@ corrplot(corr, method = 'color')
 
 # Run regression ---------------------------------------------------------------------------------------------------------------------------------------------------------------
 
+## Prepare variables -----------------------------------------------------------------------------------------------------
+rhs <- c(
+  "us", "eu", "shstaffl", "shquotal", 
+  "lnrgdpnew", "lnrgdpnewsq", 
+  "rgdpchnew", "rgdpchnewsquare", 
+  "growth1new", "reserv1", "oecd1", 
+  "year1980", "year1985", "year1990", "year1995", "year2000"
+)
+make_formula <- function(outcome) {
+  as.formula(paste(outcome, paste(rhs, collapse = " + "), sep = " ~ "))
+}
+
 ## tobit regression - IMF loan to GDP ratio ----------------------------------------------------------------------------------------------------
 loan <- tobit(
-  imfloannew100 ~
-    us +
-      eu*is_european +
-      shstaffl +
-      shquotal +
-      lnrgdpnew +
-      lnrgdpnewsq +
-      rgdpchnew +
-      rgdpchnewsquare +
-      growth1new +
-      reserv1 +
-      oecd1 +
-      year1980 +
-      year1985 +
-      year1990 +
-      year1995 +
-      year2000,
+  make_formula("imfloannew100"),
   data = data,
   left = 0,
   right = Inf,
@@ -194,23 +190,7 @@ summary(loan)
 
 ## Tobit regression - IMF participation rate ----------------------------------------------------------------------------------------------------------------
 part <- tobit(
-  imf_p ~
-    us +
-      eu*is_european +
-      shstaffl +
-      shquotal +
-      lnrgdpnew +
-      lnrgdpnewsq +
-      rgdpchnew +
-      rgdpchnewsquare +
-      growth1new +
-      reserv1 +
-      oecd1 +
-      year1980 +
-      year1985 +
-      year1990 +
-      year1995 +
-      year2000,
+  make_formula("imf_p"),
   data = data,
   left = 0,
   right = 1,
@@ -220,23 +200,7 @@ summary(part)
 
 ## Probit regression - IMF loan approval --------------------------------------------------------------------------------------------------------------------
 approval <- feglm(
-  imf5a ~
-    us +
-      eu*is_european +
-      shstaffl +
-      shquotal +
-      lnrgdpnew +
-      lnrgdpnewsq +
-      rgdpchnew +
[…]

### 2025-07-12 code/PCA.R
PCA: use msummary instead of summary

@@ -186,7 +186,7 @@ loan <- tobit(
   right = Inf,
   cluster = shcode
 )
-summary(loan)
+msummary(loan, stars = TRUE)
 
 ## Tobit regression - IMF participation rate ----------------------------------------------------------------------------------------------------------------
 part <- tobit(
@@ -196,7 +196,7 @@ part <- tobit(
   right = 1,
   cluster = shcode
 )
-summary(part)
+msummary(part, stars = TRUE)
 
 ## Probit regression - IMF loan approval --------------------------------------------------------------------------------------------------------------------
 approval <- feglm(
@@ -205,7 +205,7 @@ approval <- feglm(
   family = binomial(link = "probit"),
   cluster = 'shcode'
 )
-summary(approval)
+msummary(approval, stars = TRUE)
 
 # Tobit Total number of IMF conditions ----------------------------------------------------------------------------------------------------------------------
 condition <- tobit(
@@ -215,7 +215,7 @@ condition <- tobit(
   right = Inf,
   cluster = shcode
 )
-summary(condition)
+msummary(condition, stars = TRUE)
 
 # Table generation -------------------------------------------------------------------------------------------------------------------------------------------
 

### 2025-07-12 code/is_Europe.R
is_europe: use data with pca

@@ -2,7 +2,7 @@
 # Authors: Dianyi Yang
 # R Script
 # Purpose: This script runs separate regressions for European and non-European countries.
-# Inputs:  data/panel_data.rds
+# Inputs:  data/panel_data_pca.rds
 # Outputs: 
 
 # SETUP ----------------------------------------------------------------------------------------------------------------
@@ -20,7 +20,7 @@ have <- need %in% rownames(installed.packages()) # checks packages you have
 if (any(!have)) install.packages(need[!have]) # install missing packages
 invisible(lapply(need, library, character.only = T))
 
-data <- read_rds("data/panel_data.rds") # Load data
+data <- read_rds("data/panel_data_pca.rds") # Load data
 
 # Generate European dummy variable ------------------------------------------------------------------------------------------------
 data$is_european <- ifelse(data$continent == "Europe", 1, 0)

### 2025-07-12 code/is_Europe.R
is_europe: reuse variables

@@ -29,7 +29,7 @@ data$is_european <- ifelse(data$continent == "Europe", 1, 0)
 
 ## Prepare variables -----------------------------------------------------------------------------------------------------
 rhs <- c(
-  "us", "eu", "shstaffl", "shquotal", 
+  "us", "eu*is_european", "shstaffl", "shquotal", 
   "lnrgdpnew", "lnrgdpnewsq", 
   "rgdpchnew", "rgdpchnewsquare", 
   "growth1new", "reserv1", "oecd1", 
@@ -51,23 +51,7 @@ summary(loan)
 
 ## Tobit regression - IMF participation rate ----------------------------------------------------------------------------------------------------------------
 part <- tobit(
-  imf_p ~
-    us +
-      eu +
-      shstaffl +
-      shquotal +
-      lnrgdpnew +
-      lnrgdpnewsq +
-      rgdpchnew +
-      rgdpchnewsquare +
-      growth1new +
-      reserv1 +
-      oecd1 +
-      year1980 +
-      year1985 +
-      year1990 +
-      year1995 +
-      year2000,
+  make_formula("imf_p"),
   data = data,
   left = 0,
   right = 1,
@@ -77,23 +61,7 @@ summary(part)
 
 ## Probit regression - IMF loan approval --------------------------------------------------------------------------------------------------------------------
 approval <- feglm(
-  imf5a ~
-    us +
-      eu +
-      shstaffl +
-      shquotal +
-      lnrgdpnew +
-      lnrgdpnewsq +
-      rgdpchnew +
-      rgdpchnewsquare +
-      growth1new +
-      reserv1 +
-      oecd1 +
-      year1980 +
-      year1985 +
-      year1990 +
-      year1995 +
-      year2000,
+  make_formula("imf5a"),
   data = data,
   family = binomial(link = "probit"),
   cluster = 'shcode'
@@ -102,23 +70,7 @@ summary(approval)
 
 # Tobit Total number of IMF conditions ----------------------------------------------------------------------------------------------------------------------
 condition <- tobit(
-  tc ~
-    us +
-      eu +
-      shstaffl +
-      shquotal +
-      lnrgdpnew +
-      lnrgdpnewsq +
-      rgdpchnew +
-      rgdpchnewsquare +
-      growth1new +
-      reserv1 +
-      oecd1 +
-      year1980 +
-      year1985 +
-      year1990 +
-      year1995 +
-      year2000,
+  make_formula("tc"),
   data = data,
   left = 0,
   right = Inf,

### 2025-07-12 code/is_Europe.R
is_europe: use msummary instead of summary

@@ -47,7 +47,8 @@ loan <- tobit(
   right = Inf,
   cluster = shcode
 )
-summary(loan)
+
+msummary(loan, stars = TRUE)
 
 ## Tobit regression - IMF participation rate ----------------------------------------------------------------------------------------------------------------
 part <- tobit(
@@ -57,7 +58,7 @@ part <- tobit(
   right = 1,
   cluster = shcode
 )
-summary(part)
+msummary(part, stars = TRUE)
 
 ## Probit regression - IMF loan approval --------------------------------------------------------------------------------------------------------------------
 approval <- feglm(
@@ -66,7 +67,7 @@ approval <- feglm(
   family = binomial(link = "probit"),
   cluster = 'shcode'
 )
-summary(approval)
+msummary(approval, stars = TRUE)
 
 # Tobit Total number of IMF conditions ----------------------------------------------------------------------------------------------------------------------
 condition <- tobit(
@@ -76,4 +77,4 @@ condition <- tobit(
   right = Inf,
   cluster = shcode
 )
-summary(condition)
+msummary(condition, stars = TRUE)

### 2025-07-12 code/is_Europe.R
is_europea: save models

@@ -78,3 +78,17 @@ condition <- tobit(
   cluster = shcode
 )
 msummary(condition, stars = TRUE)
+
+# Table generation -------------------------------------------------------------------------------------------------------------------------------------------
+
+## Save the models --------------------------------------------------------------------------------------------------------------------------------------------
+regModels <- list(
+  'IMF loan to GDP ratio' = loan,
+  'IMF participation rate' = part,
+  'IMF loan approval' = approval,
+  'Number of IMF conditions' = condition
+)
+save(regModels, file = 'save/regModels_europe.RData')
+
+# load('save/regModels.RData') # Load the models again if needed
+msummary(regModels, stars = T) # Rough regression table with all variables and stars

### 2025-07-12 code/PCA.R
move f test generation to helper

@@ -233,51 +233,7 @@ msummary(regModels, stars = T) # Rough regression table with all variables and s
 
 ## F-tests generation -----------------------------------------------------------------------------------------------------------------------------------------
 
-regModels$`IMF loan approval`$df.residual <- degrees_of_freedom(
-  regModels$`IMF loan approval`
-)
-
-get_stars <- function(p){
-  if (p < 0.001) {
-    return('***')
-  } else if (p < 0.01) {
-    return('**')
-  } else if (p < 0.05) {
-    return('*')
-  } else if (p < 0.1) {
-    return('+')
-  } else {
-    return('')
-  }
-}
-
-glance_custom.tobit <- function(x, ...) {
-  p <- lht(x, test = 'F', 'us=eu')[2, 4]
-  data.frame(
-    'equality' = paste0(
-      '[',
-      round(p, 3),
-      get_stars(p),
-      ']'
-    ),
-    'periodFE' = if_else("year2000" %in% names(x$coefficients), "Yes",  ""),
-    'regression' = 'Tobit'
-  )
-}
-
-glance_custom.fixest <- function(x, ...) {
-  p <- lht(x, test = 'Chisq', 'us=eu')[2, 4]
-  data.frame(
-    'equality' = paste0(
-      '[',
-      round(p, 3),
-      get_stars(p),
-      ']'
-    ),
-    'periodFE' = if_else("year2000" %in% names(x$coefficients), "Yes",  ""),
-    'regression' = 'Probit'
-  )
-}
+source("helper/f_test.R")
 
 ## Variable renaming -------------------------------------------------------------------------------------------------------------------------------------------
 

### 2025-07-12 code/is_Europe.R
save european regression results

@@ -92,3 +92,48 @@ save(regModels, file = 'save/regModels_europe.RData')
 
 # load('save/regModels.RData') # Load the models again if needed
 msummary(regModels, stars = T) # Rough regression table with all variables and stars
+
+## Variable renaming -------------------------------------------------------------------------------------------------------------------------------------------
+
+coefmap <- c(
+  'us' = 'USA Influence',
+  'eu' = 'EUP Influence',
+  'eu:is_european' = 'EUP Influence x European recipient',
+  'is_european' = 'European recipient',
+  'shstaffl' = 'IMF Staff',
+  'shquotal' = 'IMF Quota',
+  'lnrgdpnew' = 'GDP',
+  'lnrgdpnewsq' = 'GDP$^2$',
+  'rgdpchnew' = 'GDPpc',
+  'rgdpchnewsquare' = "GDPpc$^2$",
+  'growth1new' = 'GDPpc growth',
+  'reserv1' = 'Reserves',
+  'oecd1' = 'OECD',
+  '(Intercept)' = '(Intercept)'
+)
+gof_map <- list(
+  list("raw" = "periodFE", "clean" = "Period FE", "fmt" = NULL),
+  list("raw" = "regression", "clean" = "Regression", "fmt" = NULL),
+  list("raw" = "nobs", "clean" = "$N$", "fmt" = 0)
+  # list('raw' = 'vcov.type', 'clean' = 'Std.Errors', 'fmt' = NULL)
+)
+
+getTable <- function(output = "html") {
+  msummary(
+    regModels,
+    coef_map = coefmap,
+    gof_map = gof_map,
+    stars = T,
+    escape = F,
+    output = output,
+    notes = c(
+      "Standard errors clustered at the country level."
+    )
+  )
+}
+
+getTable() # for previewing the table in IDE
+
+resultsTable <- getTable("latex")
+
+save(resultsTable, file = 'save/regTable_europe.RData') # Save the table

### 2025-07-12 code/PCA.R
Add regressions which separate European recipient from non-European ones (#28)
* PCA.R: add missing output info

* add is_europe draft

* PCA add panel_data_pca.rds as output

* Add corresponding changes in PCA.R

* PCA: add missing divider

* PCA: reuse variables

* PCA: use msummary instead of summary

* is_europe: use data with pca

* is_europe: reuse variables

* is_europe: use msummary instead of summary

* is_europea: save models

* move f test generation to helper

* save european regression results

@@ -3,7 +3,11 @@
 # R Script
 # Purpose: This script runs PCA and regression analysis for the demo.
 # Inputs:  data/panel_data.rds
-# Outputs: save/regModels.RData
+# Outputs: save/corr.RData
+#          save/PCA.RData
+#          save/scree.RData
+#          data/panel_data_pca.rds
+#          save/regModels.RData
 #          save/regTable.RData
 
 # SETUP ----------------------------------------------------------------------------------------------------------------
@@ -139,6 +143,8 @@ data$us <- predict(US, data)[, 1] |> scale()
 data$eu <- predict(EU, data)[, 1] |> scale()
 data$pca <- -predict(PCA, data)[, 1]
 
+write_rds(data, "data/panel_data_pca.rds")
+
 ## draw a kernel density plot for US and EU PC1 components  -----------------------------------------------------------------------------------------------------------------------------------
 data |> 
   select(us, eu) |>
@@ -160,108 +166,56 @@ corrplot(corr, method = 'color')
 
 # Run regression ---------------------------------------------------------------------------------------------------------------------------------------------------------------
 
-## tobit regression - IMF loan to GDP ratio
+## Prepare variables -----------------------------------------------------------------------------------------------------
+rhs <- c(
+  "us", "eu", "shstaffl", "shquotal", 
+  "lnrgdpnew", "lnrgdpnewsq", 
+  "rgdpchnew", "rgdpchnewsquare", 
+  "growth1new", "reserv1", "oecd1", 
+  "year1980", "year1985", "year1990", "year1995", "year2000"
+)
+make_formula <- function(outcome) {
+  as.formula(paste(outcome, paste(rhs, collapse = " + "), sep = " ~ "))
+}
+
+## tobit regression - IMF loan to GDP ratio ----------------------------------------------------------------------------------------------------
 loan <- tobit(
-  imfloannew100 ~
-    us +
-      eu +
-      shstaffl +
-      shquotal +
-      lnrgdpnew +
-      lnrgdpnewsq +
-      rgdpchnew +
-      rgdpchnewsquare +
-      growth1new +
-      reserv1 +
-      oecd1 +
-      year1980 +
-      year1985 +
-      year1990 +
-      year1995 +
-      year2000,
+  make_formula("imfloannew100"),
   data = data,
   left = 0,
   right = Inf,
   cluster = shcode
 )
-summary(loan)
[…]

### 2025-07-12 code/is_Europe.R

@@ -1,0 +1,139 @@
+######## INFO ########
+# Authors: Dianyi Yang
+# R Script
+# Purpose: This script runs separate regressions for European and non-European countries.
+# Inputs:  data/panel_data_pca.rds
+# Outputs: 
+
+# SETUP ----------------------------------------------------------------------------------------------------------------
+rm(list = ls()) # Clear workspace
+
+need <- c(
+  'tidyverse',
+  'AER',
+  'fixest',
+  "kableExtra",
+  'modelsummary',
+  "parameters"
+) # list packages needed
+have <- need %in% rownames(installed.packages()) # checks packages you have
+if (any(!have)) install.packages(need[!have]) # install missing packages
+invisible(lapply(need, library, character.only = T))
+
+data <- read_rds("data/panel_data_pca.rds") # Load data
+
+# Generate European dummy variable ------------------------------------------------------------------------------------------------
+data$is_european <- ifelse(data$continent == "Europe", 1, 0)
+
+# Run regression ---------------------------------------------------------------------------------------------------------------------------------------------------------------
+
+## Prepare variables -----------------------------------------------------------------------------------------------------
+rhs <- c(
+  "us", "eu*is_european", "shstaffl", "shquotal", 
+  "lnrgdpnew", "lnrgdpnewsq", 
+  "rgdpchnew", "rgdpchnewsquare", 
+  "growth1new", "reserv1", "oecd1", 
+  "year1980", "year1985", "year1990", "year1995", "year2000"
+)
+make_formula <- function(outcome) {
+  as.formula(paste(outcome, paste(rhs, collapse = " + "), sep = " ~ "))
+}
+
+## tobit regression - IMF loan to GDP ratio --------------------------------------------------------------------------------------------------------------------
+loan <- tobit(
+  make_formula("imfloannew100"),
+  data = data,
+  left = 0,
+  right = Inf,
+  cluster = shcode
+)
+
+msummary(loan, stars = TRUE)
+
+## Tobit regression - IMF participation rate ----------------------------------------------------------------------------------------------------------------
+part <- tobit(
+  make_formula("imf_p"),
+  data = data,
+  left = 0,
+  right = 1,
+  cluster = shcode
+)
+msummary(part, stars = TRUE)
+
[…]

### 2025-07-12 helper/f_test.R

@@ -1,0 +1,45 @@
+regModels$`IMF loan approval`$df.residual <- degrees_of_freedom(
+  regModels$`IMF loan approval`
+)
+
+get_stars <- function(p){
+  if (p < 0.001) {
+    return('***')
+  } else if (p < 0.01) {
+    return('**')
+  } else if (p < 0.05) {
+    return('*')
+  } else if (p < 0.1) {
+    return('+')
+  } else {
+    return('')
+  }
+}
+
+glance_custom.tobit <- function(x, ...) {
+  p <- lht(x, test = 'F', 'us=eu')[2, 4]
+  data.frame(
+    'equality' = paste0(
+      '[',
+      round(p, 3),
+      get_stars(p),
+      ']'
+    ),
+    'periodFE' = if_else("year2000" %in% names(x$coefficients), "Yes",  ""),
+    'regression' = 'Tobit'
+  )
+}
+
+glance_custom.fixest <- function(x, ...) {
+  p <- lht(x, test = 'Chisq', 'us=eu')[2, 4]
+  data.frame(
+    'equality' = paste0(
+      '[',
+      round(p, 3),
+      get_stars(p),
+      ']'
+    ),
+    'periodFE' = if_else("year2000" %in% names(x$coefficients), "Yes",  ""),
+    'regression' = 'Probit'
+  )
+}

### 2025-07-12 code/PCA.R
do not save data in regModels

@@ -226,7 +226,7 @@ regModels <- list(
   'IMF loan approval' = approval,
   'Number of IMF conditions' = condition
 )
-save(regModels, data, file = 'save/regModels.RData')
+save(regModels, file = 'save/regModels.RData')
 
 # load('save/regModels.RData') # Load the models again if needed
 msummary(regModels, stars = T) # Rough regression table with all variables and stars
@@ -268,7 +268,7 @@ getTable <- function(output = "html") {
     escape = F,
     output = output,
     notes = c(
-      "Standard errors clustered at the country level. Values in square brackets represent $p$-values from $F$-tests."
+      "Standard errors clustered at the country level. Values in square brackets represent $p$-values from $F$-tests (linear models) and Chi-squared tests (Probit model)."
     )
   )
 }

### 2025-07-12 manuscript/manuscript.qmd
Bugfix: read RDS data rather than dta
add europe-disaggregated table

@@
 #| label: fig-data
-dta <- haven::read_dta('data/panel_data.dta')
+dta <- readr::read_rds('data/panel_data.rds')
 ```
@@
 
+```{r}
+#| label: tbl-results-europe
+#| tbl-cap: Effects of U.S. and European Influence on IMF Lending Outcomes (1980–2010), Separating European and Non-European Recipients
+#| tbl-pos: H
+load('save/regTable_europe.RData')
+resultsTable |> tinytable::theme_tt("resize", width = 1)
+```
+
 # Case Study

### 2025-07-14 manuscript/manuscript.qmd
literaturereview_writing1.1
explains for macroeconomic factors as covariates.

--- 2025-07-11 manuscript/manuscript.qmd
@@
   message: false
+bibliography: references.bib
 ---
@@
 
--   IMF chief European [@kahler2001]
+As an international organization formally governed by a system of weighted voting and technocratic procedures, IMF is viewed as an “accountant” [@lutz2019], primarily concerned with the internal consistency of fiscal projections to guarantee timely repayment of its financial assistance.
+However, empirical studies suggested that countries with low international reserves, high current account deficits, and low GDP growth are more likely to receive IMF funding assistance [@barro2005, @knight1997] , despite such indicators often reflecting a limited capacity to repay.
+This paradox leads to a widely argued statement, particularly among realists, that international institutions are “basically a reflection of the distribution of power in the world” [@mearsheimer1994].
+Generally, all studies on IMF lending outcomes agree that major power impose their influence in the process.
+For example, studies have shown that countries strategically important to the United States are more likely to receive IMF loans, or to receive larger amounts.
+Specifically, this applies to countries that align with the U.S.
+in United Nations General Assembly and United Nations Security Council voting [@lipscy2018, @thacker1999, @dreher2009], are in geopolitical vital regions \[\], or hold significant debts to American commercial banks [@oatley2004].
+This is not surprising, as the United States, as the most powerful state, has long projected its influence over international financial institutions—much like it has done with the World Bank [@wade2022, @woods2023, @clark2020].
+The question this study brings about is, how this reflection of power is distributed within a given international organization, especially one that is widely perceived to be dominated by a few major powers.
+
+The projection of great power capabilities into international organizations is sometimes uneven, because of strategic preferences, the perceived inefficiency of institutional engagement, or other possible reasons.
+Existing studies provide substantial evidence of U.S.
+influence over IMF lending decisions, they leave a crucial question open-the comparison between the U.S.
+and other major powers, particularly the European countries.
+IMF is perceived to be an institution dominated by a few major power, but the relative weight of American versus European remains ambiguous.
+The organization is traditionally chaired by an European national[@kahler2001],
 
@@
 
-The only favourable treatment associated with American influence, perhaps surprisingly, is a reduction in the number of IMF conditions.
-In contrast, ties with Europe are not associated with such benefits.
-This finding echoes the argument by @lutz2014, who suggest that European actors have historically adhered more strictly to economic orthodoxy—most visibly during the Eurozone crisis—despite such policies traditionally being associated with the United States under the so-called “Washington Consensus.” Although the European adoption of neoliberal economic doctrines became especially prominent after the crisis, it is notable that similar patterns are already evident in the pre-2010 data examined here.
-
 [^12]: Based on tests of equality for the null hypothesis that $\text{USA} = \text{EUP}$.
@@
 For loan approval, the difference is nearly significant at the 10% level ($p = 0.122$).
+
+The only favourable treatment associated with American influence, perhaps surprisingly, is a reduction in the number of IMF conditions.
+In contrast, ties with Europe are not associated with such benefits.
+This finding echoes the argument by @lutz2014, who suggest that European actors have historically adhered more strictly to economic orthodoxy—most visibly during the Eurozone crisis—despite such policies traditionally being associated with the United States under the so-called “Washington Consensus.” Although the European adoption of neoliberal economic doctrines became especially prominent after the crisis, it is notable that similar patterns are already evident in the pre-2010 data examined here.
 

### 2025-07-14 manuscript/manuscript.qmd
Share references.bib with demo
Fix citations

@@
   message: false
-bibliography: references.bib
 ---
@@
 As an international organization formally governed by a system of weighted voting and technocratic procedures, IMF is viewed as an “accountant” [@lutz2019], primarily concerned with the internal consistency of fiscal projections to guarantee timely repayment of its financial assistance.
-However, empirical studies suggested that countries with low international reserves, high current account deficits, and low GDP growth are more likely to receive IMF funding assistance [@barro2005, @knight1997] , despite such indicators often reflecting a limited capacity to repay.
+However, empirical studies suggested that countries with low international reserves, high current account deficits, and low GDP growth are more likely to receive IMF funding assistance [@barro2005; @knight1997], despite such indicators often reflecting a limited capacity to repay.
 This paradox leads to a widely argued statement, particularly among realists, that international institutions are “basically a reflection of the distribution of power in the world” [@mearsheimer1994].
@@
 Specifically, this applies to countries that align with the U.S.
-in United Nations General Assembly and United Nations Security Council voting [@lipscy2018, @thacker1999, @dreher2009], are in geopolitical vital regions \[\], or hold significant debts to American commercial banks [@oatley2004].
-This is not surprising, as the United States, as the most powerful state, has long projected its influence over international financial institutions—much like it has done with the World Bank [@wade2022, @woods2023, @clark2020].
+in United Nations General Assembly and United Nations Security Council voting [@lipscy2018; @thacker1999; @dreher2009], are in geopolitical vital regions \[\], or hold significant debts to American commercial banks [@oatley2004].
+This is not surprising, as the United States, as the most powerful state, has long projected its influence over international financial institutions—much like it has done with the World Bank [@wade2002; @woods2003; @clark2020].
 The question this study brings about is, how this reflection of power is distributed within a given international organization, especially one that is widely perceived to be dominated by a few major powers.
@@
 IMF is perceived to be an institution dominated by a few major power, but the relative weight of American versus European remains ambiguous.
-The organization is traditionally chaired by an European national[@kahler2001],
+The organization is traditionally chaired by an European national [@kahler2001],
 

### 2025-07-15 manuscript/manuscript.qmd
Add robustness draft

--- 2025-07-12 manuscript/manuscript.qmd
@@
 
+## Robustness: Continental Preference
+
+One plausible alternative explanation for the strong association between European influence and favourable IMF lending outcomes is that European powers might be disproportionately favouring fellow European recipients.
+To evaluate this possibility, we examine whether the effect of European influence differs significantly between European and non-European countries.
+
+We interact the European influence variable (`EUP Influence`) with a binary indicator for whether the recipient is a European country (`European recipient`), and re-estimate the main models accordingly.
+The results are presented in @tbl-results-europe.
+
+Contrary to the hypothesis of regional favouritism, the interaction terms are either statistically insignificant or, when significant, indicate a *weaker* effect of European influence for European borrowers.
+For example, in the loan-to-GDP ratio and participation rate models, the negative and significant interaction terms suggest that European influence is associated with *less favourable* outcomes for European recipients compared to non-Europeans.
+
+These findings suggest that the observed European influence is not driven by a preference for geographically proximate countries, but instead reflects a broader institutional influence over the IMF.
+In other words, European powers appear to exercise their influence globally, rather than selectively favoring their regional neighbours.
+
 ```{r}

### 2025-07-16 code/is_Europe.R
use `\times` rather than x

@@ -98,7 +98,7 @@ msummary(regModels, stars = T) # Rough regression table with all variables and s
 coefmap <- c(
   'us' = 'USA Influence',
   'eu' = 'EUP Influence',
-  'eu:is_european' = 'EUP Influence x European recipient',
+  'eu:is_european' = 'EUP Influence $\\times$ European recipient',
   'is_european' = 'European recipient',
   'shstaffl' = 'IMF Staff',
   'shquotal' = 'IMF Quota',

### 2025-07-16 code/is_Europe.R
Fix variable line break

@@ -98,7 +98,7 @@ msummary(regModels, stars = T) # Rough regression table with all variables and s
 coefmap <- c(
   'us' = 'USA Influence',
   'eu' = 'EUP Influence',
-  'eu:is_european' = 'EUP Influence $\\times$ European recipient',
+  'eu:is_european' = 'EUP Influence \\\\~$\\times$ European recipient',
   'is_european' = 'European recipient',
   'shstaffl' = 'IMF Staff',
   'shquotal' = 'IMF Quota',

### 2025-07-16 manuscript/manuscript.qmd
Add question mark to robustness title
Rearrange paragraphs to add Greece example

@@
 
-## Robustness: Continental Preference
+## Robustness: Continental Preference?
 
-One plausible alternative explanation for the strong association between European influence and favourable IMF lending outcomes is that European powers might be disproportionately favouring fellow European recipients.
-To evaluate this possibility, we examine whether the effect of European influence differs significantly between European and non-European countries.
+One plausible alternative explanation for the strong association between European influence and favourable IMF lending outcomes is that European powers might be disproportionately favouring fellow European recipients, in contrast to the United States, for which such continental preference may not apply.
+This suspicion is not unfounded: during the Eurozone crisis, the IMF partnered with the European Commission and the European Central Bank—collectively known as the Troika—to jointly extend financial assistance to Greece and other European economies.
 
+To evaluate this possibility, we examine whether the effect of European influence varies systematically between European and non-European countries.
 We interact the European influence variable (`EUP Influence`) with a binary indicator for whether the recipient is a European country (`European recipient`), and re-estimate the main models accordingly.

### 2025-07-16 code/is_Europe.R
Fix mismatched divider in table

@@ -98,7 +98,7 @@ msummary(regModels, stars = T) # Rough regression table with all variables and s
 coefmap <- c(
   'us' = 'USA Influence',
   'eu' = 'EUP Influence',
-  'eu:is_european' = 'EUP Influence \\\\~$\\times$ European recipient',
+  'eu:is_european' = 'EUP Influence $\\times$ \\newline European recipient',
   'is_european' = 'European recipient',
   'shstaffl' = 'IMF Staff',
   'shquotal' = 'IMF Quota',

### 2025-07-16 code/is_Europe.R
Add FE and regressiopn type to table

@@ -93,6 +93,10 @@ save(regModels, file = 'save/regModels_europe.RData')
 # load('save/regModels.RData') # Load the models again if needed
 msummary(regModels, stars = T) # Rough regression table with all variables and stars
 
+## gof generation -----------------------------------------------------------------------------------------------------------------------------------------
+
+source("helper/f_test.R")
+
 ## Variable renaming -------------------------------------------------------------------------------------------------------------------------------------------
 
 coefmap <- c(

### 2025-07-16 manuscript/manuscript.qmd
move interpretation below the table
Fix table captions

@@
 #| label: tbl-results
-#| tbl-cap: Comparative Effects of US and EU Influence on IMF Lending Outcomes (1980–2010)
+#| tbl-cap: Comparative Effects of US and European Influence on IMF Lending Outcomes (1980–2010)
 #| tbl-pos: H
@@
 
+```{r}
+#| label: tbl-results-europe
+#| tbl-cap: Effects of US and European Influence on IMF Lending Outcomes (1980–2010), Separating European and Non-European Recipients
+#| tbl-pos: H
+load('save/regTable_europe.RData')
+resultsTable |> tinytable::theme_tt("resize", width = 1)
+```
+
 Contrary to the hypothesis of regional favouritism, the interaction terms are either statistically insignificant or, when significant, indicate a *weaker* effect of European influence for European borrowers.
@@
 In other words, European powers appear to exercise their influence globally, rather than selectively favoring their regional neighbours.
-
-```{r}
-#| label: tbl-results-europe
-#| tbl-cap: Effects of U.S. and European Influence on IMF Lending Outcomes (1980–2010), Separating European and Non-European Recipients
-#| tbl-pos: H
-load('save/regTable_europe.RData')
-resultsTable |> tinytable::theme_tt("resize", width = 1)
-```
 

### 2025-07-16 manuscript/manuscript.qmd
update interaction term interpretation

@@
 
-Contrary to the hypothesis of regional favouritism, the interaction terms are either statistically insignificant or, when significant, indicate a *weaker* effect of European influence for European borrowers.
-For example, in the loan-to-GDP ratio and participation rate models, the negative and significant interaction terms suggest that European influence is associated with *less favourable* outcomes for European recipients compared to non-Europeans.
+Contrary to the hypothesis of regional favouritism, the results suggest that European recipients do not receive preferential treatment from the IMF or benefit more from their ties to Western European powers.
+If anything, the estimates indicate that they may even be treated less favourably in some respects.
+The interaction terms (`EUP Influence $\times$ European recipient`), which capture whether European influence has a differential effect for European borrowers compared to non-European ones, are either statistically insignificant or, when significant, indicate a *weaker* effect for European recipients.
+For instance, in the loan-to-GDP ratio, participation rate, and IMF conditions models, the negative and significant interaction terms suggest that ties with Western European powers are associated with *less favourable* outcomes for European recipients relative to non-European countries with comparable ties.
 

### 2025-07-16 manuscript/manuscript.qmd
Add european recipient coefficient interpretation

@@
 
+Furthermore, the `European recipient` variable itself is associated with *less favourable* outcomes, even after controlling for ties to the US and European powers, as well as other covariates.
+This negative association is statistically significant at the 5% level for both the loan-to-GDP ratio and the number of IMF conditions.
+These results suggest that, on average, European countries tend to receive *less favourable* treatment from the IMF than non-European countries, despite their typically stronger connections to Western European powers.
+
 These findings suggest that the observed European influence is not driven by a preference for geographically proximate countries, but instead reflects a broader institutional influence over the IMF.

### 2025-07-16 manuscript/manuscript.qmd
Adjust spelling for last para
Robustness: Continental Preference? (#33)
* add europe-disaggregated table

* update  regtable_europe for new tidytable version

* Add robustness draft

* use `\times` rather than x

* Fix variable line break

* Add question mark to robustness title

* Rearrange paragraphs to add Greece example

* move interpretation below the table

* Fix mismatched divider in table

* Add FE and regressiopn type to table

* Fix table captions

* update interaction term interpretation

* Add european recipient coefficient interpretation

* Adjust spelling for last para

@@
 
--   IMF chief European [@kahler2001]
+As an international organization formally governed by a system of weighted voting and technocratic procedures, IMF is viewed as an “accountant” [@lutz2019], primarily concerned with the internal consistency of fiscal projections to guarantee timely repayment of its financial assistance.
+However, empirical studies suggested that countries with low international reserves, high current account deficits, and low GDP growth are more likely to receive IMF funding assistance [@barro2005; @knight1997], despite such indicators often reflecting a limited capacity to repay.
+This paradox leads to a widely argued statement, particularly among realists, that international institutions are “basically a reflection of the distribution of power in the world” [@mearsheimer1994].
+Generally, all studies on IMF lending outcomes agree that major power impose their influence in the process.
+For example, studies have shown that countries strategically important to the United States are more likely to receive IMF loans, or to receive larger amounts.
+Specifically, this applies to countries that align with the U.S.
+in United Nations General Assembly and United Nations Security Council voting [@lipscy2018; @thacker1999; @dreher2009], are in geopolitical vital regions \[\], or hold significant debts to American commercial banks [@oatley2004].
+This is not surprising, as the United States, as the most powerful state, has long projected its influence over international financial institutions—much like it has done with the World Bank [@wade2002; @woods2003; @clark2020].
+The question this study brings about is, how this reflection of power is distributed within a given international organization, especially one that is widely perceived to be dominated by a few major powers.
+
+The projection of great power capabilities into international organizations is sometimes uneven, because of strategic preferences, the perceived inefficiency of institutional engagement, or other possible reasons.
+Existing studies provide substantial evidence of U.S.
+influence over IMF lending decisions, they leave a crucial question open-the comparison between the U.S.
+and other major powers, particularly the European countries.
+IMF is perceived to be an institution dominated by a few major power, but the relative weight of American versus European remains ambiguous.
+The organization is traditionally chaired by an European national [@kahler2001],
 
@@
 These findings suggest that the observed European influence is not driven by a preference for geographically proximate countries, but instead reflects a broader institutional influence over the IMF.
-In other words, European powers appear to exercise their influence globally, rather than selectively favoring their regional neighbours.
+In other words, European powers appear to exercise their influence globally, rather than selectively favouring their regional neighbours.
 

### 2025-07-16 code/is_Europe.R

@@ -93,12 +93,16 @@ save(regModels, file = 'save/regModels_europe.RData')
 # load('save/regModels.RData') # Load the models again if needed
 msummary(regModels, stars = T) # Rough regression table with all variables and stars
 
+## gof generation -----------------------------------------------------------------------------------------------------------------------------------------
+
+source("helper/f_test.R")
+
 ## Variable renaming -------------------------------------------------------------------------------------------------------------------------------------------
 
 coefmap <- c(
   'us' = 'USA Influence',
   'eu' = 'EUP Influence',
-  'eu:is_european' = 'EUP Influence x European recipient',
+  'eu:is_european' = 'EUP Influence $\\times$ \\newline European recipient',
   'is_european' = 'European recipient',
   'shstaffl' = 'IMF Staff',
   'shquotal' = 'IMF Quota',

### 2025-07-17 manuscript/manuscript.qmd
Add Tong Hu as coauthor

@@
     correspondence: true
+  - name: Tong Hu
+    orcid: 0009-0003-5234-8164
 format:

### 2025-07-18 manuscript/manuscript.qmd
Fix $\times$

@@
 If anything, the estimates indicate that they may even be treated less favourably in some respects.
-The interaction terms (`EUP Influence $\times$ European recipient`), which capture whether European influence has a differential effect for European borrowers compared to non-European ones, are either statistically insignificant or, when significant, indicate a *weaker* effect for European recipients.
+The interaction terms (`EUP Influence` $\times$ `European recipient`), which capture whether European influence has a differential effect for European borrowers compared to non-European ones, are either statistically insignificant or, when significant, indicate a *weaker* effect for European recipients.
 For instance, in the loan-to-GDP ratio, participation rate, and IMF conditions models, the negative and significant interaction terms suggest that ties with Western European powers are associated with *less favourable* outcomes for European recipients relative to non-European countries with comparable ties.

### 2025-07-19 manuscript/manuscript.qmd
Amend para 1 and 2 in Intro

@@
 The International Monetary Fund (IMF) and the World Bank[^1] were established toward the end of World War II as part of the Bretton Woods agreement in 1944 [@jensen2004].
-While the World Bank specialised in long-term project financing, the IMF was initially designed for addressing the short-term balance-of-payment crises which would otherwise undermine the fixed exchange rate arrangements under the Bretton Woods system [@jensen2004].
+While the World Bank specialised in long-term project financing, the IMF was initially established for providing short-term balance-of-payment support and ensuring the stability of the international monetary system  [@fund2020; @jensen2004].
 
 [^1]: Familiar name for the International Bank for Reconstruction and Development (IBRD), the lending arm of the later created World Bank Group.
+
+Despite IMF's initial technical and narrow mandate, the IMF gradually expanded its role into structural adjustment and domestic economic governance, with loan conditionality often targeting fiscal policy, labour markets, and institutional reforms [@barnett2004].
+This evolution pushed the Fund into deeply political and distributional arenas far beyond its original design.
 
@@
 
+\newpage
+
 # References {.unnumbered}

### 2025-07-19 manuscript/manuscript.qmd
Add IMF consequences

@@
 This evolution pushed the Fund into deeply political and distributional arenas far beyond its original design.
+
+However, despite the far-reaching nature of IMF loan conditionality—often inspired by the tenets of neoclassical economics [@best2007; @clift2021; @chwieroth2015; @nelson2014], such as fiscal austerity, market liberalisation, and privatisation—there is little robust evidence that these prescriptions consistently lead to improved economic outcomes.
+In many cases, IMF programmes have been associated with worsening economic inequality [@garuda2000; @lang2020; @oberdabernig2013], stagnating growth [@przeworski2000; @oberdabernig2013], or social unrest [@hartzell2010; @casper2017], raising doubts about the effectiveness and legitimacy of its policy advice.
 

### 2025-07-19 manuscript/manuscript.qmd
Add 1 reference on stagnating growth

@@
 However, despite the far-reaching nature of IMF loan conditionality—often inspired by the tenets of neoclassical economics [@best2007; @clift2021; @chwieroth2015; @nelson2014], such as fiscal austerity, market liberalisation, and privatisation—there is little robust evidence that these prescriptions consistently lead to improved economic outcomes.
-In many cases, IMF programmes have been associated with worsening economic inequality [@garuda2000; @lang2020; @oberdabernig2013], stagnating growth [@przeworski2000; @oberdabernig2013], or social unrest [@hartzell2010; @casper2017], raising doubts about the effectiveness and legitimacy of its policy advice.
+In many cases, IMF programmes have been associated with worsening economic inequality [@garuda2000; @lang2020; @oberdabernig2013], stagnating growth [@dreher2006; @przeworski2000; @oberdabernig2013], or social unrest [@hartzell2010; @casper2017], raising doubts about the effectiveness and legitimacy of its policy advice.
 

### 2025-07-19 manuscript/manuscript.qmd
References on new page

--- 2025-07-19 manuscript/manuscript.qmd
@@
 The International Monetary Fund (IMF) and the World Bank[^1] were established toward the end of World War II as part of the Bretton Woods agreement in 1944 [@jensen2004].
-While the World Bank specialised in long-term project financing, the IMF was initially established for providing short-term balance-of-payment support and ensuring the stability of the international monetary system  [@fund2020; @jensen2004].
+While the World Bank specialised in long-term project financing, the IMF was initially designed for addressing the short-term balance-of-payment crises which would otherwise undermine the fixed exchange rate arrangements under the Bretton Woods system [@jensen2004].
 
 [^1]: Familiar name for the International Bank for Reconstruction and Development (IBRD), the lending arm of the later created World Bank Group.
-
-Despite IMF's initial technical and narrow mandate, the IMF gradually expanded its role into structural adjustment and domestic economic governance, with loan conditionality often targeting fiscal policy, labour markets, and institutional reforms [@barnett2004].
-This evolution pushed the Fund into deeply political and distributional arenas far beyond its original design.
 

### 2025-07-19 manuscript/manuscript.qmd
Add neoliberalism

--- 2025-07-19 manuscript/manuscript.qmd
@@
 In many cases, IMF programmes have been associated with worsening economic inequality [@garuda2000; @lang2020; @oberdabernig2013], stagnating growth [@dreher2006; @przeworski2000; @oberdabernig2013], or social unrest [@hartzell2010; @casper2017], raising doubts about the effectiveness and legitimacy of its policy advice.
+
+This performance problem becomes a *legitimacy* issue when compounded by the Fund’s institutional design, which structurally privileges the interests of its most powerful member states.
+With a weighted voting system [@fund2020] and governance norms that concentrate decision-making among a handful of advanced economies—particularly those in the West—the IMF has long been criticised as a platform for great power projection [@barnett2004].
+Scholars have drawn parallels between the Fund’s influence and forms of informal empire or *neo-colonialism*, where powerful donor states impose conditions on economically weaker recipients through seemingly technocratic institutions [@cain2022; @chen2021; @hickel2020; @moyo2024].
 

### 2025-07-19 manuscript/manuscript.qmd
Fix transition to transatlantic comparison

@@
 
-One renowned fact about the IMF and its sister institution, the World Bank, is the long-standing "gentlemen's agreement" of the IMF managing director being a European, and the World Bank president being an American [@keating2024; @weiss2023], despite no such provisions in either institution's articles of agreements [@fund2020; @WorldBank2021].
+While critiques of Western dominance in the IMF are longstanding, they often treat “the West” as a monolithic bloc.
+In practice, however, the institutional structure of the IMF reflects a careful transatlantic balancing act.
+One of the most symbolic manifestations of this arrangement is the long-standing “gentlemen’s agreement” under which the IMF’s Managing Director is always a European, while the World Bank is headed by an American [@keating2024; @weiss2023].
+This tacit division of leadership—despite being absent from the formal Articles of Agreement [@fund2020; @WorldBank2021]—raises an underexplored but important question: *who actually wields more influence within the IMF, Europe or the United States?* Addressing this question helps clarify the nature of global economic governance and sheds light on the dynamics within what is often treated as a unified Western core.
 
-Such arrangements inspire a straightforward question that is yet difficult to answer: who has more influence over the IMF, the US or Europe?
-Whereas the US dominance in the World Bank is undisputed [@clark2020; @wade2002; @woods2003], the case of the IMF is much more complicated.
-As the [Literature Review](#sec-litreview) section demonstrates, there has been abundant evidence of both European and American influence on the IMF, but short of a qualitative or quantitative comparison between the two.
+The question is straightforward but surprisingly difficult to answer.
+Whereas the US dominance in the World Bank is undisputed [@clark2020; @wade2002; @woods2003], the IMF presents a more ambiguous case.
+As the [Literature Review](#sec-litreview) section shows, existing research offers abundant evidence of both American and European influence, yet rarely attempts a direct comparison between the two.
 

### 2025-07-19 manuscript/manuscript.qmd
Fix qualititative research para for more natural language

@@
 
-Qualitative attempts at this question may have been thwarted by the closed-door and consensus-based style of governance of the IMF, as epitomised by the "gentlemen's agreement".
-On the one hand, the frequent use of conventions and consensus renders any analyses on the formal structure of the IMF (e.g.
-veto power) insufficient on its own.
-On the other hand, the closed-door negotiations imply that we often observe the outcome, but not the process of the bargaining.
-While specific case studies (e.g.
-Mexico and European debt crisis, see @lipscy2018 and @hodson2014) may have shed light on the influence of the US or Europe, they rarely allow for a transatlantic comparison and their representativeness can be questioned.
+Qualitative approaches to this issue are often hampered by the Fund’s closed-door, consensus-based decision-making—precisely the kind of opaque process exemplified by the “gentlemen’s agreement.” On the one hand, the prevalence of informal conventions limits the utility of formal institutional analyses (e.g., voting power or veto rights).
+On the other, the secretive nature of negotiations means scholars often observe the outcomes, but not the underlying power struggles.
+While case studies such as Mexico’s bailout or the European debt crisis [@lipscy2018; @hodson2014] provide compelling narratives, they lack transatlantic comparability and generalisability.
 

### 2025-07-19 manuscript/manuscript.qmd
Add contribution para

@@
 
-This paper therefore attempts to quantitatively compare the influence of the US and Europe on the IMF.
+This paper contributes to the literature in three key ways.
+First, it offers a direct empirical comparison of US and European influence over IMF lending outcomes—an angle surprisingly overlooked in existing work.
+Second, by disaggregating “the West” into its US and European components, it uncovers variation in how influence is exercised and to what ends.
+Third, it challenges the assumption of geographic bias, showing that European powers do not disproportionately favour European borrowers.
+These findings shed new light on the internal politics of the IMF and speak to wider debates on institutional legitimacy, great power contestation, and the geopolitics of global finance.
 

### 2025-07-21 manuscript/manuscript.qmd
Add opening para

@@
 # Literature Review {#sec-litreview}
+
+International Relations theory, particularly realism, has long emphasised that international organisations reflect underlying distributions of power rather than operating as neutral or autonomous actors.
+As Mearsheimer famously argued, international institutions are “basically a reflection of the distribution of power in the world” [@mearsheimer1994].
+The International Monetary Fund (IMF), despite its formal presentation as a rules-based and technocratic institution [@barnett2004; @chwieroth2015; @lutz2019], is no exception [@copelovitch2010a].
+Empirical studies have consistently shown that geopolitical considerations play a central role in IMF lending decisions.
 

### 2025-07-21 manuscript/manuscript.qmd
Add "Western" bloc

@@
 Empirical studies have consistently shown that geopolitical considerations play a central role in IMF lending decisions.
+
+In the case of the IMF, there is broad consensus that the institution is dominated by “Western” powers.
+This is reflected in both historical accounts of its founding and institutional design—especially its weighted voting system, which concentrates power in the hands of the United States and European countries [@kahler2001; @barnett2004; @copelovitch2010].
+Given the ideological, geopolitical, and economic proximity between the US and major European states, many studies treat “the West” as a unified bloc and do not disaggregate their respective influences [e.g.
+@lipscy2018; @vreeland2007; @copelovitch2010a].
 

### 2025-07-21 manuscript/manuscript.qmd
Add citation to intro

@@
 
-This performance problem becomes a *legitimacy* issue when compounded by the Fund’s institutional design, which structurally privileges the interests of its most powerful member states.
+This performance problem becomes a *legitimacy* issue when compounded by the Fund’s institutional design, which structurally privileges the interests of its most powerful member states [@stone2011].
 With a weighted voting system [@fund2020] and governance norms that concentrate decision-making among a handful of advanced economies—particularly those in the West—the IMF has long been criticised as a platform for great power projection [@barnett2004].

### 2025-07-21 manuscript/manuscript.qmd
add stone reference to 1st para

@@
 As Mearsheimer famously argued, international institutions are “basically a reflection of the distribution of power in the world” [@mearsheimer1994].
-The International Monetary Fund (IMF), despite its formal presentation as a rules-based and technocratic institution [@barnett2004; @chwieroth2015; @lutz2019], is no exception [@copelovitch2010a].
+The International Monetary Fund (IMF), despite its formal presentation as a rules-based and technocratic institution [@barnett2004; @chwieroth2015; @lutz2019], is no exception [@copelovitch2010a, @stone2011].
 Empirical studies have consistently shown that geopolitical considerations play a central role in IMF lending decisions.

### 2025-07-22 manuscript/manuscript.qmd
fix 2nd para references
Add copelovitch reference to intro

@@
 On the other, the secretive nature of negotiations means scholars often observe the outcomes, but not the underlying power struggles.
-While case studies such as Mexico’s bailout or the European debt crisis [@lipscy2018; @hodson2014] provide compelling narratives, they lack transatlantic comparability and generalisability.
+While case studies such as Mexico’s bailout or the European debt crisis [@lipscy2018; @copelovitch2010a; @hodson2014] provide compelling narratives, they lack transatlantic comparability and generalisability.
 
@@
 Given the ideological, geopolitical, and economic proximity between the US and major European states, many studies treat “the West” as a unified bloc and do not disaggregate their respective influences [e.g.
-@lipscy2018; @vreeland2007; @copelovitch2010a].
+@copelovitch2010a; @lipscy2018; @barro2005].
 

### 2025-07-22 manuscript/manuscript.qmd
remove mearsheimer para from Hu Tong's para
Add stone reference of consensus-based decision making to intro

@@
 
-Qualitative approaches to this issue are often hampered by the Fund’s closed-door, consensus-based decision-making—precisely the kind of opaque process exemplified by the “gentlemen’s agreement.” On the one hand, the prevalence of informal conventions limits the utility of formal institutional analyses (e.g., voting power or veto rights).
+Qualitative approaches to this issue are often hampered by the Fund’s closed-door, consensus-based decision-making [@stone2011, pp.55-56]—precisely the kind of opaque process exemplified by the “gentlemen’s agreement.” On the one hand, the prevalence of informal conventions limits the utility of formal institutional analyses (e.g., voting power or veto rights).
 On the other, the secretive nature of negotiations means scholars often observe the outcomes, but not the underlying power struggles.
@@
 However, empirical studies suggested that countries with low international reserves, high current account deficits, and low GDP growth are more likely to receive IMF funding assistance [@barro2005; @knight1997], despite such indicators often reflecting a limited capacity to repay.
-This paradox leads to a widely argued statement, particularly among realists, that international institutions are “basically a reflection of the distribution of power in the world” [@mearsheimer1994].
 Generally, all studies on IMF lending outcomes agree that major power impose their influence in the process.

### 2025-07-22 manuscript/manuscript.qmd
add vreeland to 1st para

@@
 As Mearsheimer famously argued, international institutions are “basically a reflection of the distribution of power in the world” [@mearsheimer1994].
-The International Monetary Fund (IMF), despite its formal presentation as a rules-based and technocratic institution [@barnett2004; @chwieroth2015; @lutz2019], is no exception [@copelovitch2010a, @stone2011].
+The International Monetary Fund (IMF), despite its formal presentation as a rules-based and technocratic institution [@barnett2004; @chwieroth2015; @lutz2019], is no exception [@copelovitch2010a, @stone2011, @vreeland2009].
 Empirical studies have consistently shown that geopolitical considerations play a central role in IMF lending decisions.

### 2025-07-22 manuscript/manuscript.qmd
Add US qualitative

@@
 @copelovitch2010a; @lipscy2018; @barro2005].
+
+For the strand of literature that examines American and European influence separately, nearly all studies identify the United States as the dominant actor within the IMF [@stone2011, p.51].
+Qualitative accounts emphasise the United States’ foundational role in shaping the IMF’s governance structure [@stone2011], the Fund’s practice of making direct enquiries with Washington for major lending decisions [@southard1979; @woods2003], and its unique status as the only country-level veto player, stemming from its position as the largest financial contributor [@woods2003; @stone2011].
+The US also benefits from significant informational advantages over other members [@stone2011, p.57] and proximity-based access, with the IMF headquartered in Washington, D.C.
+[@stone2011, p.57].
+Numerous case studies further demonstrate instances of US political pressure resulting in lenient loan conditions for key geopolitical allies, including Zaire, Egypt, Mexico, Argentina, and South Korea [@vreeland2009; @copelovitch2010a; @lipscy2018].
 

### 2025-07-23 manuscript/manuscript.qmd
add draft for 3rd and 4th para

@@
 
-As an international organization formally governed by a system of weighted voting and technocratic procedures, IMF is viewed as an “accountant” [@lutz2019], primarily concerned with the internal consistency of fiscal projections to guarantee timely repayment of its financial assistance.
-However, empirical studies suggested that countries with low international reserves, high current account deficits, and low GDP growth are more likely to receive IMF funding assistance [@barro2005; @knight1997], despite such indicators often reflecting a limited capacity to repay.
-Generally, all studies on IMF lending outcomes agree that major power impose their influence in the process.
-For example, studies have shown that countries strategically important to the United States are more likely to receive IMF loans, or to receive larger amounts.
-Specifically, this applies to countries that align with the U.S.
-in United Nations General Assembly and United Nations Security Council voting [@lipscy2018; @thacker1999; @dreher2009], are in geopolitical vital regions \[\], or hold significant debts to American commercial banks [@oatley2004].
-This is not surprising, as the United States, as the most powerful state, has long projected its influence over international financial institutions—much like it has done with the World Bank [@wade2002; @woods2003; @clark2020].
-The question this study brings about is, how this reflection of power is distributed within a given international organization, especially one that is widely perceived to be dominated by a few major powers.
+Quantitative studies also provide evidence of great power influence over IMF lending decisions.
+Such studies often use a variety of proxies to measure the recipient country's ties to the major powers, such as UN voting alignment [@thacker1999; @dreher2009], bilateral trade intensity [@barro2005], and financial exposure [@oatley2004].
+These proxies aim to capture the strategic incentives of the great powers to support IMF lending to countries that align with its geopolitical interests.
+Such ties have been proven to be associated with more favourable IMF lending outcomes, such as larger loan amounts, higher participation rates, as well as less and more lenient conditionality [@lipscy2018; @barro2005; @dreher2009].
+
+Yet, much of the quantitative literature has remained US-centric, often using only American alignment as the independent variable of interest [@thacker1999; @oatley2004; @dreher2009].
+Others pool European indicators with US ones, often under the umbrella of “Western” influence [@lipscy2018, @barro2005].
 

### 2025-07-25 manuscript/manuscript.qmd
Add European para

@@
 
-Quantitative studies also provide evidence of great power influence over IMF lending decisions.
-Such studies often use a variety of proxies to measure the recipient country's ties to the major powers, such as UN voting alignment [@thacker1999; @dreher2009], bilateral trade intensity [@barro2005], and financial exposure [@oatley2004].
-These proxies aim to capture the strategic incentives of the great powers to support IMF lending to countries that align with its geopolitical interests.
-Such ties have been proven to be associated with more favourable IMF lending outcomes, such as larger loan amounts, higher participation rates, as well as less and more lenient conditionality [@lipscy2018; @barro2005; @dreher2009].
+In comparison, the literature on European influence is notably smaller.
+The most visible manifestation of European power within the Fund is the tradition of European leadership: all Managing Directors of the IMF since its inception have been European nationals [@kahler2001].
+Scholars have also pointed to Europe-backed lending decisions, such as those during the Eurozone crisis (notably Greece) [@lipscy2018; @hodson2014], and loans to former colonies in Africa [@stone2004].
+However, while these studies offer important insights, they tend to remain case-specific or anecdotal.
+Crucially, they stop short of offering a systematic comparison between American and European influence, leaving open the question of how these two centres of power operate within the IMF.
 

### 2025-07-25 manuscript/manuscript.qmd
Rewrite quantitative section

@@
 
-Yet, much of the quantitative literature has remained US-centric, often using only American alignment as the independent variable of interest [@thacker1999; @oatley2004; @dreher2009].
-Others pool European indicators with US ones, often under the umbrella of “Western” influence [@lipscy2018, @barro2005].
+In principle, quantitative approaches are well suited for directly comparing the influence of different great powers within international organisations.
+In the IMF context, however, most such studies focus on American influence, with little systematic attention paid to European power.
+These studies typically proxy the strategic interests of major powers using indicators such as UN voting alignment [@thacker1999; @dreher2009], bilateral trade intensity [@barro2005], and financial exposure [@oatley2004].
+Yet, much of this literature has remained US-centric, often using only American alignment as the independent variable of interest [@thacker1999; @oatley2004; @dreher2009].
+Other studies pool European indicators with US ones under the broader category of “Western” influence, thus masking potential differences between the two [@lipscy2018; @barro2005].
+These ties have been shown to correlate with more favourable IMF lending outcomes—larger loans, higher approval rates, and fewer or less stringent conditions [@lipscy2018; @barro2005; @dreher2009].
+Yet, despite the availability of these tools, comparative assessments between the US and Europe remain largely absent from this literature.
 

### 2025-07-25 manuscript/manuscript.qmd
Add contribution

@@
 Yet, despite the availability of these tools, comparative assessments between the US and Europe remain largely absent from this literature.
+
+This paper addresses this gap by conducting a disaggregated, quantitative comparison of US and European influence on IMF lending outcomes.
+By distinguishing between the two major transatlantic actors, we provide new insights into the internal dynamics of power within the IMF and contribute to broader debates over institutional legitimacy, hegemony, and the geopolitics of economic governance.
 

### 2025-07-25 manuscript/manuscript.qmd
Add WB comparison
Add original section on IMF economic lendnig criteria in red

@@
 Numerous case studies further demonstrate instances of US political pressure resulting in lenient loan conditions for key geopolitical allies, including Zaire, Egypt, Mexico, Argentina, and South Korea [@vreeland2009; @copelovitch2010a; @lipscy2018].
+This pattern of influence is consistent with the United States’ broader role across international financial institutions, particularly the World Bank, where American dominance is widely acknowledged [@wade2002; @woods2003; @clark2020].
 
@@
 
-The projection of great power capabilities into international organizations is sometimes uneven, because of strategic preferences, the perceived inefficiency of institutional engagement, or other possible reasons.
-Existing studies provide substantial evidence of U.S.
-influence over IMF lending decisions, they leave a crucial question open-the comparison between the U.S.
-and other major powers, particularly the European countries.
-IMF is perceived to be an institution dominated by a few major power, but the relative weight of American versus European remains ambiguous.
-The organization is traditionally chaired by an European national [@kahler2001],
+\color{red}
+However, empirical studies suggested that countries with low international reserves, high current account deficits, and low GDP growth are more likely to receive IMF funding assistance [@barro2005; @knight1997], despite such indicators often reflecting a limited capacity to repay.
+\color{black}
 

### 2025-07-25 manuscript/manuscript.qmd
Fix citations in para 1

@@
 As Mearsheimer famously argued, international institutions are “basically a reflection of the distribution of power in the world” [@mearsheimer1994].
-The International Monetary Fund (IMF), despite its formal presentation as a rules-based and technocratic institution [@barnett2004; @chwieroth2015; @lutz2019], is no exception [@copelovitch2010a, @stone2011, @vreeland2009].
+The International Monetary Fund (IMF), despite its formal presentation as a rules-based and technocratic institution [@barnett2004; @chwieroth2015; @lutz2019], is no exception [@copelovitch2010a; @stone2011; @vreeland2009].
 Empirical studies have consistently shown that geopolitical considerations play a central role in IMF lending decisions.

### 2025-07-25 manuscript/manuscript.qmd
Move economic indicators to covariate section

@@
 
-\color{red}
-However, empirical studies suggested that countries with low international reserves, high current account deficits, and low GDP growth are more likely to receive IMF funding assistance [@barro2005; @knight1997], despite such indicators often reflecting a limited capacity to repay.
-\color{black}
-
 # Empirical Analysis
@@
 
-We also include economic performance indicators such as GDP, per capita GDP, and lagged GDP growth as determinants of lending and borrowing behaviour.
-International reserves as a share of imports are used to capture a country’s ability to finance its balance of payments without IMF support.
-Finally, a binary indicator for OECD membership is included to proxy for institutional and economic differences among countries.
+We include a range of economic performance indicators that have been shown to systematically affect IMF lending decisions [@barro2005; @knight1997].
+To capture these dynamics, we include GDP, per capita GDP, and lagged GDP growth as proxies for macroeconomic performance.
+International reserves as a share of imports are used to measure a country’s external liquidity and ability to manage balance-of-payments pressures without IMF support.
+Finally, a binary variable for OECD membership serves as a proxy for broader institutional and economic differences among countries.
 

### 2025-07-28 manuscript/manuscript.qmd
europe lit revie

--- 2025-07-25 manuscript/manuscript.qmd
@@
 
-Qualitative approaches to this issue are often hampered by the Fund’s closed-door, consensus-based decision-making [@stone2011, pp.55-56]—precisely the kind of opaque process exemplified by the “gentlemen’s agreement.” On the one hand, the prevalence of informal conventions limits the utility of formal institutional analyses (e.g., voting power or veto rights).
+Qualitative approaches to this issue are often hampered by the Fund’s closed-door, consensus-based decision-making—precisely the kind of opaque process exemplified by the “gentlemen’s agreement.” On the one hand, the prevalence of informal conventions limits the utility of formal institutional analyses (e.g., voting power or veto rights).
 On the other, the secretive nature of negotiations means scholars often observe the outcomes, but not the underlying power struggles.
@@
 As Mearsheimer famously argued, international institutions are “basically a reflection of the distribution of power in the world” [@mearsheimer1994].
-The International Monetary Fund (IMF), despite its formal presentation as a rules-based and technocratic institution [@barnett2004; @chwieroth2015; @lutz2019], is no exception [@copelovitch2010a; @stone2011; @vreeland2009].
+The International Monetary Fund (IMF), despite its formal presentation as a rules-based and technocratic institution [@barnett2004; @chwieroth2015; @lutz2019], is no exception [@copelovitch2010a, @stone2011].
 Empirical studies have consistently shown that geopolitical considerations play a central role in IMF lending decisions.
@@
 
-For the strand of literature that examines American and European influence separately, nearly all studies identify the United States as the dominant actor within the IMF [@stone2011, p.51].
-Qualitative accounts emphasise the United States’ foundational role in shaping the IMF’s governance structure [@stone2011], the Fund’s practice of making direct enquiries with Washington for major lending decisions [@southard1979; @woods2003], and its unique status as the only country-level veto player, stemming from its position as the largest financial contributor [@woods2003; @stone2011].
-The US also benefits from significant informational advantages over other members [@stone2011, p.57] and proximity-based access, with the IMF headquartered in Washington, D.C.
-[@stone2011, p.57].
-Numerous case studies further demonstrate instances of US political pressure resulting in lenient loan conditions for key geopolitical allies, including Zaire, Egypt, Mexico, Argentina, and South Korea [@vreeland2009; @copelovitch2010a; @lipscy2018].
-This pattern of influence is consistent with the United States’ broader role across international financial institutions, particularly the World Bank, where American dominance is widely acknowledged [@wade2002; @woods2003; @clark2020].
+As an international organization formally governed by a system of weighted voting and technocratic procedures, IMF is viewed as an “accountant” [@lutz2019], primarily concerned with the internal consistency of fiscal projections to guarantee timely repayment of its financial assistance.
+However, empirical studies suggested that countries with low international reserves, high current account deficits, and low GDP growth are more likely to receive IMF funding assistance [@barro2005; @knight1997], despite such indicators often reflecting a limited capacity to repay.
+This paradox leads to a widely argued statement, particularly among realists, that international institutions are “basically a reflection of the distribution of power in the world” [@mearsheimer1994].
+Generally, all studies on IMF lending outcomes agree that major power impose their influence in the process.
+For example, studies have shown that countries strategically important to the United States are more likely to receive IMF loans, or to receive larger amounts.
+Specifically, this applies to countries that align with the U.S.
+in United Nations General Assembly and United Nations Security Council voting [@lipscy2018; @thacker1999; @dreher2009], are in geopolitical vital regions \[\], or hold significant debts to American commercial banks [@oatley2004].
+This is not surprising, as the United States, as the most powerful state, has long projected its influence over international financial institutions—much like it has done with the World Bank [@wade2002; @woods2003; @clark2020].
+The question this study brings about is, how this reflection of power is distributed within a given international organization, especially one that is widely perceived to be dominated by a few major powers.
 
-In comparison, the literature on European influence is notably smaller.
-The most visible manifestation of European power within the Fund is the tradition of European leadership: all Managing Directors of the IMF since its inception have been European nationals [@kahler2001].
-Scholars have also pointed to Europe-backed lending decisions, such as those during the Eurozone crisis (notably Greece) [@lipscy2018; @hodson2014], and loans to former colonies in Africa [@stone2004].
-However, while these studies offer important insights, they tend to remain case-specific or anecdotal.
-Crucially, they stop short of offering a systematic comparison between American and European influence, leaving open the question of how these two centres of power operate within the IMF.
+The projection of great power capabilities into international organizations is sometimes uneven, because of strategic preferences, the perceived inefficiency of institutional engagement, or other possible reasons.
+Existing studies provide substantial evidence of U.S.
+influence over IMF lending decisions, they leave a crucial question open-the comparison between the U.S.
+and other major powers, particularly the European countries.
+IMF is perceived to be an institution dominated by a few major power, but the relative weight of American versus European remains ambiguous.
+The organization is traditionally chaired by an European national [@kahler2001],
 
-In principle, quantitative approaches are well suited for directly comparing the influence of different great powers within international organisations.
-In the IMF context, however, most such studies focus on American influence, with little systematic attention paid to European power.
-These studies typically proxy the strategic interests of major powers using indicators such as UN voting alignment [@thacker1999; @dreher2009], bilateral trade intensity [@barro2005], and financial exposure [@oatley2004].
-Yet, much of this literature has remained US-centric, often using only American alignment as the independent variable of interest [@thacker1999; @oatley2004; @dreher2009].
-Other studies pool European indicators with US ones under the broader category of “Western” influence, thus masking potential differences between the two [@lipscy2018; @barro2005].
-These ties have been shown to correlate with more favourable IMF lending outcomes—larger loans, higher approval rates, and fewer or less stringent conditions [@lipscy2018; @barro2005; @dreher2009].
-Yet, despite the availability of these tools, comparative assessments between the US and Europe remain largely absent from this literature.
+The literature on Europe's influence in the IMF is rather fragmented and lacks systematic research.
+Most scholars (such as [@smaghi2004; @thacker1999 ; @miescu2014]) argue that the representatives from Europe are rather scattered, while the United States remains the dominant force.
+However, there are also scholars (such as [@keating2024; @lutz2014; @copelovitch2010a; @stone2004]) who argue that Europe has significant advantages in leadership, voting power, agenda setting, overseas influence, and promoting its own interests, which enables it to play an indispensable role in the issue of IMF loans.
 
-This paper addresses this gap by conducting a disaggregated, quantitative comparison of US and European influence on IMF lending outcomes.
-By distinguishing between the two major transatlantic actors, we provide new insights into the internal dynamics of power within the IMF and contribute to broader debates over institutional legitimacy, hegemony, and the geopolitics of economic governance.
+First, European countries have structural advantages in IMF institutions.
+As mentioned in the introduction, the IMF's leadership has long been held by Europeans, and this gentleman's agreement between the United States and Europe has always worked in practice [@keating2024].
+EU countries hold more than 30% of the voting rights in the IMF, among which Britain, France and Germany hold seats alone [@leech2005a].
+Research shows that if European countries unify their seats, it is highly possible that they will have equal veto power compared to the United States [@brandner2009a].
 
-\color{red}
-However, empirical studies suggested that countries with low international reserves, high current account deficits, and low GDP growth are more likely to receive IMF funding assistance [@barro2005; @knight1997], despite such indicators often reflecting a limited capacity to repay.
-\color{black}
+Second, European countries dominate policy design and technical bureaucracy.
+[@lutz2014]'s research shows that the EU and the European Central Bank promoted the "Washington Consensus"-style austerity policy in the IMF's loans to Central and Eastern European countries (such as Latvia and Romania).
+This is because France and Germany have high exposure to Eastern European banking risks.
+Specifically, European countries have formed a consistent position within the IMF through the EURIMF (European Representatives to the IMF) and SCIMF (Sub-committee on the IMF) mechanisms, which makes it easier to put pressure on the IMF to promote policies that are beneficial to European interests (ibid.).
+Meanwhile, during the European debt crisis, France and Germany pushed for the participation of the IMF in rescue efforts and established a joint mechanism of the IMF, the European Commission and the European Central Bank to deal with the crisis [@veron2016].
+European countries were both lenders and borrowers.
+However, this move has also been criticized for the IMF overly obeying the will of major European powers, and it is believed that the failure of European governments to adopt the IMF's reform suggestions is the core cause of the outbreak of the European debt crisis (ibid.).
 

### 2025-07-29 manuscript/manuscript.qmd
add EU lit review (pro stance)

--- 2025-07-25 manuscript/manuscript.qmd
@@
 
-In comparison, the literature on European influence is notably smaller.
-The most visible manifestation of European power within the Fund is the tradition of European leadership: all Managing Directors of the IMF since its inception have been European nationals [@kahler2001].
-Scholars have also pointed to Europe-backed lending decisions, such as those during the Eurozone crisis (notably Greece) [@lipscy2018; @hodson2014], and loans to former colonies in Africa [@stone2004].
+Compared to the extensive literature on the United States, scholarly analyses of European influence within the IMF remain relatively limited and fragmented.
+A number of scholars (such as @smaghi2004; @thacker1999; @miescu2014) emphasize that European representation within the Fund is institutionally dispersed, resulting in a fragmented voice that limits collective influence.
+In contrast, the United States retains a centralized and coherent presence, thereby maintaining a dominant position in key decision-making processes.
+Nevertheless, other contributions (such as @keating2024; @lutz2014; @copelovitch2010a; @stone2004) argue that European states possess significant structural and institutional advantages—including leadership appointments, voting weight, agenda-setting capacity, transnational financial exposure, and diplomatic leverage—which enable them to exert substantive influence over IMF lending practices.
+
+First, European countries benefit from long-standing institutional privileges within the IMF’s governance architecture.
+As noted in the introduction, the Fund’s leadership has traditionally been allocated to Europe under an informal transatlantic arrangement that dates back to the Bretton Woods settlement (@kahler2001).
+Collectively, EU member states control over 30% of the IMF’s voting rights, with the United Kingdom, France, and Germany each occupying individual executive board seats (@leech2005a).
+Quantitative studies suggest that, under a scenario of unified representation, the EU could potentially attain voting power equivalent to a de facto veto similar to that of the United States (@brandner2009a).
+These structural features provide European countries with disproportionate influence relative to their global economic weight.
+
+Second, European countries have played a leading role in shaping policy design and operational implementation within the IMF.
+@lutz2014 demonstrates that the EU and the European Central Bank were instrumental in embedding austerity-oriented conditionalities—reminiscent of the Washington Consensus—in IMF lending to Central and Eastern European states such as Latvia and Romania.
+This policy orientation was motivated by the significant exposure of French and German banks to these economies.
+At the request of EU institutions, IMF conditionality in these cases expanded beyond standard macroeconomic stabilization to encompass fiscal rules, structural adjustment, and financial sector supervision (@hodson2014).
+Coordination platforms such as EURIMF (European Representatives to the IMF) and SCIMF (Sub-Committee on the IMF) have enabled EU member states to form cohesive negotiating positions and act collectively to advance shared preferences within the Fund (@lutz2014).
+
+In this context, @copelovitch2010 argues that IMF lending decisions are fundamentally political, with European influence manifesting not only through formal mechanisms such as quota shares, but also through informal channels including diplomatic lobbying, technical staffing appointments, and the selection of mission team leaders.
+During the Eurozone sovereign debt crisis, major European powers, particularly France and Germany, actively solicited IMF participation in stabilization packages.
+The resulting “troika” mechanism—comprising the IMF, the European Commission, and the European Central Bank—was emblematic of Europe's dual status as both lender and borrower in the crisis (@veron2016).
+However, this arrangement attracted criticism for compromising IMF independence.
+Some scholars contend that excessive deference to European preferences undermined the Fund’s credibility, and that the reluctance of euro area governments to implement IMF-prescribed reforms was a key driver of the crisis’s deepening (@veron2016).
+
+Third, European influence over IMF lending has often served to advance strategic financial and geopolitical interests.
+This was particularly evident during the Eurozone crisis, as mentioned earlier, when major European economies—most notably Germany and France—sought to mitigate systemic risks arising from their banking sector’s exposure to highly indebted member states.
+Fiscal surveillance mechanisms, such as the EU’s 3% deficit ceiling, were complemented by IMF-backed assistance programs aimed at restoring financial stability (@heimberger2025, p.
+9).
+The initiation of the IMF’s financial assistance to Greece, for example, was undertaken at Germany’s explicit request (@kickert2019).
+Beyond economic stabilization, European policymakers also leveraged IMF involvement to enhance the legitimacy of domestic austerity programs, using the Fund’s reputation as an external technocratic authority to depoliticize painful fiscal adjustments (@kickert2019).
+
+Moreover, European states have extended their influence to extra-European contexts, particularly in Africa.
+Countries such as the United Kingdom and France have supported IMF concessional lending to former colonies, often motivated by a combination of commercial and strategic interests, including access to natural resources and the preservation of post-colonial diplomatic ties (@stone2004).
+However, this form of influence has not been without controversy.
+Critics such as @moyo2024 argue that European-backed IMF programs have promoted neoliberal policy frameworks—including market liberalization and privatization—without adequate consideration of local development priorities.
+
 However, while these studies offer important insights, they tend to remain case-specific or anecdotal.

### 2025-07-29 manuscript/manuscript.qmd
Add EU lit review (pro and cons)

@@
 First, European countries benefit from long-standing institutional privileges within the IMF’s governance architecture.
-As noted in the introduction, the Fund’s leadership has traditionally been allocated to Europe under an informal transatlantic arrangement that dates back to the Bretton Woods settlement (@kahler2001).
-Collectively, EU member states control over 30% of the IMF’s voting rights, with the United Kingdom, France, and Germany each occupying individual executive board seats (@leech2005a).
-Quantitative studies suggest that, under a scenario of unified representation, the EU could potentially attain voting power equivalent to a de facto veto similar to that of the United States (@brandner2009a).
+As noted in the introduction, the Fund’s leadership has traditionally been allocated to Europe under an informal transatlantic arrangement that dates back to the Bretton Woods settlement [@kahler2001].
+Collectively, EU member states control over 30% of the IMF’s voting rights, with the United Kingdom, France, and Germany each occupying individual executive board seats [@leech2005a].
+Quantitative studies suggest that, under a scenario of unified representation, the EU could potentially attain voting power equivalent to a de facto veto similar to that of the United States [@brandner2009a].
 These structural features provide European countries with disproportionate influence relative to their global economic weight.
@@
 This policy orientation was motivated by the significant exposure of French and German banks to these economies.
-At the request of EU institutions, IMF conditionality in these cases expanded beyond standard macroeconomic stabilization to encompass fiscal rules, structural adjustment, and financial sector supervision (@hodson2014).
-Coordination platforms such as EURIMF (European Representatives to the IMF) and SCIMF (Sub-Committee on the IMF) have enabled EU member states to form cohesive negotiating positions and act collectively to advance shared preferences within the Fund (@lutz2014).
+At the request of EU institutions, IMF conditionality in these cases expanded beyond standard macroeconomic stabilization to encompass fiscal rules, structural adjustment, and financial sector supervision [@hodson2014].
+Coordination platforms such as EURIMF (European Representatives to the IMF) and SCIMF (Sub-Committee on the IMF) have enabled EU member states to form cohesive negotiating positions and act collectively to advance shared preferences within the Fund [@lutz2014].
 
@@
 During the Eurozone sovereign debt crisis, major European powers, particularly France and Germany, actively solicited IMF participation in stabilization packages.
-The resulting “troika” mechanism—comprising the IMF, the European Commission, and the European Central Bank—was emblematic of Europe's dual status as both lender and borrower in the crisis (@veron2016).
+The resulting “troika” mechanism—comprising the IMF, the European Commission, and the European Central Bank—was emblematic of Europe's dual status as both lender and borrower in the crisis [@veron2016].
 However, this arrangement attracted criticism for compromising IMF independence.
-Some scholars contend that excessive deference to European preferences undermined the Fund’s credibility, and that the reluctance of euro area governments to implement IMF-prescribed reforms was a key driver of the crisis’s deepening (@veron2016).
+Some scholars contend that excessive deference to European preferences undermined the Fund’s credibility, and that the reluctance of euro area governments to implement IMF-prescribed reforms was a key driver of the crisis’s deepening [@veron2016].
 
@@
 This was particularly evident during the Eurozone crisis, as mentioned earlier, when major European economies—most notably Germany and France—sought to mitigate systemic risks arising from their banking sector’s exposure to highly indebted member states.
-Fiscal surveillance mechanisms, such as the EU’s 3% deficit ceiling, were complemented by IMF-backed assistance programs aimed at restoring financial stability (@heimberger2025, p.
-9).
-The initiation of the IMF’s financial assistance to Greece, for example, was undertaken at Germany’s explicit request (@kickert2019).
-Beyond economic stabilization, European policymakers also leveraged IMF involvement to enhance the legitimacy of domestic austerity programs, using the Fund’s reputation as an external technocratic authority to depoliticize painful fiscal adjustments (@kickert2019).
+Fiscal surveillance mechanisms, such as the EU’s 3% deficit ceiling, were complemented by IMF-backed assistance programs aimed at restoring financial stability [@heimberger2025, p.
+9].
+The initiation of the IMF’s financial assistance to Greece, for example, was undertaken at Germany’s explicit request [@kickert2019].
+Beyond economic stabilization, European policymakers also leveraged IMF involvement to enhance the legitimacy of domestic austerity programs, using the Fund’s reputation as an external technocratic authority to depoliticize painful fiscal adjustments [@kickert2019].
 
 Moreover, European states have extended their influence to extra-European contexts, particularly in Africa.
-Countries such as the United Kingdom and France have supported IMF concessional lending to former colonies, often motivated by a combination of commercial and strategic interests, including access to natural resources and the preservation of post-colonial diplomatic ties (@stone2004).
+Countries such as the United Kingdom and France have supported IMF concessional lending to former colonies, often motivated by a combination of commercial and strategic interests, including access to natural resources and the preservation of post-colonial diplomatic ties [@stone2004].
 However, this form of influence has not been without controversy.
-Critics such as @moyo2024 argue that European-backed IMF programs have promoted neoliberal policy frameworks—including market liberalization and privatization—without adequate consideration of local development priorities.
+Critics such as @moyo2024 and @cain2022 argue that European-backed IMF programs have promoted neoliberal policy frameworks—including market liberalization and privatization—without adequate consideration of local development priorities, undermining African countries' economic sovereighty and state capacity.
+
+As noted earlier, numerous scholars have questioned both the extent and effectiveness of Europe’s influence within the IMF.
+First, Europe faces persistent challenges in internal coordination.
+EU member states are represented through multiple, uncoordinated constituencies within the Fund, resulting in a fragmented and often inconsistent policy voice [@smaghi2004].
+This institutional fragmentation is compounded by substantive policy divergences among key member states—for instance, France tends to prioritize pro-growth strategies, while Germany places greater emphasis on fiscal discipline [@miescu2014].
+
+Second, despite Europe’s substantial voting power, the United States continues to enjoy structural advantages, most notably its effective veto power over major IMF decisions and its ability to shape key institutional outcomes [@thacker1999].
+Moreover, European influence is often constrained by the need to coordinate positions with other members of the G5, which may necessitate concessions that dilute European preferences or undermine coherence [@copelovitch2010a].
+
+Third, concerns have been raised regarding the democratic legitimacy and global representativeness of European countries within the IMF governance structure.
+@woodward2024 highlights the disproportionate relationship between the population size of EU member states and their aggregate voting power in the Fund, arguing that this imbalance structurally disadvantages developing countries.
+@leech2005a further suggests that while a consolidation of European seats could potentially establish a bipolar governance configuration capable of counterbalancing U.S.
+dominance, such a shift might also open space for increased participation and influence by small and medium-sized states within the IMF’s decision-making processes.
 

### 2025-07-29 manuscript/manuscript.qmd
EU lit review revised_0729

@@
     orcid: 0009-0003-5234-8164
+  - name: Zhenyi Chen
+    orcid: 0009-0008-2910-5187
 format:
@@
 9].
-The initiation of the IMF’s financial assistance to Greece, for example, was undertaken at Germany’s explicit request [@kickert2019].
-Beyond economic stabilization, European policymakers also leveraged IMF involvement to enhance the legitimacy of domestic austerity programs, using the Fund’s reputation as an external technocratic authority to depoliticize painful fiscal adjustments [@kickert2019].
+The initiation of the IMF’s financial assistance to Greece, for example, was undertaken at Germany’s explicit request [@kickert2019a].
+Beyond economic stabilization, European policymakers also leveraged IMF involvement to enhance the legitimacy of domestic austerity programs, using the Fund’s reputation as an external technocratic authority to depoliticize painful fiscal adjustments [@kickert2019a].
 

### 2025-07-30 manuscript/manuscript.qmd
europe lit review revised 2.0

@@
 However, while these studies offer important insights, they tend to remain case-specific or anecdotal.
-Crucially, they stop short of offering a systematic comparison between American and European influence, leaving open the question of how these two centres of power operate within the IMF.
+While existing literature offers valuable case-based insights into either U.S.
+or European influence in IMF operations, there is a notable absence of systematic empirical comparisons between the two.
+This study addresses that gap by directly evaluating and contrasting how U.S.
+and European political ties affect IMF lending outcomes, including program access, conditionality, and loan volumes.
+Secondly, much of the literature treats the West as a monolithic bloc of creditor powers.
+However, this study disaggregates “the West” into the United States and Europe, revealing divergent motivations, instruments, and patterns of influence.
+In doing so, it challenges the assumption of homogeneous Western preferences and highlights variation in how and why influence is exerted.
 

### 2025-07-30 manuscript/manuscript.qmd
shortened EU lit review part

--- 2025-07-29 manuscript/manuscript.qmd
@@
 
-Compared to the extensive literature on the United States, scholarly analyses of European influence within the IMF remain relatively limited and fragmented.
-A number of scholars (such as @smaghi2004; @thacker1999; @miescu2014) emphasize that European representation within the Fund is institutionally dispersed, resulting in a fragmented voice that limits collective influence.
-In contrast, the United States retains a centralized and coherent presence, thereby maintaining a dominant position in key decision-making processes.
-Nevertheless, other contributions (such as @keating2024; @lutz2014; @copelovitch2010a; @stone2004) argue that European states possess significant structural and institutional advantages—including leadership appointments, voting weight, agenda-setting capacity, transnational financial exposure, and diplomatic leverage—which enable them to exert substantive influence over IMF lending practices.
+Compared to the extensive literature on U.S.
+dominance in the IMF, scholarship on European influence is relatively limited and often fragmented.
+Many studies underscore that EU member states are institutionally dispersed across multiple constituencies, hindering coherent collective action [@smaghi2004; @miescu2014; @thacker1999].
+Still, other works argue that European countries wield significant structural advantages—including leadership appointments, voting power, and coordination mechanisms—that enable them to shape lending outcomes in line with their interests [@copelovitch2010a; @keating2024; @lutz2014; @stone2004].
 
-First, European countries benefit from long-standing institutional privileges within the IMF’s governance architecture.
-As noted in the introduction, the Fund’s leadership has traditionally been allocated to Europe under an informal transatlantic arrangement that dates back to the Bretton Woods settlement [@kahler2001].
-Collectively, EU member states control over 30% of the IMF’s voting rights, with the United Kingdom, France, and Germany each occupying individual executive board seats [@leech2005a].
-Quantitative studies suggest that, under a scenario of unified representation, the EU could potentially attain voting power equivalent to a de facto veto similar to that of the United States [@brandner2009a].
-These structural features provide European countries with disproportionate influence relative to their global economic weight.
+European influence reflects a strategic combination of formal power and coordinated action.
+EU countries collectively control over 30% of IMF voting rights, and if they were to unify, they could potentially counterbalance U.S.
+veto power [@brandner2009a; @leech2005a; @smaghi2004].
+This is reinforced by the long-standing transatlantic norm allocating IMF leadership to a European national [@kahler2001].
+Through platforms such as EURIMF and SCIMF, EU members have harmonized policy positions and influenced loan design.
+During the Eurozone crisis, France and Germany pushed for IMF involvement to stabilize the region while embedding austerity-oriented conditionality that protected their banking systems [@hodson2014; @kickert2019a; @veron2016].
+Lending terms were shaped to reflect creditor priorities, with evidence that conditionality was both intensified or relaxed based on geopolitical alignment [@lutz2014; @heimberger2025, p.
+9; @dreher2015].
+Beyond Europe, the UK and France have used IMF concessional lending in Africa to maintain post-colonial ties and commercial access [@stone2004].
 
-Second, European countries have played a leading role in shaping policy design and operational implementation within the IMF.
-@lutz2014 demonstrates that the EU and the European Central Bank were instrumental in embedding austerity-oriented conditionalities—reminiscent of the Washington Consensus—in IMF lending to Central and Eastern European states such as Latvia and Romania.
-This policy orientation was motivated by the significant exposure of French and German banks to these economies.
-At the request of EU institutions, IMF conditionality in these cases expanded beyond standard macroeconomic stabilization to encompass fiscal rules, structural adjustment, and financial sector supervision [@hodson2014].
-Coordination platforms such as EURIMF (European Representatives to the IMF) and SCIMF (Sub-Committee on the IMF) have enabled EU member states to form cohesive negotiating positions and act collectively to advance shared preferences within the Fund [@lutz2014].
+Despite Europe’s institutional advantages, scholars remain skeptical of its actual effectiveness in the IMF.
+Internal divergences—such as France’s growth-oriented stance versus Germany’s fiscal conservatism—combined with fragmented constituencies, often impede coherent action and reduce policy influence [@miescu2014; @dreher2015; @copelovitch2010a].
+Moreover, Europe’s disproportionate voting power relative to its population raises legitimacy concerns and is seen as disadvantaging developing countries [@woodward2024; @leech2005a].
+While proposals to consolidate EU seats aim to improve efficiency and counterbalance U.S.
+dominance, critics warn such reforms may merely reinforce dominance by a few large European states rather than broaden representation [@leech2005a; @copelovitch2010a].
 
-In this context, @copelovitch2010 argues that IMF lending decisions are fundamentally political, with European influence manifesting not only through formal mechanisms such as quota shares, but also through informal channels including diplomatic lobbying, technical staffing appointments, and the selection of mission team leaders.
-During the Eurozone sovereign debt crisis, major European powers, particularly France and Germany, actively solicited IMF participation in stabilization packages.
-The resulting “troika” mechanism—comprising the IMF, the European Commission, and the European Central Bank—was emblematic of Europe's dual status as both lender and borrower in the crisis [@veron2016].
-However, this arrangement attracted criticism for compromising IMF independence.
-Some scholars contend that excessive deference to European preferences undermined the Fund’s credibility, and that the reluctance of euro area governments to implement IMF-prescribed reforms was a key driver of the crisis’s deepening [@veron2016].
-
-Third, European influence over IMF lending has often served to advance strategic financial and geopolitical interests.
-This was particularly evident during the Eurozone crisis, as mentioned earlier, when major European economies—most notably Germany and France—sought to mitigate systemic risks arising from their banking sector’s exposure to highly indebted member states.
-Fiscal surveillance mechanisms, such as the EU’s 3% deficit ceiling, were complemented by IMF-backed assistance programs aimed at restoring financial stability [@heimberger2025, p.
-9].
-The initiation of the IMF’s financial assistance to Greece, for example, was undertaken at Germany’s explicit request [@kickert2019a].
-Beyond economic stabilization, European policymakers also leveraged IMF involvement to enhance the legitimacy of domestic austerity programs, using the Fund’s reputation as an external technocratic authority to depoliticize painful fiscal adjustments [@kickert2019a].
-
-Moreover, European states have extended their influence to extra-European contexts, particularly in Africa.
-Countries such as the United Kingdom and France have supported IMF concessional lending to former colonies, often motivated by a combination of commercial and strategic interests, including access to natural resources and the preservation of post-colonial diplomatic ties [@stone2004].
-However, this form of influence has not been without controversy.
-Critics such as @moyo2024 and @cain2022 argue that European-backed IMF programs have promoted neoliberal policy frameworks—including market liberalization and privatization—without adequate consideration of local development priorities, undermining African countries' economic sovereighty and state capacity.
-
-As noted earlier, numerous scholars have questioned both the extent and effectiveness of Europe’s influence within the IMF.
-First, Europe faces persistent challenges in internal coordination.
-EU member states are represented through multiple, uncoordinated constituencies within the Fund, resulting in a fragmented and often inconsistent policy voice [@smaghi2004].
-This institutional fragmentation is compounded by substantive policy divergences among key member states—for instance, France tends to prioritize pro-growth strategies, while Germany places greater emphasis on fiscal discipline [@miescu2014].
-
-Second, despite Europe’s substantial voting power, the United States continues to enjoy structural advantages, most notably its effective veto power over major IMF decisions and its ability to shape key institutional outcomes [@thacker1999].
-Moreover, European influence is often constrained by the need to coordinate positions with other members of the G5, which may necessitate concessions that dilute European preferences or undermine coherence [@copelovitch2010a].
-
-Third, concerns have been raised regarding the democratic legitimacy and global representativeness of European countries within the IMF governance structure.
-@woodward2024 highlights the disproportionate relationship between the population size of EU member states and their aggregate voting power in the Fund, arguing that this imbalance structurally disadvantages developing countries.
-@leech2005a further suggests that while a consolidation of European seats could potentially establish a bipolar governance configuration capable of counterbalancing U.S.
-dominance, such a shift might also open space for increased participation and influence by small and medium-sized states within the IMF’s decision-making processes.
-
-However, while these studies offer important insights, they tend to remain case-specific or anecdotal.
-Crucially, they stop short of offering a systematic comparison between American and European influence, leaving open the question of how these two centres of power operate within the IMF.
+While several studies examine European involvement in IMF operations, most focus on individual cases or regional crises, lacking generalizable conclusions.
+Existing literature often treats Europe as a secondary actor or subsumes it under undifferentiated “Western” influence, thereby overlooking intra-Western variation.
+By analytically separating U.S.
+and European roles, our study addresses these gaps and offers a comparative framework for understanding divergent patterns of influence within global financial governance.
 

### 2025-07-30 manuscript/manuscript.qmd
switch to Brirtish spelling

@@
 
-Compared to the extensive literature on U.S.
-dominance in the IMF, scholarship on European influence is relatively limited and often fragmented.
+Compared to the extensive literature on US dominance in the IMF, scholarship on European influence is relatively limited and often fragmented.
 Many studies underscore that EU member states are institutionally dispersed across multiple constituencies, hindering coherent collective action [@smaghi2004; @miescu2014; @thacker1999].
@@
 European influence reflects a strategic combination of formal power and coordinated action.
-EU countries collectively control over 30% of IMF voting rights, and if they were to unify, they could potentially counterbalance U.S.
-veto power [@brandner2009a; @leech2005a; @smaghi2004].
+EU countries collectively control over 30% of IMF voting rights, and if they were to unify, they could potentially counterbalance US veto power [@brandner2009a; @leech2005a; @smaghi2004].
 This is reinforced by the long-standing transatlantic norm allocating IMF leadership to a European national [@kahler2001].
-Through platforms such as EURIMF and SCIMF, EU members have harmonized policy positions and influenced loan design.
-During the Eurozone crisis, France and Germany pushed for IMF involvement to stabilize the region while embedding austerity-oriented conditionality that protected their banking systems [@hodson2014; @kickert2019a; @veron2016].
+Through platforms such as EURIMF and SCIMF, EU members have harmonised policy positions and influenced loan design.
+During the Eurozone crisis, France and Germany pushed for IMF involvement to stabilise the region while embedding austerity-oriented conditionality that protected their banking systems [@hodson2014; @kickert2019a; @veron2016].
 Lending terms were shaped to reflect creditor priorities, with evidence that conditionality was both intensified or relaxed based on geopolitical alignment [@lutz2014; @heimberger2025, p.
@@
 Moreover, Europe’s disproportionate voting power relative to its population raises legitimacy concerns and is seen as disadvantaging developing countries [@woodward2024; @leech2005a].
-While proposals to consolidate EU seats aim to improve efficiency and counterbalance U.S.
-dominance, critics warn such reforms may merely reinforce dominance by a few large European states rather than broaden representation [@leech2005a; @copelovitch2010a].
+While proposals to consolidate EU seats aim to improve efficiency and counterbalance US dominance, critics warn such reforms may merely reinforce dominance by a few large European states rather than broaden representation [@leech2005a; @copelovitch2010a].
 
-While several studies examine European involvement in IMF operations, most focus on individual cases or regional crises, lacking generalizable conclusions.
+While several studies examine European involvement in IMF operations, most focus on individual cases or regional crises, lacking generalisable conclusions.
 Existing literature often treats Europe as a secondary actor or subsumes it under undifferentiated “Western” influence, thereby overlooking intra-Western variation.
-By analytically separating U.S.
-and European roles, our study addresses these gaps and offers a comparative framework for understanding divergent patterns of influence within global financial governance.
+By analytically separating US and European roles, our study addresses these gaps and offers a comparative framework for understanding divergent patterns of influence within global financial governance.
 

### 2025-07-31 manuscript/manuscript.qmd
correct references

@@
 During the Eurozone crisis, France and Germany pushed for IMF involvement to stabilise the region while embedding austerity-oriented conditionality that protected their banking systems [@hodson2014; @kickert2019a; @veron2016].
-Lending terms were shaped to reflect creditor priorities, with evidence that conditionality was both intensified or relaxed based on geopolitical alignment [@lutz2014; @heimberger2025, p.
-9; @dreher2015].
-Beyond Europe, the UK and France have used IMF concessional lending in Africa to maintain post-colonial ties and commercial access [@stone2004].
+Lending terms were shaped to reflect creditor priorities, with evidence that conditionality was both intensified or relaxed based on geopolitical alignment [@cain2022; @dreher2015; @lutz2014].
+Beyond Europe, the UK and France have used IMF concessional lending in Africa to maintain post-colonial ties and commercial access [@moyo2024; @stone2004].
 

### 2025-07-31 manuscript/manuscript.qmd
Add rearranged version

@@
 
+Compared to the extensive literature on US dominance in the IMF, research on European influence remains more limited and fragmented.
+Nevertheless, a growing body of scholarship highlights that European countries do possess considerable institutional and structural advantages that enable them to shape lending outcomes.
+First, EU countries collectively hold over 30% of IMF voting shares—greater than the US—and, if coordinated, could theoretically counterbalance American veto power [@brandner2009a; @leech2005a; @smaghi2004].
+Second, Europe has enjoyed long-standing leadership within the Fund: all IMF Managing Directors since the institution's founding have been European nationals [@kahler2001].
+Third, coordination mechanisms such as EURIMF and SCIMF facilitate policy alignment among EU members and amplify their collective voice.
+During the Eurozone crisis, for instance, France and Germany lobbied for IMF involvement, embedding austerity-oriented conditionality that aligned with creditor interests and protected domestic banking sectors [@hodson2014; @kickert2019a; @veron2016].
+Scholars have also documented European influence beyond the continent—for example, in the use of IMF concessional lending to sustain post-colonial relationships in Africa, particularly by the UK and France [@moyo2024; @stone2004].
+These studies suggest that European actors, when effectively coordinated, are capable of steering lending priorities.
+
+At the same time, existing literature raises questions about the consistency and effectiveness of European influence.
+Internal differences—such as the often-cited contrast between France’s growth-oriented approach and Germany’s fiscal conservatism—along with the dispersion of EU members across multiple IMF constituencies, are said to limit coherence and reduce policy impact [@smaghi2004; @miescu2014; @dreher2015].
+While some proposals have called for the consolidation of EU seats to strengthen European influence, critics argue this may entrench the dominance of larger EU states without necessarily enhancing representativeness [@leech2005a; @copelovitch2010a].
+Moreover, concerns about Europe’s voting power exceeding its global economic and demographic weight have generated debates about fairness and legitimacy, particularly from developing countries [@woodward2024].
+These institutional tensions may place constraints on Europe’s ability to act decisively, though the extent of such limitations remains contested.
+
+Although many studies document specific cases of European involvement—especially during regional crises—they tend to be case-based, focus on intra-European dynamics, or treat Europe as a secondary actor.
+In addition, several key works either overlook European influence altogether or conflate it with broader “Western” alignment [@lipscy2018; @barro2005].
+As a result, generalisable, cross-national assessments of Europe’s role in shaping IMF outcomes remain scarce.
+This paper addresses that gap by analytically disaggregating US and European influence within a unified empirical framework.
+In doing so, it contributes to a more nuanced understanding of transatlantic dynamics in global financial governance—highlighting that Western influence may not be as unified or symmetrical as often assumed.
+
 Compared to the extensive literature on US dominance in the IMF, scholarship on European influence is relatively limited and often fragmented.
@@
 European influence reflects a strategic combination of formal power and coordinated action.
-EU countries collectively control over 30% of IMF voting rights, and if they were to unify, they could potentially counterbalance US veto power [@brandner2009a; @leech2005a; @smaghi2004].
-This is reinforced by the long-standing transatlantic norm allocating IMF leadership to a European national [@kahler2001].
-Through platforms such as EURIMF and SCIMF, EU members have harmonised policy positions and influenced loan design.
-During the Eurozone crisis, France and Germany pushed for IMF involvement to stabilise the region while embedding austerity-oriented conditionality that protected their banking systems [@hodson2014; @kickert2019a; @veron2016].
 Lending terms were shaped to reflect creditor priorities, with evidence that conditionality was both intensified or relaxed based on geopolitical alignment [@cain2022; @dreher2015; @lutz2014].
-Beyond Europe, the UK and France have used IMF concessional lending in Africa to maintain post-colonial ties and commercial access [@moyo2024; @stone2004].
 
@@
 While proposals to consolidate EU seats aim to improve efficiency and counterbalance US dominance, critics warn such reforms may merely reinforce dominance by a few large European states rather than broaden representation [@leech2005a; @copelovitch2010a].
-
-While several studies examine European involvement in IMF operations, most focus on individual cases or regional crises, lacking generalisable conclusions.
-Existing literature often treats Europe as a secondary actor or subsumes it under undifferentiated “Western” influence, thereby overlooking intra-Western variation.
-By analytically separating US and European roles, our study addresses these gaps and offers a comparative framework for understanding divergent patterns of influence within global financial governance.
 

### 2025-07-31 manuscript/manuscript.qmd
move arguments to new paras
Move references

@@
 First, EU countries collectively hold over 30% of IMF voting shares—greater than the US—and, if coordinated, could theoretically counterbalance American veto power [@brandner2009a; @leech2005a; @smaghi2004].
-Second, Europe has enjoyed long-standing leadership within the Fund: all IMF Managing Directors since the institution's founding have been European nationals [@kahler2001].
-Third, coordination mechanisms such as EURIMF and SCIMF facilitate policy alignment among EU members and amplify their collective voice.
+Second, Europe has enjoyed long-standing leadership within the Fund: all IMF Managing Directors since the institution's founding have been European nationals [@kahler2001; @keating2024].
+Third, coordination mechanisms such as EURIMF and SCIMF facilitate policy alignment among EU members and amplify their collective voice [@lutz2014].
 During the Eurozone crisis, for instance, France and Germany lobbied for IMF involvement, embedding austerity-oriented conditionality that aligned with creditor interests and protected domestic banking sectors [@hodson2014; @kickert2019a; @veron2016].
@@
 At the same time, existing literature raises questions about the consistency and effectiveness of European influence.
+Many studies underscore that EU member states are institutionally dispersed across multiple constituencies, hindering coherent collective action [@smaghi2004; @miescu2014; @thacker1999].
 Internal differences—such as the often-cited contrast between France’s growth-oriented approach and Germany’s fiscal conservatism—along with the dispersion of EU members across multiple IMF constituencies, are said to limit coherence and reduce policy impact [@smaghi2004; @miescu2014; @dreher2015].
@@
 
-Compared to the extensive literature on US dominance in the IMF, scholarship on European influence is relatively limited and often fragmented.
-Many studies underscore that EU member states are institutionally dispersed across multiple constituencies, hindering coherent collective action [@smaghi2004; @miescu2014; @thacker1999].
-Still, other works argue that European countries wield significant structural advantages—including leadership appointments, voting power, and coordination mechanisms—that enable them to shape lending outcomes in line with their interests [@copelovitch2010a; @keating2024; @lutz2014; @stone2004].
+voting power, and coordination mechanisms—that enable them to shape lending outcomes in line with their interests [@copelovitch2010a].
 

### 2025-07-31 manuscript/manuscript.qmd
remove a para
Add abstract draft

--- 2025-07-31 manuscript/manuscript.qmd
@@
     orcid: 0009-0003-5234-8164
-  - name: Zhenyi Chen
-    orcid: 0009-0008-2910-5187
 format:
@@
           \AtAppendix{\counterwithin{figure}{section}}
-abstract: This is an abstract.
+abstract: "This paper investigates whether the United States or Europe holds greater influence over the International Monetary Fund (IMF)—a longstanding yet underexamined question in the study of global economic governance.
+While the Fund is widely regarded as dominated by “Western” powers, most existing research either focuses exclusively on American influence or treats the West as a unified bloc, masking important variation between its two key actors.
+Drawing on panel data from 1980 to 2010, we construct separate measures of US and European influence using principal component analysis of commonly used alignment indicators—UN voting patterns, bilateral trade, and banking exposure—and estimate a series of Tobit and Probit models to assess their impact on four key IMF lending outcomes: loan-to-GDP ratios, participation rates, loan approvals, and conditionality.
+We find that European influence is strongly associated with higher loan volumes, more frequent participation, and greater likelihood of loan approval, while US influence is not.
+However, American ties are significantly correlated with fewer conditions attached to loans.
+Additional analysis shows that European influence is not driven by regional or colonial favouritism: neither European nor African borrowers benefit more from European ties—and may, in fact, fare worse—than other countries with comparable alignment.
+These findings complicate existing assumptions about IMF decision-making, highlighting the need to distinguish between types of Western influence.
+They also suggest that Europe wields broader institutional influence over access to IMF resources, while US power manifests more selectively through conditionality.
+In doing so, the paper contributes to debates on hegemony, institutional legitimacy, and intra-Western dynamics in global economic governance."
 keywords: IMF, US hegemony, Europe
@@
 
-Compared to the extensive literature on US dominance in the IMF, scholarship on European influence is relatively limited and often fragmented.
-Many studies underscore that EU member states are institutionally dispersed across multiple constituencies, hindering coherent collective action [@smaghi2004; @miescu2014; @thacker1999].
-Still, other works argue that European countries wield significant structural advantages—including leadership appointments, voting power, and coordination mechanisms—that enable them to shape lending outcomes in line with their interests [@copelovitch2010a; @keating2024; @lutz2014; @stone2004].
-
-European influence reflects a strategic combination of formal power and coordinated action.
-EU countries collectively control over 30% of IMF voting rights, and if they were to unify, they could potentially counterbalance US veto power [@brandner2009a; @leech2005a; @smaghi2004].
-This is reinforced by the long-standing transatlantic norm allocating IMF leadership to a European national [@kahler2001].
-Through platforms such as EURIMF and SCIMF, EU members have harmonised policy positions and influenced loan design.
-During the Eurozone crisis, France and Germany pushed for IMF involvement to stabilise the region while embedding austerity-oriented conditionality that protected their banking systems [@hodson2014; @kickert2019a; @veron2016].
-Lending terms were shaped to reflect creditor priorities, with evidence that conditionality was both intensified or relaxed based on geopolitical alignment [@cain2022; @dreher2015; @lutz2014].
-Beyond Europe, the UK and France have used IMF concessional lending in Africa to maintain post-colonial ties and commercial access [@moyo2024; @stone2004].
-
-Despite Europe’s institutional advantages, scholars remain skeptical of its actual effectiveness in the IMF.
-Internal divergences—such as France’s growth-oriented stance versus Germany’s fiscal conservatism—combined with fragmented constituencies, often impede coherent action and reduce policy influence [@miescu2014; @dreher2015; @copelovitch2010a].
-Moreover, Europe’s disproportionate voting power relative to its population raises legitimacy concerns and is seen as disadvantaging developing countries [@woodward2024; @leech2005a].
-While proposals to consolidate EU seats aim to improve efficiency and counterbalance US dominance, critics warn such reforms may merely reinforce dominance by a few large European states rather than broaden representation [@leech2005a; @copelovitch2010a].
-
-While several studies examine European involvement in IMF operations, most focus on individual cases or regional crises, lacking generalisable conclusions.
-Existing literature often treats Europe as a secondary actor or subsumes it under undifferentiated “Western” influence, thereby overlooking intra-Western variation.
-By analytically separating US and European roles, our study addresses these gaps and offers a comparative framework for understanding divergent patterns of influence within global financial governance.
+In comparison, the literature on European influence is notably smaller.
+The most visible manifestation of European power within the Fund is the tradition of European leadership: all Managing Directors of the IMF since its inception have been European nationals [@kahler2001].
+Scholars have also pointed to Europe-backed lending decisions, such as those during the Eurozone crisis (notably Greece) [@lipscy2018; @hodson2014], and loans to former colonies in Africa [@stone2004].
+However, while these studies offer important insights, they tend to remain case-specific or anecdotal.
+Crucially, they stop short of offering a systematic comparison between American and European influence, leaving open the question of how these two centres of power operate within the IMF.
 

### 2025-07-31 manuscript/manuscript.qmd
delete a para

--- 2025-07-31 manuscript/manuscript.qmd
@@
 In doing so, it contributes to a more nuanced understanding of transatlantic dynamics in global financial governance—highlighting that Western influence may not be as unified or symmetrical as often assumed.
-
-voting power, and coordination mechanisms—that enable them to shape lending outcomes in line with their interests [@copelovitch2010a].
-
-European influence reflects a strategic combination of formal power and coordinated action.
-Lending terms were shaped to reflect creditor priorities, with evidence that conditionality was both intensified or relaxed based on geopolitical alignment [@cain2022; @dreher2015; @lutz2014].
 

### 2025-08-01 manuscript/manuscript.qmd
Merge the standalone para into existing ones

@@
 Many studies underscore that EU member states are institutionally dispersed across multiple constituencies, hindering coherent collective action [@smaghi2004; @miescu2014; @thacker1999].
-Internal differences—such as the often-cited contrast between France’s growth-oriented approach and Germany’s fiscal conservatism—along with the dispersion of EU members across multiple IMF constituencies, are said to limit coherence and reduce policy impact [@smaghi2004; @miescu2014; @dreher2015].
+Internal differences—such as the often-cited contrast between France’s growth-oriented approach and Germany’s fiscal conservatism—along with the dispersion of EU members across multiple IMF constituencies, are said to limit coherence and reduce policy impact [@smaghi2004; @miescu2014; @dreher2015; @copelovitch2010a].
 While some proposals have called for the consolidation of EU seats to strengthen European influence, critics argue this may entrench the dominance of larger EU states without necessarily enhancing representativeness [@leech2005a; @copelovitch2010a].
-Moreover, concerns about Europe’s voting power exceeding its global economic and demographic weight have generated debates about fairness and legitimacy, particularly from developing countries [@woodward2024].
+Moreover, concerns about Europe’s voting power exceeding its global economic and demographic weight have generated debates about fairness and legitimacy, particularly from developing countries [@woodward2024; @leech2005a].
 These institutional tensions may place constraints on Europe’s ability to act decisively, though the extent of such limitations remains contested.
@@
 In doing so, it contributes to a more nuanced understanding of transatlantic dynamics in global financial governance—highlighting that Western influence may not be as unified or symmetrical as often assumed.
-
-Despite Europe’s institutional advantages, scholars remain skeptical of its actual effectiveness in the IMF.
-Internal divergences—such as France’s growth-oriented stance versus Germany’s fiscal conservatism—combined with fragmented constituencies, often impede coherent action and reduce policy influence [@miescu2014; @dreher2015; @copelovitch2010a].
-Moreover, Europe’s disproportionate voting power relative to its population raises legitimacy concerns and is seen as disadvantaging developing countries [@woodward2024; @leech2005a].
-While proposals to consolidate EU seats aim to improve efficiency and counterbalance US dominance, critics warn such reforms may merely reinforce dominance by a few large European states rather than broaden representation [@leech2005a; @copelovitch2010a].
 

### 2025-08-01 manuscript/manuscript.qmd
revied EU paras

@@
 Internal differences—such as the often-cited contrast between France’s growth-oriented approach and Germany’s fiscal conservatism—along with the dispersion of EU members across multiple IMF constituencies, are said to limit coherence and reduce policy impact [@smaghi2004; @miescu2014; @dreher2015; @copelovitch2010a].
+Moreover, some scholars argue that European influence within the IMF is frequently constrained by the need to align with other G5 powers (refers to the US, Japan, the UK, France, Germany) — particularly the United States — which often necessitates policy compromises and limits Europe’s ability to assert unified preferences [@copelovitch2010].
 While some proposals have called for the consolidation of EU seats to strengthen European influence, critics argue this may entrench the dominance of larger EU states without necessarily enhancing representativeness [@leech2005a; @copelovitch2010a].
-Moreover, concerns about Europe’s voting power exceeding its global economic and demographic weight have generated debates about fairness and legitimacy, particularly from developing countries [@woodward2024; @leech2005a].
 These institutional tensions may place constraints on Europe’s ability to act decisively, though the extent of such limitations remains contested.

### 2025-08-04 manuscript/manuscript.qmd
Update Dianyi's email address (#43)

@@
   - name: Dianyi Yang
-    email: d.yang17@lse.ac.uk
+    email: dianyi.yang@politics.ox.ac.uk
     orcid: 0009-0004-4652-3429

### 2025-08-05 code/is_Africa.R
Add outputs to comments

@@ -3,7 +3,7 @@
 # R Script
 # Purpose: This script runs separate regressions for African and non-African countries.
 # Inputs:  data/panel_data_pca.rds
-# Outputs: 
+# Outputs: save/regTable_europe.RData
 
 # SETUP ----------------------------------------------------------------------------------------------------------------
 rm(list = ls()) # Clear workspace

### 2025-08-05 code/is_Europe.R

@@ -3,7 +3,7 @@
 # R Script
 # Purpose: This script runs separate regressions for European and non-European countries.
 # Inputs:  data/panel_data_pca.rds
-# Outputs: 
+# Outputs: save/regTable_europe.RData
 
 # SETUP ----------------------------------------------------------------------------------------------------------------
 rm(list = ls()) # Clear workspace

### 2025-08-05 code/is_Africa.R
Add missing outputs in comments

@@ -3,7 +3,8 @@
 # R Script
 # Purpose: This script runs separate regressions for African and non-African countries.
 # Inputs:  data/panel_data_pca.rds
-# Outputs: save/regTable_europe.RData
+# Outputs: save/regModels_africa.RData
+#          save/regTable_europe.RData
 
 # SETUP ----------------------------------------------------------------------------------------------------------------
 rm(list = ls()) # Clear workspace

### 2025-08-05 code/is_Europe.R

@@ -3,7 +3,8 @@
 # R Script
 # Purpose: This script runs separate regressions for European and non-European countries.
 # Inputs:  data/panel_data_pca.rds
-# Outputs: save/regTable_europe.RData
+# Outputs: save/regModels_europe.RData
+#          save/regTable_europe.RData
 
 # SETUP ----------------------------------------------------------------------------------------------------------------
 rm(list = ls()) # Clear workspace

### 2025-08-05 code/is_Africa.R
Africa: correct output names

@@ -89,7 +89,7 @@ regModels <- list(
   'IMF loan approval' = approval,
   'Number of IMF conditions' = condition
 )
-save(regModels, file = 'save/regModels_europe.RData')
+save(regModels, file = 'save/regModels_africa.RData')
 
 # load('save/regModels.RData') # Load the models again if needed
 msummary(regModels, stars = T) # Rough regression table with all variables and stars
@@ -141,4 +141,4 @@ getTable() # for previewing the table in IDE
 
 resultsTable <- getTable("latex")
 
-save(resultsTable, file = 'save/regTable_europe.RData') # Save the table
+save(resultsTable, file = 'save/regTable_africa.RData') # Save the table

### 2025-08-06 manuscript/manuscript.qmd
Add African table to manuscript
Add GPT robustness intro

@@
 
-## Robustness: Continental Preference?
+## Robustness Checks
+
+While our baseline results indicate that European powers exert substantial influence over IMF lending outcomes, two common explanations in the literature could potentially account for these patterns without implying broad, institutionalised influence.
+
+First, European powers may favour geographically proximate borrowers — in other words, a continental preference — leading to disproportionately favourable treatment of European recipients.
+Second, given the historical legacy of colonialism, particularly in Africa, European powers might be more inclined to support former colonies (colonial preference), especially those with enduring political, economic, or cultural ties.
+
+In this section, we test both explanations by interacting our measure of European influence with indicators for European and African recipients, respectively.
+If either hypothesis holds, we would expect the interaction terms to be positive and significant, indicating stronger European influence in those contexts.
+As shown below, neither hypothesis is supported: in both cases, the interaction terms are insignificant or negative, and our core finding — that European influence is associated with more favourable IMF lending outcomes — remains robust across the recipient groups.
+This suggests that European influence is not geographically or historically confined, but instead reflects a global institutional role within the IMF.
+
+### Continental Preference?
 
@@
 In other words, European powers appear to exercise their influence globally, rather than selectively favouring their regional neighbours.
+
+### Colonial Preference?
+
+```{r}
+#| label: tbl-results-africa
+#| tbl-cap: Effects of US and European Influence on IMF Lending Outcomes (1980–2010), Separating African and Non-African Recipients
+#| tbl-pos: H
+load('save/regTable_africa.RData')
+resultsTable |> tinytable::theme_tt("resize", width = 1)
+```
 

### 2025-08-06 manuscript/manuscript.qmd
Add GPT version of colonial preference

@@
 
+A second possible explanation is that European influence is concentrated in lending to former colonies, particularly in Africa, where historical, linguistic, and economic ties might motivate preferential treatment.
+To assess this, we interact `EUP Influence` with a binary variable for African recipients and re-estimate our main models.
+The results are reported in @tbl-results-africa.
+
 ```{r}
@@
 ```
+
+Contrary to the colonial preference hypothesis, the interaction terms (`EUP Influence` $\times$ `African recipient`) are statistically insignificant in most specifications, and when significant, they are negative — indicating that African recipients, if anything, benefit less from European influence than non-African borrowers with comparable ties.
+Likewise, the `African recipient` variable itself is not associated with systematically more favourable lending outcomes, and in some models is linked to less favourable terms.
+
+These findings parallel our results for European recipients: they indicate that the observed association between European influence and IMF lending is not driven by a concentration of support for countries with historic colonial ties.
+Instead, European influence appears to be exercised in a globally consistent manner, extending beyond specific regional or historical relationships.
 

### 2025-08-06 manuscript/manuscript.qmd
Add pro-Africa references

@@
 First, European powers may favour geographically proximate borrowers — in other words, a continental preference — leading to disproportionately favourable treatment of European recipients.
-Second, given the historical legacy of colonialism, particularly in Africa, European powers might be more inclined to support former colonies (colonial preference), especially those with enduring political, economic, or cultural ties.
+Second, given the historical legacy of colonialism, particularly in Africa, European powers might be more inclined to support former colonies (colonial preference), especially those with enduring political, economic, or cultural ties [@moyo2024; @stone2004; @stone2011, p.51].
 

### 2025-08-07 code/is_Africa.R
is_africa: fix output name

@@ -4,7 +4,7 @@
 # Purpose: This script runs separate regressions for African and non-African countries.
 # Inputs:  data/panel_data_pca.rds
 # Outputs: save/regModels_africa.RData
-#          save/regTable_europe.RData
+#          save/regTable_africa.RData
 
 # SETUP ----------------------------------------------------------------------------------------------------------------
 rm(list = ls()) # Clear workspace

### 2025-08-07 manuscript/manuscript.qmd
clarify outcomes

@@
 In this section, we test both explanations by interacting our measure of European influence with indicators for European and African recipients, respectively.
-If either hypothesis holds, we would expect the interaction terms to be positive and significant, indicating stronger European influence in those contexts.
+If either hypothesis holds, we would expect the interaction terms to be positive and significant for the first three lending outcomes — loan-to-GDP ratio, participation rate, and loan approval, indicating stronger European influence in those contexts.
 As shown below, neither hypothesis is supported: in both cases, the interaction terms are insignificant or negative, and our core finding — that European influence is associated with more favourable IMF lending outcomes — remains robust across the recipient groups.

### 2025-08-07 manuscript/manuscript.qmd
Add stone2004 citation

@@
 
-A second possible explanation is that European influence is concentrated in lending to former colonies, particularly in Africa, where historical, linguistic, and economic ties might motivate preferential treatment.
+A second possible explanation is that European influence is concentrated in lending to former colonies, particularly in Africa, where historical, linguistic, and economic ties might motivate preferential treatment, as suggested by @stone2004.
 To assess this, we interact `EUP Influence` with a binary variable for African recipients and re-estimate our main models.

### 2025-08-07 manuscript/manuscript.qmd
rephrase colonial results interpretation

@@
 
-Contrary to the colonial preference hypothesis, the interaction terms (`EUP Influence` $\times$ `African recipient`) are statistically insignificant in most specifications, and when significant, they are negative — indicating that African recipients, if anything, benefit less from European influence than non-African borrowers with comparable ties.
-Likewise, the `African recipient` variable itself is not associated with systematically more favourable lending outcomes, and in some models is linked to less favourable terms.
+Contrary to the colonial preference hypothesis, the interaction terms (`EUP Influence` $\times$ `African recipient`) are statistically insignificant in most specifications, and when significant (as in the participation rate model), they are *negative* — indicating that ties with Western European powers, if anything, bring fewer benefits to African recipients than to non-African borrowers.
+Similarly, the `African recipient` variable itself is not consistently associated with more favourable IMF lending outcomes; in fact, in some models (participation rate and conditionality), it is significantly correlated with *less* favourable terms.
+Taken together, these results suggest that African recipients do not benefit disproportionately from their historical or post-colonial relationships with European powers—and may, on average, receive less favourable treatment than non-African countries with comparable political ties and economic profiles.
 

### 2025-08-07 code/by_continent.R
Add colonial preference robustness check (#44)
* Add Africa regressions

* Add plot for all continents

* Add outputs to comments

* Add missing outputs in comments

* Africa: correct output names

* save output files

* Add African table to manuscript

* Add GPT robustness intro

* Add GPT version of colonial preference

* Add pro-Africa references

* is_africa: fix output name

* clarify outcomes

* Add stone2004 citation

* rephrase colonial results interpretation

@@ -1,0 +1,191 @@
+######## INFO ########
+# Authors: Dianyi Yang
+# R Script
+# Purpose: This script runs separate regressions for each continent.
+# Inputs:  data/panel_data_pca.rds
+# Outputs:
+
+# SETUP ----------------------------------------------------------------------------------------------------------------
+rm(list = ls()) # Clear workspace
+
+need <- c(
+  'tidyverse',
+  'AER',
+  'fixest',
+  "kableExtra",
+  'modelsummary',
+  "parameters"
+) # list packages needed
+have <- need %in% rownames(installed.packages()) # checks packages you have
+if (any(!have)) {
+  install.packages(need[!have])
+} # install missing packages
+invisible(lapply(need, library, character.only = T))
+
+data <- read_rds("data/panel_data_pca.rds") # Load data
+
+# Run regression ---------------------------------------------------------------------------------------------------------------------------------------------------------------
+
+## Prepare variables -----------------------------------------------------------------------------------------------------
+rhs <- c(
+  "us*continent",
+  "eu*continent",
+  "shstaffl",
+  "shquotal",
+  "lnrgdpnew",
+  "lnrgdpnewsq",
+  "rgdpchnew",
+  "rgdpchnewsquare",
+  "growth1new",
+  "reserv1",
+  "oecd1",
+  "year1980",
+  "year1985",
+  "year1990",
+  "year1995",
+  "year2000"
+)
+make_formula <- function(outcome) {
+  as.formula(paste(outcome, paste(rhs, collapse = " + "), sep = " ~ "))
+}
+
+## tobit regression - IMF loan to GDP ratio --------------------------------------------------------------------------------------------------------------------
+loan <- tobit(
+  make_formula("imfloannew100"),
+  data = data,
+  left = 0,
+  right = Inf,
+  cluster = shcode
+)
+
+msummary(loan, stars = TRUE)
+
+## Tobit regression - IMF participation rate ----------------------------------------------------------------------------------------------------------------
+part <- tobit(
+  make_formula("imf_p"),
+  data = data,
+  left = 0,
+  right = 1,
+  cluster = shcode
+)
+msummary(part, stars = TRUE)
+
+## Probit regression - IMF loan approval --------------------------------------------------------------------------------------------------------------------
+approval <- feglm(
+  make_formula("imf5a"),
+  data = data,
+  family = binomial(link = "probit"),
[…]

### 2025-08-07 code/is_Africa.R

@@ -1,10 +1,10 @@
 ######## INFO ########
 # Authors: Dianyi Yang
 # R Script
-# Purpose: This script runs separate regressions for European and non-European countries.
+# Purpose: This script runs separate regressions for African and non-African countries.
 # Inputs:  data/panel_data_pca.rds
-# Outputs: save/regModels_europe.RData
-#          save/regTable_europe.RData
+# Outputs: save/regModels_africa.RData
+#          save/regTable_africa.RData
 
 # SETUP ----------------------------------------------------------------------------------------------------------------
 rm(list = ls()) # Clear workspace
@@ -23,14 +23,14 @@ invisible(lapply(need, library, character.only = T))
 
 data <- read_rds("data/panel_data_pca.rds") # Load data
 
-# Generate European dummy variable ------------------------------------------------------------------------------------------------
-data$is_european <- ifelse(data$continent == "Europe", 1, 0)
+# Generate African dummy variable ------------------------------------------------------------------------------------------------
+data$is_african <- ifelse(data$continent == "Africa", 1, 0)
 
 # Run regression ---------------------------------------------------------------------------------------------------------------------------------------------------------------
 
 ## Prepare variables -----------------------------------------------------------------------------------------------------
 rhs <- c(
-  "us", "eu*is_european", "shstaffl", "shquotal", 
+  "us", "eu*is_african", "shstaffl", "shquotal", 
   "lnrgdpnew", "lnrgdpnewsq", 
   "rgdpchnew", "rgdpchnewsquare", 
   "growth1new", "reserv1", "oecd1", 
@@ -89,7 +89,7 @@ regModels <- list(
   'IMF loan approval' = approval,
   'Number of IMF conditions' = condition
 )
-save(regModels, file = 'save/regModels_europe.RData')
+save(regModels, file = 'save/regModels_africa.RData')
 
 # load('save/regModels.RData') # Load the models again if needed
 msummary(regModels, stars = T) # Rough regression table with all variables and stars
@@ -103,8 +103,8 @@ source("helper/f_test.R")
 coefmap <- c(
   'us' = 'USA Influence',
   'eu' = 'EUP Influence',
-  'eu:is_european' = 'EUP Influence $\\times$ \\newline European recipient',
-  'is_european' = 'European recipient',
+  'eu:is_african' = 'EUP Influence $\\times$ \\newline African recipient',
[…]

### 2025-08-12 manuscript/manuscript.qmd
Add pisany-ferry to European part

@@
 One plausible alternative explanation for the strong association between European influence and favourable IMF lending outcomes is that European powers might be disproportionately favouring fellow European recipients, in contrast to the United States, for which such continental preference may not apply.
-This suspicion is not unfounded: during the Eurozone crisis, the IMF partnered with the European Commission and the European Central Bank—collectively known as the Troika—to jointly extend financial assistance to Greece and other European economies.
+This suspicion is not unfounded: during the Eurozone crisis, the IMF partnered with the European Commission and the European Central Bank—collectively known as the Troika—to jointly extend financial assistance to Greece and other European economies [@pisani-ferry2013].
 

### 2025-08-13 manuscript/manuscript.qmd
Add hodson to European section

--- 2025-08-07 manuscript/manuscript.qmd
@@
 One plausible alternative explanation for the strong association between European influence and favourable IMF lending outcomes is that European powers might be disproportionately favouring fellow European recipients, in contrast to the United States, for which such continental preference may not apply.
-This suspicion is not unfounded: during the Eurozone crisis, the IMF partnered with the European Commission and the European Central Bank—collectively known as the Troika—to jointly extend financial assistance to Greece and other European economies.
+This suspicion is not unfounded: during the Eurozone crisis, the IMF partnered with the European Commission and the European Central Bank—collectively known as the Troika—to jointly extend financial assistance to Greece and other European economies [@hodson2014; @pisani-ferry2013].
 

### 2025-08-13 manuscript/manuscript.qmd
remove "explanations in the literature"

@@
 
-While our baseline results indicate that European powers exert substantial influence over IMF lending outcomes, two common explanations in the literature could potentially account for these patterns without implying broad, institutionalised influence.
+While our baseline results indicate that European powers exert substantial influence over IMF lending outcomes, two plausible explanations could potentially account for these patterns without implying broad, institutionalised influence.
 

### 2025-08-13 manuscript/manuscript.qmd
Reframe the short hypothesis part

@@
 
-First, European powers may favour geographically proximate borrowers — in other words, a continental preference — leading to disproportionately favourable treatment of European recipients.
+First, European powers may favour geographically proximate borrowers — a *continental preference* — resulting in disproportionately favourable treatment of European recipients.
+Such behaviour might be expected given the EU’s special Troika arrangements with the IMF during the Eurozone crisis [@hodson2014; @pisani-ferry2013].
 Second, given the historical legacy of colonialism, particularly in Africa, European powers might be more inclined to support former colonies (colonial preference), especially those with enduring political, economic, or cultural ties [@moyo2024; @stone2004; @stone2011, p.51].

### 2025-08-13 manuscript/manuscript.qmd
Add timing disclaimer

@@
 This suspicion is not unfounded: during the Eurozone crisis, the IMF partnered with the European Commission and the European Central Bank—collectively known as the Troika—to jointly extend financial assistance to Greece and other European economies [@hodson2014; @pisani-ferry2013].
+Although the Eurozone crisis itself overlaps only slightly with the end of our sample period (1980-2010), and the Troika’s coordinated lending response occurred largely afterwards, we nevertheless test this hypothesis to assess whether similar continental preference dynamics could explain our findings.
 

### 2025-08-13 manuscript/manuscript.qmd
Remove "potentially"

@@
 
-While our baseline results indicate that European powers exert substantial influence over IMF lending outcomes, two plausible explanations could potentially account for these patterns without implying broad, institutionalised influence.
+While our baseline results indicate that European powers exert substantial influence over IMF lending outcomes, two plausible explanations could account for these patterns without implying broad, institutionalised influence.
 

### 2025-08-20 manuscript/manuscript.qmd
Add footnote explaining youssef2023

@@
 date: last-modified
+date-format: |
+    | DD MMMM YYYY
 authors:
@@
     orcid: 0009-0004-4652-3429
-    correspondence: true
+    corresponding: true
+    affiliations:
+      - ref: ox    
   - name: Tong Hu
     orcid: 0009-0003-5234-8164
+    affiliations:
+      - ref: bw
+    equal-contributor: true
   - name: Zhenyi Chen
     orcid: 0009-0008-2910-5187
+    affiliations:
+      - ref: rmu
+    equal-contributor: true
+affiliations:
+  - id: ox
+    name: Department of Politics and International Relations, University of Oxford
+  - id: bw
+    name: School of International Relations and Diplomacy, Beijing Foreign Studies University
+  - id: rmu
+    name: School of International Studies, Renmin University of China
 format:
@@
 Yet, much of this literature has remained US-centric, often using only American alignment as the independent variable of interest [@thacker1999; @oatley2004; @dreher2009].
-Other studies pool European indicators with US ones under the broader category of “Western” influence, thus masking potential differences between the two [@lipscy2018; @barro2005].
-These ties have been shown to correlate with more favourable IMF lending outcomes—larger loans, higher approval rates, and fewer or less stringent conditions [@lipscy2018; @barro2005; @dreher2009].
+Other studies pool European indicators with US ones under the broader category of “Western” influence, thus masking potential differences between the two [@lipscy2018; @barro2005; @youssef2023][^2].
+These ties have been shown to correlate with more favourable IMF lending outcomes—larger loans, higher approval rates, and fewer or less stringent conditions [@lipscy2018; @barro2005; @dreher2009; @youssef2023].
 Yet, despite the availability of these tools, comparative assessments between the US and Europe remain largely absent from this literature.
+
+[^2]: @youssef2023 also include China as a major shareholder when calculating political influence.
 
@@
 
-# Empirical Analysis
+# Empirical Strategy
 
@@
 This paper uses the dataset compiled by @lipscy2018, which extends the @barro2005 dataset through 2010.
-It provides panel data for 130 countries in five-year intervals from 1980 to 2010[^2], including information on IMF borrowing and participation outcomes, economic indicators, and measures of each country's relationship with the United States and major European powers.
+It provides panel data for 130 countries in five-year intervals from 1980 to 2010[^3], including information on IMF borrowing and participation outcomes, economic indicators, and measures of each country's relationship with the United States and major European powers.
 
-[^2]: e.g. 1980-1985, 1985-1990.
+[^3]: e.g. 1980-1985, 1985-1990.
 
@@
 These propensities are based on widely used proxies from the existing literature, which reflect the degree of a country’s political and economic alignment with the US or Europe [@lipscy2018].
-This approach assumes that both the US and European powers[^3] are more likely to advocate for lending decisions that benefit their close partners.
+This approach assumes that both the US and European powers[^4] are more likely to advocate for lending decisions that benefit their close partners.
 
-[^3]: For operationalisation, we focus only on ties to the three largest Western European powers—namely, the United Kingdom, France, and Germany—and use their average values.
+[^4]: For operationalisation, we focus only on ties to the three largest Western European powers—namely, the United Kingdom, France, and Germany—and use their average values.
 This approach reflects the assumption that smaller European countries are unlikely to exert meaningful influence over IMF decisions on behalf of their partners.
@@
 
--   UN General Assembly voting similarity[^4]
--   Bilateral trade intensity[^5]
--   Banking sector exposure[^6]
+-   UN General Assembly voting similarity[^5]
+-   Bilateral trade intensity[^6]
+-   Banking sector exposure[^7]
 
-[^4]: Originallly used in @thacker1999.
+[^5]: Originallly used in @thacker1999.
 
-[^5]: From @barro2005.
+[^6]: From @barro2005.
 
-[^6]: See @broz2006; @lipscy2017; @oatley2004.
+[^7]: See @broz2006; @lipscy2017; @oatley2004.
 
@@
 
-However, as noted by @lipscy2018 and @belsley2004, including multiple proxies for the same latent construct in a single regression model leads to multicollinearity, which reduces statistical power and inflates standard errors, as the variables are highly correlated in our data[^7].
+However, as noted by @lipscy2018 and @belsley2004, including multiple proxies for the same latent construct in a single regression model leads to multicollinearity, which reduces statistical power and inflates standard errors, as the variables are highly correlated in our data[^8].
 To address this issue, we apply principal component analysis (PCA) to reduce the dimensionality of the data [@dunteman2008].
-Specifically, we extract the first principal component — a linear combination of the original variables that captures the greatest share of variance — to serve as a single measure of influence[^8].
+Specifically, we extract the first principal component — a linear combination of the original variables that captures the greatest share of variance — to serve as a single measure of influence[^9].
 
-[^7]: See Appendix [-@sec-corr].
+[^8]: See Appendix [-@sec-corr].
 
-[^8]: The first principal components capture the majority of the variance (`r format(EU_PC1_percentage, digits = 1)`% for Europe and `r format(US_PC1_percentage, digits = 1)`% for the United States).
+[^9]: The first principal components capture the majority of the variance (`r format(EU_PC1_percentage, digits = 1)`% for Europe and `r format(US_PC1_percentage, digits = 1)`% for the United States).
 See Appendix [-@sec-scree] for scree plots and additional details.
@@
 
-The selection of covariates, including their polynomial transformations[^9], closely follows @lipscy2018 and @barro2005 to ensure comparability with previous findings.
+The selection of covariates, including their polynomial transformations[^10], closely follows @lipscy2018 and @barro2005 to ensure comparability with previous findings.
 
-[^9]: Quadratic terms are included for GDP and per capita GDP to capture potential non-linear effects. See [Model Specification](#sec-model) for details.
+[^10]: Quadratic terms are included for GDP and per capita GDP to capture potential non-linear effects. See [Model Specification](#sec-model) for details.
 
@@
 -   $\lambda_t$ denotes five-year period fixed effects.
--   $\epsilon_{it}$ is the error term, assumed to be independently and identically distributed as $\mathcal{N}(0, \sigma^2)$[^10].
+-   $\epsilon_{it}$ is the error term, assumed to be independently and identically distributed as $\mathcal{N}(0, \sigma^2)$[^11].
 
-[^10]: Country fixed effects are not included because the number of time periods ($T$) is small relative to the number of countries ($N$).
+[^11]: Country fixed effects are not included because the number of time periods ($T$) is small relative to the number of countries ($N$).
 In such cases, including country fixed effects would lead to inconsistent estimates of the disturbance variance $\sigma$ [@greene2004].
@@
 This allows us to conveniently reuse the same latent variable specification $Y^*_{it}$ as in the Tobit model.
-The key differences are that the error term is now assumed to follow a standard normal distribution $\mathcal{N}(0, 1)$, and that the latent outcome maps to a binary observed outcome as follows[^11]:
+The key differences are that the error term is now assumed to follow a standard normal distribution $\mathcal{N}(0, 1)$, and that the latent outcome maps to a binary observed outcome as follows[^12]:
 
-[^11]: Alternatively, the Probit model can be expressed in terms of probabilities as $\Pr(Y_{it}=1)=\Phi(\alpha + \beta_1 \text{USA}_{it} + \beta_2 \text{EUP}_{it} + \gamma^\top \mathbf{P}_{it} + \delta^\top \mathbf{Z}_{it} + \lambda_t)$, where $\Phi(\cdot)$ is the standard normal cumulative distribution function.
+[^12]: Alternatively, the Probit model can be expressed in terms of probabilities as $\Pr(Y_{it}=1)=\Phi(\alpha + \beta_1 \text{USA}_{it} + \beta_2 \text{EUP}_{it} + \gamma^\top \mathbf{P}_{it} + \delta^\top \mathbf{Z}_{it} + \lambda_t)$, where $\Phi(\cdot)$ is the standard normal cumulative distribution function.
 
@@
 
-## Results
+# Results
 
@@
 In contrast, similar ties with the United States do not yield statistically significant effects on these outcomes.
-The differences between American and European influence on the first two outcomes are statistically significant at the 5% level[^12].
+The differences between American and European influence on the first two outcomes are statistically significant at the 5% level[^13].
 
-[^12]: Based on tests of equality for the null hypothesis that $\text{USA} = \text{EUP}$.
+[^13]: Based on tests of equality for the null hypothesis that $\text{USA} = \text{EUP}$.
 $F$-tests are used for the linear models; for the Probit model, the test corresponds to an asymptotic Chi-squared statistic, which is appropriate given the large sample size ($N = 522$).
@@
 load('save/regTable.RData')
-resultsTable |> tinytable::theme_tt("resize", width = 1)
+resultsTable |> tinytable::theme_latex(resize_width = 1, resize_direction = "down")
 ```
 
-## Robustness Checks
+# Robustness Checks
 
@@
 
-### Continental Preference?
+## Continental Preference?
 
@@
 load('save/regTable_europe.RData')
-resultsTable |> tinytable::theme_tt("resize", width = 1)
+resultsTable |> tinytable::theme_latex(resize_width = 1, resize_direction = "down")
 ```
@@
 
-### Colonial Preference?
+## Colonial Preference?
 
@@
 load('save/regTable_africa.RData')
-resultsTable |> tinytable::theme_tt("resize", width = 1)
+resultsTable |> tinytable::theme_latex(resize_width = 1, resize_direction = "down")
 ```

### 2025-08-20 manuscript/manuscript.qmd
Add youssef2023 to economic indicators

@@
 
-We include a range of economic performance indicators that have been shown to systematically affect IMF lending decisions [@barro2005; @knight1997].
+We include a range of economic performance indicators that have been shown to systematically affect IMF lending decisions [@barro2005; @knight1997; @youssef2023].
 To capture these dynamics, we include GDP, per capita GDP, and lagged GDP growth as proxies for macroeconomic performance.

### 2025-08-20 manuscript/manuscript.qmd
Add guimaraes (2021) to lit review

@@
 These studies typically proxy the strategic interests of major powers using indicators such as UN voting alignment [@thacker1999; @dreher2009], bilateral trade intensity [@barro2005], and financial exposure [@oatley2004].
-Yet, much of this literature has remained US-centric, often using only American alignment as the independent variable of interest [@thacker1999; @oatley2004; @dreher2009].
-Other studies pool European indicators with US ones under the broader category of “Western” influence, thus masking potential differences between the two [@lipscy2018; @barro2005; @youssef2023][^2].
-These ties have been shown to correlate with more favourable IMF lending outcomes—larger loans, higher approval rates, and fewer or less stringent conditions [@lipscy2018; @barro2005; @dreher2009; @youssef2023].
+Yet, much of this literature has remained US-centric, often using only American alignment as the independent variable of interest [@thacker1999; @oatley2004; @dreher2009; @guimaraes2021].
+Other studies pool European indicators with US ones under the broader category of “Western” influence, thus masking potential differences between the two [@lipscy2018; @barro2005; @youssef2023; @guimaraes2021][^2].
+These ties have been shown to correlate with more favourable IMF lending outcomes—larger loans, higher approval rates, and fewer or less stringent conditions [@lipscy2018; @barro2005; @dreher2009; @youssef2023; @guimaraes2021].
 Yet, despite the availability of these tools, comparative assessments between the US and Europe remain largely absent from this literature.

### 2025-08-20 manuscript/manuscript.qmd
Add harrigan to qualitative studies

@@
 [@stone2011, p.57].
-Numerous case studies further demonstrate instances of US political pressure resulting in lenient loan conditions for key geopolitical allies, including Zaire, Egypt, Mexico, Argentina, and South Korea [@vreeland2009; @copelovitch2010a; @lipscy2018].
+Numerous case studies further demonstrate instances of US political pressure resulting in lenient loan conditions for key geopolitical allies, including Zaire, Egypt, Mexico, Argentina, South Korea and MENA countries [@vreeland2009; @copelovitch2010a; @lipscy2018; @harrigan2006].
 This pattern of influence is consistent with the United States’ broader role across international financial institutions, particularly the World Bank, where American dominance is widely acknowledged [@wade2002; @woods2003; @clark2020].

### 2025-08-20 manuscript/manuscript.qmd
Apply GPT review suggestions

--- 2025-08-20 manuscript/manuscript.qmd
@@
 [@stone2011, p.57].
-Numerous case studies further demonstrate instances of US political pressure resulting in lenient loan conditions for key geopolitical allies, including Zaire, Egypt, Mexico, Argentina, and South Korea [@vreeland2009; @copelovitch2010a; @lipscy2018].
+Numerous case studies further demonstrate instances of US political pressure resulting in lenient loan conditions for key geopolitical allies, including Zaire, Egypt, Mexico, Argentina, South Korea, and MENA countries [@vreeland2009; @copelovitch2010a; @lipscy2018; @harrigan2006].
 This pattern of influence is consistent with the United States’ broader role across international financial institutions, particularly the World Bank, where American dominance is widely acknowledged [@wade2002; @woods2003; @clark2020].
@@
 
-[^5]: Originallly used in @thacker1999.
+[^5]: Originally used in @thacker1999.
 

### 2025-08-21 manuscript/manuscript.qmd
Move Results and Robustness out of Empirical section (#50)

--- 2025-08-20 manuscript/manuscript.qmd
@@
 date: last-modified
-date-format: |
-    | DD MMMM YYYY
 authors:
@@
     orcid: 0009-0004-4652-3429
-    corresponding: true
-    affiliations:
-      - ref: ox    
+    correspondence: true
   - name: Tong Hu
     orcid: 0009-0003-5234-8164
-    affiliations:
-      - ref: bw
-    equal-contributor: true
   - name: Zhenyi Chen
     orcid: 0009-0008-2910-5187
-    affiliations:
-      - ref: rmu
-    equal-contributor: true
-affiliations:
-  - id: ox
-    name: Department of Politics and International Relations, University of Oxford
-  - id: bw
-    name: School of International Relations and Diplomacy, Beijing Foreign Studies University
-  - id: rmu
-    name: School of International Studies, Renmin University of China
 format:
@@
 These studies typically proxy the strategic interests of major powers using indicators such as UN voting alignment [@thacker1999; @dreher2009], bilateral trade intensity [@barro2005], and financial exposure [@oatley2004].
-Yet, much of this literature has remained US-centric, often using only American alignment as the independent variable of interest [@thacker1999; @oatley2004; @dreher2009; @guimaraes2021].
-Other studies pool European indicators with US ones under the broader category of “Western” influence, thus masking potential differences between the two [@lipscy2018; @barro2005; @youssef2023; @guimaraes2021][^2].
-These ties have been shown to correlate with more favourable IMF lending outcomes—larger loans, higher approval rates, and fewer or less stringent conditions [@lipscy2018; @barro2005; @dreher2009; @youssef2023; @guimaraes2021].
+Yet, much of this literature has remained US-centric, often using only American alignment as the independent variable of interest [@thacker1999; @oatley2004; @dreher2009].
+Other studies pool European indicators with US ones under the broader category of “Western” influence, thus masking potential differences between the two [@lipscy2018; @barro2005].
+These ties have been shown to correlate with more favourable IMF lending outcomes—larger loans, higher approval rates, and fewer or less stringent conditions [@lipscy2018; @barro2005; @dreher2009].
 Yet, despite the availability of these tools, comparative assessments between the US and Europe remain largely absent from this literature.
-
-[^2]: @youssef2023 also include China as a major shareholder when calculating political influence.
 
@@
 This paper uses the dataset compiled by @lipscy2018, which extends the @barro2005 dataset through 2010.
-It provides panel data for 130 countries in five-year intervals from 1980 to 2010[^3], including information on IMF borrowing and participation outcomes, economic indicators, and measures of each country's relationship with the United States and major European powers.
+It provides panel data for 130 countries in five-year intervals from 1980 to 2010[^2], including information on IMF borrowing and participation outcomes, economic indicators, and measures of each country's relationship with the United States and major European powers.
 
-[^3]: e.g. 1980-1985, 1985-1990.
+[^2]: e.g. 1980-1985, 1985-1990.
 
@@
 These propensities are based on widely used proxies from the existing literature, which reflect the degree of a country’s political and economic alignment with the US or Europe [@lipscy2018].
-This approach assumes that both the US and European powers[^4] are more likely to advocate for lending decisions that benefit their close partners.
+This approach assumes that both the US and European powers[^3] are more likely to advocate for lending decisions that benefit their close partners.
 
-[^4]: For operationalisation, we focus only on ties to the three largest Western European powers—namely, the United Kingdom, France, and Germany—and use their average values.
+[^3]: For operationalisation, we focus only on ties to the three largest Western European powers—namely, the United Kingdom, France, and Germany—and use their average values.
 This approach reflects the assumption that smaller European countries are unlikely to exert meaningful influence over IMF decisions on behalf of their partners.
@@
 
--   UN General Assembly voting similarity[^5]
--   Bilateral trade intensity[^6]
--   Banking sector exposure[^7]
+-   UN General Assembly voting similarity[^4]
+-   Bilateral trade intensity[^5]
+-   Banking sector exposure[^6]
 
-[^5]: Originallly used in @thacker1999.
+[^4]: Originallly used in @thacker1999.
 
-[^6]: From @barro2005.
+[^5]: From @barro2005.
 
-[^7]: See @broz2006; @lipscy2017; @oatley2004.
+[^6]: See @broz2006; @lipscy2017; @oatley2004.
 
@@
 
-However, as noted by @lipscy2018 and @belsley2004, including multiple proxies for the same latent construct in a single regression model leads to multicollinearity, which reduces statistical power and inflates standard errors, as the variables are highly correlated in our data[^8].
+However, as noted by @lipscy2018 and @belsley2004, including multiple proxies for the same latent construct in a single regression model leads to multicollinearity, which reduces statistical power and inflates standard errors, as the variables are highly correlated in our data[^7].
 To address this issue, we apply principal component analysis (PCA) to reduce the dimensionality of the data [@dunteman2008].
-Specifically, we extract the first principal component — a linear combination of the original variables that captures the greatest share of variance — to serve as a single measure of influence[^9].
+Specifically, we extract the first principal component — a linear combination of the original variables that captures the greatest share of variance — to serve as a single measure of influence[^8].
 
-[^8]: See Appendix [-@sec-corr].
+[^7]: See Appendix [-@sec-corr].
 
-[^9]: The first principal components capture the majority of the variance (`r format(EU_PC1_percentage, digits = 1)`% for Europe and `r format(US_PC1_percentage, digits = 1)`% for the United States).
+[^8]: The first principal components capture the majority of the variance (`r format(EU_PC1_percentage, digits = 1)`% for Europe and `r format(US_PC1_percentage, digits = 1)`% for the United States).
 See Appendix [-@sec-scree] for scree plots and additional details.
@@
 
-We include a range of economic performance indicators that have been shown to systematically affect IMF lending decisions [@barro2005; @knight1997; @youssef2023].
+We include a range of economic performance indicators that have been shown to systematically affect IMF lending decisions [@barro2005; @knight1997].
 To capture these dynamics, we include GDP, per capita GDP, and lagged GDP growth as proxies for macroeconomic performance.
@@
 
-The selection of covariates, including their polynomial transformations[^10], closely follows @lipscy2018 and @barro2005 to ensure comparability with previous findings.
+The selection of covariates, including their polynomial transformations[^9], closely follows @lipscy2018 and @barro2005 to ensure comparability with previous findings.
 
-[^10]: Quadratic terms are included for GDP and per capita GDP to capture potential non-linear effects. See [Model Specification](#sec-model) for details.
+[^9]: Quadratic terms are included for GDP and per capita GDP to capture potential non-linear effects. See [Model Specification](#sec-model) for details.
 
@@
 -   $\lambda_t$ denotes five-year period fixed effects.
--   $\epsilon_{it}$ is the error term, assumed to be independently and identically distributed as $\mathcal{N}(0, \sigma^2)$[^11].
+-   $\epsilon_{it}$ is the error term, assumed to be independently and identically distributed as $\mathcal{N}(0, \sigma^2)$[^10].
 
-[^11]: Country fixed effects are not included because the number of time periods ($T$) is small relative to the number of countries ($N$).
+[^10]: Country fixed effects are not included because the number of time periods ($T$) is small relative to the number of countries ($N$).
 In such cases, including country fixed effects would lead to inconsistent estimates of the disturbance variance $\sigma$ [@greene2004].
@@
 This allows us to conveniently reuse the same latent variable specification $Y^*_{it}$ as in the Tobit model.
-The key differences are that the error term is now assumed to follow a standard normal distribution $\mathcal{N}(0, 1)$, and that the latent outcome maps to a binary observed outcome as follows[^12]:
+The key differences are that the error term is now assumed to follow a standard normal distribution $\mathcal{N}(0, 1)$, and that the latent outcome maps to a binary observed outcome as follows[^11]:
 
-[^12]: Alternatively, the Probit model can be expressed in terms of probabilities as $\Pr(Y_{it}=1)=\Phi(\alpha + \beta_1 \text{USA}_{it} + \beta_2 \text{EUP}_{it} + \gamma^\top \mathbf{P}_{it} + \delta^\top \mathbf{Z}_{it} + \lambda_t)$, where $\Phi(\cdot)$ is the standard normal cumulative distribution function.
+[^11]: Alternatively, the Probit model can be expressed in terms of probabilities as $\Pr(Y_{it}=1)=\Phi(\alpha + \beta_1 \text{USA}_{it} + \beta_2 \text{EUP}_{it} + \gamma^\top \mathbf{P}_{it} + \delta^\top \mathbf{Z}_{it} + \lambda_t)$, where $\Phi(\cdot)$ is the standard normal cumulative distribution function.
 
@@
 In contrast, similar ties with the United States do not yield statistically significant effects on these outcomes.
-The differences between American and European influence on the first two outcomes are statistically significant at the 5% level[^13].
+The differences between American and European influence on the first two outcomes are statistically significant at the 5% level[^12].
 
-[^13]: Based on tests of equality for the null hypothesis that $\text{USA} = \text{EUP}$.
+[^12]: Based on tests of equality for the null hypothesis that $\text{USA} = \text{EUP}$.
 $F$-tests are used for the linear models; for the Probit model, the test corresponds to an asymptotic Chi-squared statistic, which is appropriate given the large sample size ($N = 522$).
@@
 load('save/regTable.RData')
-resultsTable |> tinytable::theme_latex(resize_width = 1, resize_direction = "down")
+resultsTable |> tinytable::theme_tt("resize", width = 1)
 ```
@@
 load('save/regTable_europe.RData')
-resultsTable |> tinytable::theme_latex(resize_width = 1, resize_direction = "down")
+resultsTable |> tinytable::theme_tt("resize", width = 1)
 ```
@@
 load('save/regTable_africa.RData')
-resultsTable |> tinytable::theme_latex(resize_width = 1, resize_direction = "down")
+resultsTable |> tinytable::theme_tt("resize", width = 1)
 ```

### 2025-08-21 manuscript/manuscript.qmd
Fix corresponding

@@
     orcid: 0009-0004-4652-3429
-    correspondence: true
+    corresponding: true
   - name: Tong Hu
@@
 load('save/regTable.RData')
-resultsTable |> tinytable::theme_tt("resize", width = 1)
+resultsTable |> tinytable::theme_latex(resize_width = 1, resize_direction = "down")
 ```

### 2025-08-21 manuscript/manuscript.qmd
Add affiliations

@@
     corresponding: true
+    affiliations:
+      - ref: ox    
   - name: Tong Hu
     orcid: 0009-0003-5234-8164
+    affiliations:
+      - ref: bw
+    equal-contributor: true
   - name: Zhenyi Chen
     orcid: 0009-0008-2910-5187
+    affiliations:
+      - ref: rmu
+    equal-contributor: true
+affiliations:
+  - id: ox
+    name: Department of Politics and International Relations, University of Oxford
+  - id: bw
+    name: School of International Relations and Diplomacy, Beijing Foreign Studies University
+  - id: rmu
+    name: School of International Studies, Renmin University of China
 format:

### 2025-08-22 code/PCA.R
PCA.R: use library only for renv

@@ -13,20 +13,15 @@
 # SETUP ----------------------------------------------------------------------------------------------------------------
 rm(list = ls()) # Clear workspace
 
-need <- c(
-  'tidyverse',
-  'haven',
-  'AER',
-  'fixest',
-  "kableExtra",
-  'modelsummary',
-  'corrplot',
-  "ggfortify",
-  "parameters"
-) # list packages needed
-have <- need %in% rownames(installed.packages()) # checks packages you have
-if (any(!have)) install.packages(need[!have]) # install missing packages
-invisible(lapply(need, library, character.only = T))
+library('tidyverse')
+library('haven')
+library('AER')
+library('fixest')
+library("kableExtra")
+library('modelsummary')
+library('corrplot')
+library("ggfortify")
+library("parameters")
 
 data <- read_rds("data/panel_data.rds") # Load data
 

### 2025-08-22 code/PCA.R
PCA: reformat

@@ -102,9 +102,8 @@ save(US, EU, file = "save/PCA.RData") # Save PCA results
     y = "Percentage of explained variances"
   ) +
   scale_x_continuous(breaks = 1:3, limits = c(1, 3.1)) +
-  scale_y_continuous(labels = scales::percent,
-                     limits = c(0, 0.55)) +
-  theme_bw()+
+  scale_y_continuous(labels = scales::percent, limits = c(0, 0.55)) +
+  theme_bw() +
   theme(panel.grid.minor.x = element_blank()))
 
 (scree_EU <- data.frame(
@@ -125,9 +124,8 @@ save(US, EU, file = "save/PCA.RData") # Save PCA results
     y = "Percentage of explained variances"
   ) +
   scale_x_continuous(breaks = 1:3, limits = c(1, 3.1)) +
-  scale_y_continuous(labels = scales::percent,
-                     limits = c(0, 0.65)) +
-  theme_bw()+
+  scale_y_continuous(labels = scales::percent, limits = c(0, 0.65)) +
+  theme_bw() +
   theme(panel.grid.minor.x = element_blank()))
 
 save(scree_US, scree_EU, file = "save/scree.RData") # Save scree plot results
@@ -139,17 +137,20 @@ data$eu <- predict(EU, data)[, 1] |> scale()
 data$pca <- -predict(PCA, data)[, 1]
 
 write_rds(data, "data/panel_data_pca.rds")
+data <- read_rds("data/panel_data_pca.rds") # Reload data with PCA variables
 
 ## draw a kernel density plot for US and EU PC1 components  -----------------------------------------------------------------------------------------------------------------------------------
-data |> 
+data |>
   select(us, eu) |>
   na.omit() |>
   pivot_longer(cols = everything(), names_to = 'region', values_to = 'value') |>
   ggplot(aes(x = value, fill = region)) +
   geom_density(alpha = 0.5) +
-  labs(title = "Kernel Density Plot of US and EU variables",
-       x = "PC1",
-       y = "Density") +
+  labs(
+    title = "Kernel Density Plot of US and EU variables",
+    x = "PC1",
+    y = "Density"
+  ) +
   theme_minimal()
 
 ## plot correlation between our variables with the original one. ---------------------------------------------------------------------------------------------------------------
[…]

### 2025-08-22 code/is_Europe.R
is_europe: reformat + use library()

@@ -9,17 +9,13 @@
 # SETUP ----------------------------------------------------------------------------------------------------------------
 rm(list = ls()) # Clear workspace
 
-need <- c(
-  'tidyverse',
-  'AER',
-  'fixest',
-  "kableExtra",
-  'modelsummary',
-  "parameters"
-) # list packages needed
-have <- need %in% rownames(installed.packages()) # checks packages you have
-if (any(!have)) install.packages(need[!have]) # install missing packages
-invisible(lapply(need, library, character.only = T))
+
+library('tidyverse')
+library('AER')
+library('fixest')
+library("kableExtra")
+library('modelsummary')
+library("parameters")
 
 data <- read_rds("data/panel_data_pca.rds") # Load data
 
@@ -30,11 +26,22 @@ data$is_european <- ifelse(data$continent == "Europe", 1, 0)
 
 ## Prepare variables -----------------------------------------------------------------------------------------------------
 rhs <- c(
-  "us", "eu*is_european", "shstaffl", "shquotal", 
-  "lnrgdpnew", "lnrgdpnewsq", 
-  "rgdpchnew", "rgdpchnewsquare", 
-  "growth1new", "reserv1", "oecd1", 
-  "year1980", "year1985", "year1990", "year1995", "year2000"
+  "us",
+  "eu*is_european",
+  "shstaffl",
+  "shquotal",
+  "lnrgdpnew",
+  "lnrgdpnewsq",
+  "rgdpchnew",
+  "rgdpchnewsquare",
+  "growth1new",
+  "reserv1",
+  "oecd1",
+  "year1980",
+  "year1985",
+  "year1990",
+  "year1995",
+  "year2000"
 )
 make_formula <- function(outcome) {
   as.formula(paste(outcome, paste(rhs, collapse = " + "), sep = " ~ "))

### 2025-08-22 code/is_Africa.R
process is_africa

@@ -9,17 +9,12 @@
 # SETUP ----------------------------------------------------------------------------------------------------------------
 rm(list = ls()) # Clear workspace
 
-need <- c(
-  'tidyverse',
-  'AER',
-  'fixest',
-  "kableExtra",
-  'modelsummary',
-  "parameters"
-) # list packages needed
-have <- need %in% rownames(installed.packages()) # checks packages you have
-if (any(!have)) install.packages(need[!have]) # install missing packages
-invisible(lapply(need, library, character.only = T))
+library('tidyverse')
+library('AER')
+library('fixest')
+library("kableExtra")
+library('modelsummary')
+library("parameters")
 
 data <- read_rds("data/panel_data_pca.rds") # Load data
 
@@ -30,11 +25,22 @@ data$is_african <- ifelse(data$continent == "Africa", 1, 0)
 
 ## Prepare variables -----------------------------------------------------------------------------------------------------
 rhs <- c(
-  "us", "eu*is_african", "shstaffl", "shquotal", 
-  "lnrgdpnew", "lnrgdpnewsq", 
-  "rgdpchnew", "rgdpchnewsquare", 
-  "growth1new", "reserv1", "oecd1", 
-  "year1980", "year1985", "year1990", "year1995", "year2000"
+  "us",
+  "eu*is_african",
+  "shstaffl",
+  "shquotal",
+  "lnrgdpnew",
+  "lnrgdpnewsq",
+  "rgdpchnew",
+  "rgdpchnewsquare",
+  "growth1new",
+  "reserv1",
+  "oecd1",
+  "year1980",
+  "year1985",
+  "year1990",
+  "year1995",
+  "year2000"
 )
 make_formula <- function(outcome) {
   as.formula(paste(outcome, paste(rhs, collapse = " + "), sep = " ~ "))

### 2025-08-22 code/PCA.R
PCA.R: add data loading comment

@@ -137,7 +137,7 @@ data$eu <- predict(EU, data)[, 1] |> scale()
 data$pca <- -predict(PCA, data)[, 1]
 
 write_rds(data, "data/panel_data_pca.rds")
-data <- read_rds("data/panel_data_pca.rds") # Reload data with PCA variables
+# data <- read_rds("data/panel_data_pca.rds") # Reload data with PCA variables
 
 ## draw a kernel density plot for US and EU PC1 components  -----------------------------------------------------------------------------------------------------------------------------------
 data |>

### 2025-08-22 code/is_Europe.R
is_europe: fix reg loading comment

@@ -98,7 +98,7 @@ regModels <- list(
 )
 save(regModels, file = 'save/regModels_europe.RData')
 
-# load('save/regModels.RData') # Load the models again if needed
+# load('save/regModels_europe.RData') # Load the models again if needed
 msummary(regModels, stars = T) # Rough regression table with all variables and stars
 
 ## gof generation -----------------------------------------------------------------------------------------------------------------------------------------

### 2025-08-22 code/is_Africa.R
is_Africa: fix reg loading comment

@@ -97,7 +97,7 @@ regModels <- list(
 )
 save(regModels, file = 'save/regModels_africa.RData')
 
-# load('save/regModels.RData') # Load the models again if needed
+# load('save/regModels_africa.RData') # Load the models again if needed
 msummary(regModels, stars = T) # Rough regression table with all variables and stars
 
 ## gof generation -----------------------------------------------------------------------------------------------------------------------------------------

### 2025-08-22 manuscript/manuscript.qmd
Fix manuscript with updated tinytable commands

--- 2025-08-21 manuscript/manuscript.qmd
@@
     orcid: 0009-0004-4652-3429
-    corresponding: true
+    correspondence: true
   - name: Tong Hu

### 2025-08-22 code/PCA.R
BREAKING: Use Renv (#54)
and tinytable 0.13.0.

* init renv

* PCA.R: use library only for renv

* PCA: reformat

* update lockfile with PCA.R

* is_europe: reformat + use library()

* process is_africa

* process continents

* update main table output

* PCA.R: add data loading comment

* is_europe: fix reg loading comment

* update regTable_europe

* is_Africa: fix reg loading comment

* update regTable_africa

* Fix manuscript with updated tinytable commands

* update README to erxplain renv

@@ -13,20 +13,15 @@
 # SETUP ----------------------------------------------------------------------------------------------------------------
 rm(list = ls()) # Clear workspace
 
-need <- c(
-  'tidyverse',
-  'haven',
-  'AER',
-  'fixest',
-  "kableExtra",
-  'modelsummary',
-  'corrplot',
-  "ggfortify",
-  "parameters"
-) # list packages needed
-have <- need %in% rownames(installed.packages()) # checks packages you have
-if (any(!have)) install.packages(need[!have]) # install missing packages
-invisible(lapply(need, library, character.only = T))
+library('tidyverse')
+library('haven')
+library('AER')
+library('fixest')
+library("kableExtra")
+library('modelsummary')
+library('corrplot')
+library("ggfortify")
+library("parameters")
 
 data <- read_rds("data/panel_data.rds") # Load data
 
@@ -107,9 +102,8 @@ save(US, EU, file = "save/PCA.RData") # Save PCA results
     y = "Percentage of explained variances"
   ) +
   scale_x_continuous(breaks = 1:3, limits = c(1, 3.1)) +
-  scale_y_continuous(labels = scales::percent,
-                     limits = c(0, 0.55)) +
-  theme_bw()+
+  scale_y_continuous(labels = scales::percent, limits = c(0, 0.55)) +
+  theme_bw() +
   theme(panel.grid.minor.x = element_blank()))
 
 (scree_EU <- data.frame(
@@ -130,9 +124,8 @@ save(US, EU, file = "save/PCA.RData") # Save PCA results
     y = "Percentage of explained variances"
   ) +
   scale_x_continuous(breaks = 1:3, limits = c(1, 3.1)) +
-  scale_y_continuous(labels = scales::percent,
-                     limits = c(0, 0.65)) +
-  theme_bw()+
+  scale_y_continuous(labels = scales::percent, limits = c(0, 0.65)) +
+  theme_bw() +
   theme(panel.grid.minor.x = element_blank()))
 
 save(scree_US, scree_EU, file = "save/scree.RData") # Save scree plot results
@@ -144,17 +137,20 @@ data$eu <- predict(EU, data)[, 1] |> scale()
 data$pca <- -predict(PCA, data)[, 1]
 
 write_rds(data, "data/panel_data_pca.rds")
+# data <- read_rds("data/panel_data_pca.rds") # Reload data with PCA variables
 
[…]

### 2025-08-22 code/by_continent.R

@@ -8,19 +8,12 @@
 # SETUP ----------------------------------------------------------------------------------------------------------------
 rm(list = ls()) # Clear workspace
 
-need <- c(
-  'tidyverse',
-  'AER',
-  'fixest',
-  "kableExtra",
-  'modelsummary',
-  "parameters"
-) # list packages needed
-have <- need %in% rownames(installed.packages()) # checks packages you have
-if (any(!have)) {
-  install.packages(need[!have])
-} # install missing packages
-invisible(lapply(need, library, character.only = T))
+library('tidyverse')
+library('AER')
+library('fixest')
+library("kableExtra")
+library('modelsummary')
+library("parameters")
 
 data <- read_rds("data/panel_data_pca.rds") # Load data
 
@@ -126,28 +119,31 @@ get_continent_estimate <- function(continent, power, mod) {
 ## IMF loan to GDP ratio --------------------------------------------------------------------------------------------------
 loan_summary_df <- get_estimates(loan) |>
   dplyr::select(-term) |>
-  slice(0) |>  # empties the dataframe while preserving structure
+  slice(0) |> # empties the dataframe while preserving structure
   mutate(
     continent = character(),
     power = character()
   ) |>
   relocate(continent, power, .before = 1)
 
-  # Loop through each continent and power to compute estimates
+# Loop through each continent and power to compute estimates
 for (ct in continents) {
   for (pw in c("us", "eu")) {
     row <- get_continent_estimate(ct, pw, loan) |>
       mutate(continent = ct, power = pw) |>
       relocate(continent, power, .before = 1)
-    
+
     loan_summary_df <- bind_rows(loan_summary_df, row)
   }
 }
 
 ggplot(loan_summary_df, aes(x = power, y = estimate, color = power)) +
   geom_point(position = position_dodge(width = 0.5)) +
-  geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
-                width = 0.2, position = position_dodge(width = 0.5)) +
+  geom_errorbar(
+    aes(ymin = conf.low, ymax = conf.high),
+    width = 0.2,
+    position = position_dodge(width = 0.5)
+  ) +
   facet_wrap(~continent, ncol = 3) +
   theme_bw() +
   labs(
@@ -160,7 +156,7 @@ ggplot(loan_summary_df, aes(x = power, y = estimate, color = power)) +
[…]

### 2025-08-22 code/continent.R

@@ -9,13 +9,8 @@
 # SETUP ----------------------------------------------------------------------------------------------------------------
 rm(list = ls()) # Clear workspace
 
-need <- c(
-  'tidyverse',
-  'jsonlite'
-) # list packages needed
-have <- need %in% rownames(installed.packages()) # checks packages you have
-if (any(!have)) install.packages(need[!have]) # install missing packages
-invisible(lapply(need, library, character.only = T))
+library('tidyverse')
+library('jsonlite')
 
 # Read data ----------------------------------------------------------------------------------------------------------------
 raw_data <- read_dta("data/panel_data_raw.dta")

### 2025-08-22 code/is_Africa.R

@@ -9,17 +9,12 @@
 # SETUP ----------------------------------------------------------------------------------------------------------------
 rm(list = ls()) # Clear workspace
 
-need <- c(
-  'tidyverse',
-  'AER',
-  'fixest',
-  "kableExtra",
-  'modelsummary',
-  "parameters"
-) # list packages needed
-have <- need %in% rownames(installed.packages()) # checks packages you have
-if (any(!have)) install.packages(need[!have]) # install missing packages
-invisible(lapply(need, library, character.only = T))
+library('tidyverse')
+library('AER')
+library('fixest')
+library("kableExtra")
+library('modelsummary')
+library("parameters")
 
 data <- read_rds("data/panel_data_pca.rds") # Load data
 
@@ -30,11 +25,22 @@ data$is_african <- ifelse(data$continent == "Africa", 1, 0)
 
 ## Prepare variables -----------------------------------------------------------------------------------------------------
 rhs <- c(
-  "us", "eu*is_african", "shstaffl", "shquotal", 
-  "lnrgdpnew", "lnrgdpnewsq", 
-  "rgdpchnew", "rgdpchnewsquare", 
-  "growth1new", "reserv1", "oecd1", 
-  "year1980", "year1985", "year1990", "year1995", "year2000"
+  "us",
+  "eu*is_african",
+  "shstaffl",
+  "shquotal",
+  "lnrgdpnew",
+  "lnrgdpnewsq",
+  "rgdpchnew",
+  "rgdpchnewsquare",
+  "growth1new",
+  "reserv1",
+  "oecd1",
+  "year1980",
+  "year1985",
+  "year1990",
+  "year1995",
+  "year2000"
 )
 make_formula <- function(outcome) {
   as.formula(paste(outcome, paste(rhs, collapse = " + "), sep = " ~ "))
@@ -91,7 +97,7 @@ regModels <- list(
 )
 save(regModels, file = 'save/regModels_africa.RData')
 
-# load('save/regModels.RData') # Load the models again if needed
+# load('save/regModels_africa.RData') # Load the models again if needed
 msummary(regModels, stars = T) # Rough regression table with all variables and stars
 
 ## gof generation -----------------------------------------------------------------------------------------------------------------------------------------

### 2025-08-22 code/is_Europe.R

@@ -9,17 +9,13 @@
 # SETUP ----------------------------------------------------------------------------------------------------------------
 rm(list = ls()) # Clear workspace
 
-need <- c(
-  'tidyverse',
-  'AER',
-  'fixest',
-  "kableExtra",
-  'modelsummary',
-  "parameters"
-) # list packages needed
-have <- need %in% rownames(installed.packages()) # checks packages you have
-if (any(!have)) install.packages(need[!have]) # install missing packages
-invisible(lapply(need, library, character.only = T))
+
+library('tidyverse')
+library('AER')
+library('fixest')
+library("kableExtra")
+library('modelsummary')
+library("parameters")
 
 data <- read_rds("data/panel_data_pca.rds") # Load data
 
@@ -30,11 +26,22 @@ data$is_european <- ifelse(data$continent == "Europe", 1, 0)
 
 ## Prepare variables -----------------------------------------------------------------------------------------------------
 rhs <- c(
-  "us", "eu*is_european", "shstaffl", "shquotal", 
-  "lnrgdpnew", "lnrgdpnewsq", 
-  "rgdpchnew", "rgdpchnewsquare", 
-  "growth1new", "reserv1", "oecd1", 
-  "year1980", "year1985", "year1990", "year1995", "year2000"
+  "us",
+  "eu*is_european",
+  "shstaffl",
+  "shquotal",
+  "lnrgdpnew",
+  "lnrgdpnewsq",
+  "rgdpchnew",
+  "rgdpchnewsquare",
+  "growth1new",
+  "reserv1",
+  "oecd1",
+  "year1980",
+  "year1985",
+  "year1990",
+  "year1995",
+  "year2000"
 )
 make_formula <- function(outcome) {
   as.formula(paste(outcome, paste(rhs, collapse = " + "), sep = " ~ "))
@@ -91,7 +98,7 @@ regModels <- list(
 )
 save(regModels, file = 'save/regModels_europe.RData')
 
-# load('save/regModels.RData') # Load the models again if needed
+# load('save/regModels_europe.RData') # Load the models again if needed
 msummary(regModels, stars = T) # Rough regression table with all variables and stars
 
 ## gof generation -----------------------------------------------------------------------------------------------------------------------------------------

### 2025-08-28 code/fiscal.R
Add fiscal.R

@@ -1,0 +1,184 @@
+######## INFO ########
+# Authors: Dianyi Yang
+# R Script
+# Purpose: This script runs the robustness check with required fiscal adjustments (RFA) as the dependent variable for conditionality.
+# Inputs:  data/fiscal.dta
+#          data/panel_data_pca.rds
+#          save/regModels.RData
+#          helper/f_test_fiscal.R
+# Outputs:
+#          data/panel_data_fiscal.rds
+#          save/regModels_fiscal.RData
+#          save/regTable_fiscal.RData
+
+# SETUP ----------------------------------------------------------------------------------------------------------------
+rm(list = ls()) # Clear workspace
+
+library('tidyverse')
+library('AER')
+library('fixest')
+library("kableExtra")
+library('modelsummary')
+library("parameters")
+
+# Clean data -----------------------------------------------------------------------------------------------------------
+fiscal <- read_dta("data/fiscal.dta") # Load rfa data
+data <- read_rds("data/panel_data_pca.rds") # Load data
+
+## Check name consistency -----------------------------------------------------------------------------------------------------
+fiscal_names <- fiscal |>
+  distinct(countryn) |>
+  pull() |>
+  str_to_lower() |>
+  sort()
+data_names <- data |>
+  distinct(country) |>
+  pull() |>
+  str_to_lower() |>
+  sort()
+fiscal_names[!fiscal_names %in% data_names] # only cote d'ivoire needs to be fixed
+fiscal <- fiscal |>
+  mutate(countryn = ifelse(countryn == "COTE D'IVOIRE", "COTE D`IVOIRE", countryn))
+
+## summarise to 5 year intervals -----------------------------------------------------------------------------------------------------
+fiscal_sum <- fiscal |>
+  mutate(panel = case_when(
+    approval >= 1980 & approval < 1985 ~ 1980,
+    approval >= 1985 & approval < 1990 ~ 1985,
+    approval >= 1990 & approval < 1995 ~ 1990,
+    approval >= 1995 & approval < 2000 ~ 1995,
+    approval >= 2000 & approval < 2005 ~ 2000,
+    approval >= 2005 & approval < 2010 ~ 2005,
+    TRUE ~ NA_real_
+  )) |>
+  group_by(countryn, panel) |>
+  summarise(rfa1 = mean(rfa1, na.rm = TRUE),
+            rfa2 = mean(rfa2, na.rm = TRUE),
+            numb_fiscal = sum(numb_fiscal, na.rm = TRUE)) |>
+  ungroup() |> 
[…]

### 2025-08-28 code/fiscal.R
T+1 in maths mode

@@ -130,7 +130,7 @@ msummary(rfa2, stars = TRUE)
 regModels <- list(
   'Number of IMF conditions' = condition,
   'Number of Fiscal conditions' = fiscal_condition,
-  'RFA (T+1)' = rfa1,
+  'RFA ($T$+1)' = rfa1,
   'RFA (end year)' = rfa2
 )
 save(regModels, file = 'save/regModels_fiscal.RData')

### 2025-08-29 code/fiscal.R
Add line to reload data

@@ -68,6 +68,7 @@ data_new <- data |>
   select(-country_lower)
 
 saveRDS(data_new, "data/panel_data_fiscal.rds") # save the new data
+# data_new <- read_rds("data/panel_data_fiscal.rds") # reload to avoid any potential issues
 
 # Run regressions -----------------------------------------------------------------------------------------------------------
 ## Load old condition regression model ---------------------------------------------------------------------------------

### 2025-08-29 manuscript/manuscript.qmd
Add fiscal table to appendix
Add 1st para to robustness checks

--- 2025-08-21 manuscript/manuscript.qmd
@@
 
+## Alternative Measure of Conditionality
+
+Another concern is that our results on conditionality could be sensitive to the way IMF programme conditions are measured.
+Our baseline analysis follows @lipscy2018, who aggregate the total number of conditions across five-year periods.
+Yet, as @guimaraes2021 note, the overall count of IMF conditions, although widely used, may not fully capture the *stringency* of conditionality.
+They therefore propose alternative indicators, including the number of *fiscal* conditions and the required fiscal adjustment (RFA), to reassess the relationship between political alignment and conditionality.
+To address this concern, we replicate our conditionality model (the fourth specification in @tbl-results) using these alternative measures as dependent variables, and present the results in Appendix [-@sec-fiscal].
+
 # Case Study
@@
 ```
+
+# Fiscal Conditions {#sec-fiscal .appendix}
+
+```{r}
+#| label: tbl-results-fiscal
+#| tbl-cap: Effects of US and European Influence on IMF fiscal conditionality (1999–2012)
+#| tbl-pos: H
+load('save/regTable_fiscal.RData')
+resultsTable |> tinytable::theme_latex(resize_width = 1, resize_direction = "down")
+```

### 2025-09-01 manuscript/manuscript.qmd
Fix appendix table counter

@@
           \AtAppendix{\counterwithin{figure}{section}}
+          \AtAppendix{\counterwithin{table}{section}}
 abstract: "This paper investigates whether the United States or Europe holds greater influence over the International Monetary Fund (IMF)—a longstanding yet underexamined question in the study of global economic governance.

### 2025-09-01 manuscript/manuscript.qmd
Add 2nd para

@@
 
+The results, presented in @tbl-results-fiscal, suggest that our finding that closer relations with the US, rather than European powers, are associated with fewer IMF conditions also applies to the *fiscal* subset of conditions.
+This also echoes @guimaraes2021, who note that political alignment with the US reduces the number of fiscal conditions, although they do not control for European influence.
+Nevertheless, a caveat is that these findings should not be interpreted as evidence that US influence reduces the *stringency* of fiscal conditionality, as our results on the required fiscal adjustment (RFA) are inconclusive, again similar to the results reported by @guimaraes2021.
+Also, one should note that the data for fiscal conditions are available only from 1999 onwards.
+When merged to our data, the number of observations is extremely limited ($N \approx 30$) and should not be over-interpreted.
+
 # Case Study

### 2025-09-01 manuscript/manuscript.qmd
2nd para rephrased

--- 2025-09-01 manuscript/manuscript.qmd
@@
 
+The results in @tbl-results-fiscal indicate that our earlier finding, that closer relations with the US, rather than European powers, are associated with fewer IMF conditions, extends to the *fiscal* subset of conditions.
+This echoes @guimaraes2021, who also find that political alignment with the US reduces fiscal conditions, though they do not control for European influence.
+Importantly, fewer fiscal conditions should not be taken as evidence of less stringent conditionality: our results on the required fiscal adjustment (RFA) are inconclusive, consistent with @guimaraes2021.
+Finally, fiscal-conditions data are available only from 1999 onwards, leaving a very limited sample ($N \approx 30$), so these findings should be interpreted with caution.
+
 # Case Study

### 2025-09-02 manuscript/manuscript.qmd
Appendix: add 1st para

@@
 
+This appendix extends our analysis of IMF conditionality by examining alternative measures that focus specifically on fiscal policy requirements.
+Following @guimaraes2021, we consider not only the total number of conditions (as in our main analysis), but also the number of *fiscal* conditions and the required fiscal adjustment (RFA)—the extent of fiscal tightening mandated by IMF programmes, expressed as a percentage of GDP.
+We report RFA at two time horizons: one year after programme approval ($T+1$) and at the programme’s end year.
+Together, these measures provide a more nuanced understanding of the nature and stringency of IMF conditionality, particularly in the fiscal domain.
+
 ```{r}

### 2025-09-02 manuscript/manuscript.qmd
Appendix: 2nd para

@@
 
+@tbl-results-fiscal reveals patterns that largely reinforce our main findings while adding important nuance.
+Consistent with our baseline results, US influence is significantly associated with fewer overall IMF conditions.
+This pattern extends to fiscal conditions, where US influence is associated with a substantial reduction, though this effect is only marginally significant ($p < 0.1$), likely due to the much smaller sample size ($N=31$ vs.
+$N=93$ for total conditions).
+Interestingly, European influence shows the opposite pattern: it is associated with a significant increase in fiscal conditions ($p < 0.05$), suggesting that European powers may advocate for more extensive fiscal reforms when they have influence over IMF lending decisions.
+This finding supports arguments in the current literature that Europe, if anything, has more fully internalized neoliberal economic orthodoxy and the principles of the "Washington Consensus" than the United States, and may view austerity-oriented policies as intrinsically beneficial [@lutz2014; @dyson2010].
+
 ```{r}

### 2025-09-02 code/fiscal.R
fiscal: add haven package for read_dta

@@ -20,6 +20,7 @@ library('fixest')
 library("kableExtra")
 library('modelsummary')
 library("parameters")
+library("haven")
 
 # Clean data -----------------------------------------------------------------------------------------------------------
 fiscal <- read_dta("data/fiscal.dta") # Load rfa data

### 2025-09-02 manuscript/manuscript.qmd
appendix: my suggestion for para 3 and 4

@@
 
+The required fiscal adjustment (RFA), however, shows inconsistent results.
+US influence correlates with significantly more stringent fiscal adjustment requirements in the short term ($p<0.05$ for T+1), while European influence is associated with more lenient short-term adjustment requirements ($p<0.1$).
+This is at odds with the findings on the number of (fiscal) conditions.
+Moreover, such patterns do not persist in the longer term (end year), where neither US nor European influence shows statistically significant effects.
+These mixed results suggest that while political alignment may influence the *number* of fiscal conditions imposed, it does not consistently affect the *stringency* of fiscal adjustment requirements over time.
+This echoes the findings of @guimaraes2021 and indicates that the relationship between political influence and the nature of conditionality is complex and warrants further investigation.
+
+We note that these fiscal conditionality analyses are based on substantially smaller samples ($N\approxeq30$) and cover a more limited time period (1999-2012) than our main analyses.
+This reduction in sample size is due to the limited availability of disaggregated fiscal conditionality data.
+Consequently, these results should be interpreted with appropriate caution.
+
 ```{r}

### 2025-09-02 manuscript/manuscript.qmd
Appendix: revise para 3 and 4

@@
 The required fiscal adjustment (RFA), however, shows inconsistent results.
-US influence correlates with significantly more stringent fiscal adjustment requirements in the short term ($p<0.05$ for T+1), while European influence is associated with more lenient short-term adjustment requirements ($p<0.1$).
-This is at odds with the findings on the number of (fiscal) conditions.
-Moreover, such patterns do not persist in the longer term (end year), where neither US nor European influence shows statistically significant effects.
+US influence correlates with significantly more stringent fiscal adjustment requirements in the short term ($p<0.05$ for $T+1$), while European influence is associated with more lenient short-term adjustment requirements ($p<0.1$).
+This stands in contrast to the findings on the number of (fiscal) conditions.
+Moreover, these patterns do not persist in the longer term (end year), where neither US nor European influence shows statistically significant effects.
 These mixed results suggest that while political alignment may influence the *number* of fiscal conditions imposed, it does not consistently affect the *stringency* of fiscal adjustment requirements over time.
@@
 
-We note that these fiscal conditionality analyses are based on substantially smaller samples ($N\approxeq30$) and cover a more limited time period (1999-2012) than our main analyses.
+We note that these fiscal conditionality analyses are based on substantially smaller samples ($N\approx 30$) and cover a more limited time period (1999–2012) than our main analyses.
 This reduction in sample size is due to the limited availability of disaggregated fiscal conditionality data.

### 2025-09-02 code/fiscal.R
Fiscal Appendix: fix note in table

@@ -174,7 +174,7 @@ getTable <- function(output = "html") {
     escape = F,
     output = output,
     notes = c(
-      "Standard errors clustered at the country level. Values in square brackets represent $p$-values from $F$-tests (linear models) and Chi-squared tests (Probit model)."
+      "Standard errors clustered at the country level. Values in square brackets represent $p$-values from $F$-tests."
     )
   )
 }

### 2025-09-02 manuscript/manuscript.qmd
Appendix: change section title
Fiscal Conditionality Analysis (#56)
* Add fiscal data

* Add cleaned data

* Add fiscal.R

* adapt f_test helper for fiscal

* save reg models

* save reg table

* T+1 in maths mode

* Add fiscal table to appendix

* Add line to reload data

* Add 1st para to robustness checks

* Fix appendix table counter

* Add 2nd para

* 2nd para rephrased

* Appendix: add 1st para

* Appendix: 2nd para

* fiscal: add haven package for read_dta

* appendix: my suggestion for para 3 and 4

* Appendix: revise para 3 and 4

* Appendix: change section title

* Add footnote to ex

@@
 
-# Fiscal Conditions {#sec-fiscal .appendix}
+# Fiscal Conditionality Analysis {#sec-fiscal .appendix}
 
 This appendix extends our analysis of IMF conditionality by examining alternative measures that focus specifically on fiscal policy requirements.
-Following @guimaraes2021, we consider not only the total number of conditions (as in our main analysis), but also the number of *fiscal* conditions and the required fiscal adjustment (RFA)—the extent of fiscal tightening mandated by IMF programmes, expressed as a percentage of GDP.
+Following @guimaraes2021, we consider not only the total number of conditions (as in our main analysis), but also the number of *fiscal* conditions and the required fiscal adjustment (RFA)[^13]—the extent of fiscal tightening mandated by IMF programmes, expressed as a percentage of GDP.
 We report RFA at two time horizons: one year after programme approval ($T+1$) and at the programme’s end year.
 Together, these measures provide a more nuanced understanding of the nature and stringency of IMF conditionality, particularly in the fiscal domain.
+
+[^13]: To merge these variables with our dataset, aggregation is necessary: our data are at the five-year country level, whereas theirs are at the programme level.
+We sum the number of fiscal conditions and average the RFA variables.
 
@@
 Interestingly, European influence shows the opposite pattern: it is associated with a significant increase in fiscal conditions ($p < 0.05$), suggesting that European powers may advocate for more extensive fiscal reforms when they have influence over IMF lending decisions.
-This finding supports arguments in the current literature that Europe, if anything, has more fully internalized neoliberal economic orthodoxy and the principles of the "Washington Consensus" than the United States, and may view austerity-oriented policies as intrinsically beneficial [@lutz2014; @dyson2010].
+This finding supports arguments in the current literature that Europe, if anything, has more fully internalised neoliberal economic orthodoxy and the principles of the "Washington Consensus" than the United States, and may view austerity-oriented policies as intrinsically beneficial [@lutz2014; @dyson2010].
 

### 2025-09-02 code/fiscal.R

@@ -1,0 +1,186 @@
+######## INFO ########
+# Authors: Dianyi Yang
+# R Script
+# Purpose: This script runs the robustness check with required fiscal adjustments (RFA) as the dependent variable for conditionality.
+# Inputs:  data/fiscal.dta
+#          data/panel_data_pca.rds
+#          save/regModels.RData
+#          helper/f_test_fiscal.R
+# Outputs:
+#          data/panel_data_fiscal.rds
+#          save/regModels_fiscal.RData
+#          save/regTable_fiscal.RData
+
+# SETUP ----------------------------------------------------------------------------------------------------------------
+rm(list = ls()) # Clear workspace
+
+library('tidyverse')
+library('AER')
+library('fixest')
+library("kableExtra")
+library('modelsummary')
+library("parameters")
+library("haven")
+
+# Clean data -----------------------------------------------------------------------------------------------------------
+fiscal <- read_dta("data/fiscal.dta") # Load rfa data
+data <- read_rds("data/panel_data_pca.rds") # Load data
+
+## Check name consistency -----------------------------------------------------------------------------------------------------
+fiscal_names <- fiscal |>
+  distinct(countryn) |>
+  pull() |>
+  str_to_lower() |>
+  sort()
+data_names <- data |>
+  distinct(country) |>
+  pull() |>
+  str_to_lower() |>
+  sort()
+fiscal_names[!fiscal_names %in% data_names] # only cote d'ivoire needs to be fixed
+fiscal <- fiscal |>
+  mutate(countryn = ifelse(countryn == "COTE D'IVOIRE", "COTE D`IVOIRE", countryn))
+
+## summarise to 5 year intervals -----------------------------------------------------------------------------------------------------
+fiscal_sum <- fiscal |>
+  mutate(panel = case_when(
+    approval >= 1980 & approval < 1985 ~ 1980,
+    approval >= 1985 & approval < 1990 ~ 1985,
+    approval >= 1990 & approval < 1995 ~ 1990,
+    approval >= 1995 & approval < 2000 ~ 1995,
+    approval >= 2000 & approval < 2005 ~ 2000,
+    approval >= 2005 & approval < 2010 ~ 2005,
+    TRUE ~ NA_real_
+  )) |>
+  group_by(countryn, panel) |>
+  summarise(rfa1 = mean(rfa1, na.rm = TRUE),
+            rfa2 = mean(rfa2, na.rm = TRUE),
+            numb_fiscal = sum(numb_fiscal, na.rm = TRUE)) |>
[…]

### 2025-09-06 manuscript/manuscript.qmd
Add summary stats as a section

@@
 
-## Model Specification {#sec-model}
+## Summary Statistics
+
+We present summary statistics for key variables in @tbl-summary.
+
+```{r}
+#| label: tbl-summary
+#| tbl-cap: Summary Statistics of Key Variables (1980–2010)
+#| tbl-pos: H
+library(modelsummary)
+dsummary(
+  (`IMF loan to GDP ratio` = imfloannew100) +
+    (`IMF participation rate` = imf_p) +
+    (`IMF loan approval` = imf5a) +
+    (`Number of IMF conditions` = tc) +
+    (`USA Influence` = us) +
+    (`EUP Influence` = eu) +
+    (`IMF Staff` = shstaffl) +
+    (`IMF Quota` = `shquotal`) +
+    (GDP = lnrgdpnew) +
+    (GDPpc = rgdpchnew) +
+    (`GDPpc growth` = growth1new) +
+    (Reserves = reserv1) +
+    (OECD = oecd1) ~ 
+    Mean + SD + Min + (Q1 = P25) + Median + (Q3 = P75) + Max + N,
+  data = dta
+) |> tinytable::theme_latex(resize_width = 1, resize_direction = "down")
+```
+
+# Model Specification {#sec-model}
 

### 2025-09-06 manuscript/manuscript.qmd
use data with pca var
More description

@@
 
-# Empirical Strategy
-
-## Hypotheses
+# Hypotheses
 
@@
 
-## Data
+# Data
 
@@
 #| label: fig-data
-dta <- readr::read_rds('data/panel_data.rds')
+dta <- readr::read_rds('data/panel_data_pca.rds')
 ```
@@
 
-We present summary statistics for key variables in @tbl-summary.
+@tbl-summary presents summary statistics for the key variables.
+IMF lending outcomes are generally sparse: both the loan-to-GDP ratio and the participation rate have means close to zero, with medians at zero, indicating that most country-periods did not involve IMF programmes.
+However, when loans are approved, they can be large, with the loan-to-GDP ratio reaching over 12 percent and the number of conditions ranging up to 179.
+The influence measures derived from PCA are centered at zero by construction, with wide dispersion across countries.
+Among the covariates, GDP per capita and its growth show substantial variation, while OECD membership is rare in the sample (16 percent of observations).
+International reserves also vary widely, suggesting heterogeneous external liquidity and self-insurance capacity across countries.
+These patterns underscore the skewed and censored nature of IMF outcomes, motivating our use of Tobit and Probit models in the subsequent analysis.
 

### 2025-09-07 manuscript/manuscript.qmd
Add units

@@
 dsummary(
-  (`IMF loan to GDP ratio` = imfloannew100) +
+  (`IMF loan to GDP ratio (%)` = imfloannew100) +
     (`IMF participation rate` = imf_p) +
@@
     (`EUP Influence` = eu) +
-    (`IMF Staff` = shstaffl) +
-    (`IMF Quota` = `shquotal`) +
-    (GDP = lnrgdpnew) +
-    (GDPpc = rgdpchnew) +
-    (`GDPpc growth` = growth1new) +
-    (Reserves = reserv1) +
+    (`IMF Staff (%)` = shstaffl) +
+    (`IMF Quota (%)` = `shquotal`) +
+    (`GDP (log)` = lnrgdpnew) +
+    (`GDPpc (log)` = rgdpchnew) +
+    (`GDPpc growth (%)` = growth1new) +
+    (`Reserves (months of imports)` = reserv1) +
     (OECD = oecd1) ~ 

### 2025-09-07 manuscript/manuscript.qmd
update corresponding description in text

@@
 We include a range of economic performance indicators that have been shown to systematically affect IMF lending decisions [@barro2005; @knight1997].
-To capture these dynamics, we include GDP, per capita GDP, and lagged GDP growth as proxies for macroeconomic performance.
-International reserves as a share of imports are used to measure a country’s external liquidity and ability to manage balance-of-payments pressures without IMF support.
+To capture these dynamics, we include log GDP, log per capita GDP, and lagged GDP growth as proxies for macroeconomic performance.
+International reserves measured in months of imports are used to measure a country’s external liquidity and ability to manage balance-of-payments pressures without IMF support.
 Finally, a binary variable for OECD membership serves as a proxy for broader institutional and economic differences among countries.

### 2025-09-07 manuscript/manuscript.qmd
Add logged share explanation

@@
 Notably, countries with larger IMF quotas (and thus greater voting power), as well as those with nationals among IMF staff, are expected to receive more favourable treatment from the Fund [@barro2005].
+Therefore, we include the share of nationals among IMF economists and the share of IMF quota, both in logged form, as control covariates.
 
@@
     (`EUP Influence` = eu) +
-    (`IMF Staff (%)` = shstaffl) +
-    (`IMF Quota (%)` = `shquotal`) +
+    (`IMF Staff (log)` = shstaffl) +
+    (`IMF Quota (log)` = `shquotal`) +
     (`GDP (log)` = lnrgdpnew) +

### 2025-09-07 manuscript/manuscript.qmd
Add summary statistics (#58)
* Add summary stats as a section

* use data with pca var

* More description

* Add units

* update corresponding description in text

* Add logged share explanation

* correct GDP pc measurement

@@
 We include a range of economic performance indicators that have been shown to systematically affect IMF lending decisions [@barro2005; @knight1997].
-To capture these dynamics, we include log GDP, log per capita GDP, and lagged GDP growth as proxies for macroeconomic performance.
+To capture these dynamics, we include log GDP, per capita GDP, and lagged GDP growth as proxies for macroeconomic performance[^9].
 International reserves measured in months of imports are used to measure a country’s external liquidity and ability to manage balance-of-payments pressures without IMF support.
@@
 
-The selection of covariates, including their polynomial transformations[^9], closely follows @lipscy2018 and @barro2005 to ensure comparability with previous findings.
+[^9]: GDP, before log transformation, is measured in millions of constant 2002 US dollars, while GDP per capita is measured in thousands of constant 2002 US dollars.
 
-[^9]: Quadratic terms are included for GDP and per capita GDP to capture potential non-linear effects. See [Model Specification](#sec-model) for details.
+The selection of covariates, including their polynomial transformations[^10], closely follows @lipscy2018 and @barro2005 to ensure comparability with previous findings.
+
+[^10]: Quadratic terms are included for GDP and per capita GDP to capture potential non-linear effects. See [Model Specification](#sec-model) for details.
 
@@
     (`GDP (log)` = lnrgdpnew) +
-    (`GDPpc (log)` = rgdpchnew) +
+    (`GDPpc (thousands USD)` = rgdpchnew) +
     (`GDPpc growth (%)` = growth1new) +
@@
 This appendix extends our analysis of IMF conditionality by examining alternative measures that focus specifically on fiscal policy requirements.
-Following @guimaraes2021, we consider not only the total number of conditions (as in our main analysis), but also the number of *fiscal* conditions and the required fiscal adjustment (RFA)[^13]—the extent of fiscal tightening mandated by IMF programmes, expressed as a percentage of GDP.
+Following @guimaraes2021, we consider not only the total number of conditions (as in our main analysis), but also the number of *fiscal* conditions and the required fiscal adjustment (RFA)[^14]—the extent of fiscal tightening mandated by IMF programmes, expressed as a percentage of GDP.
 We report RFA at two time horizons: one year after programme approval ($T+1$) and at the programme’s end year.
@@
 
-[^13]: To merge these variables with our dataset, aggregation is necessary: our data are at the five-year country level, whereas theirs are at the programme level.
+[^14]: To merge these variables with our dataset, aggregation is necessary: our data are at the five-year country level, whereas theirs are at the programme level.
 We sum the number of fiscal conditions and average the RFA variables.

### 2025-09-08 manuscript/manuscript.qmd
add 2 literatures to the current version

@@
 Numerous case studies further demonstrate instances of US political pressure resulting in lenient loan conditions for key geopolitical allies, including Zaire, Egypt, Mexico, Argentina, and South Korea [@vreeland2009; @copelovitch2010a; @lipscy2018].
-This pattern of influence is consistent with the United States’ broader role across international financial institutions, particularly the World Bank, where American dominance is widely acknowledged [@wade2002; @woods2003; @clark2020].
+Closer allies of the United States enjoy IMF loans with less fiscal conditions[@dreher2003].
+This pattern of influence is consistent with the United States’ broader role across international financial institutions, particularly the World Bank, where American dominance is widely acknowledged \[@wade2002; @woods2003; @clark2020\].
 
@@
 This appendix extends our analysis of IMF conditionality by examining alternative measures that focus specifically on fiscal policy requirements.
+Fiscal conditionality, under a comprehensive review of both theoretical approaches and empirical evidence, is "the practice of giving financial assistance contingent on the implementation of specific policies", and it continues to rise over the years[@dreher2009a].
 Following @guimaraes2021, we consider not only the total number of conditions (as in our main analysis), but also the number of *fiscal* conditions and the required fiscal adjustment (RFA)[^14]—the extent of fiscal tightening mandated by IMF programmes, expressed as a percentage of GDP.

### 2025-09-09 manuscript/manuscript.qmd
Fix escaping

--- 2025-09-07 manuscript/manuscript.qmd
@@
 Numerous case studies further demonstrate instances of US political pressure resulting in lenient loan conditions for key geopolitical allies, including Zaire, Egypt, Mexico, Argentina, and South Korea [@vreeland2009; @copelovitch2010a; @lipscy2018].
+Closer allies of the United States enjoy IMF loans with less fiscal conditions[@dreher2003].
 This pattern of influence is consistent with the United States’ broader role across international financial institutions, particularly the World Bank, where American dominance is widely acknowledged [@wade2002; @woods2003; @clark2020].

### 2025-09-09 manuscript/manuscript.qmd
spaces before brackets

--- 2025-09-07 manuscript/manuscript.qmd
@@
 Numerous case studies further demonstrate instances of US political pressure resulting in lenient loan conditions for key geopolitical allies, including Zaire, Egypt, Mexico, Argentina, and South Korea [@vreeland2009; @copelovitch2010a; @lipscy2018].
+Closer allies of the United States enjoy IMF loans with less fiscal conditions [@dreher2003].
 This pattern of influence is consistent with the United States’ broader role across international financial institutions, particularly the World Bank, where American dominance is widely acknowledged [@wade2002; @woods2003; @clark2020].
@@
 This appendix extends our analysis of IMF conditionality by examining alternative measures that focus specifically on fiscal policy requirements.
+Fiscal conditionality, under a comprehensive review of both theoretical approaches and empirical evidence, is "the practice of giving financial assistance contingent on the implementation of specific policies", and it continues to rise over the years [@dreher2009a].
 Following @guimaraes2021, we consider not only the total number of conditions (as in our main analysis), but also the number of *fiscal* conditions and the required fiscal adjustment (RFA)[^14]—the extent of fiscal tightening mandated by IMF programmes, expressed as a percentage of GDP.

### 2025-09-09 manuscript/manuscript.qmd
seems linux being too case-sensitive is the reason

--- 2025-09-09 manuscript/manuscript.qmd
@@
   message: false
+  dir: project
 ---
@@
 Numerous case studies further demonstrate instances of US political pressure resulting in lenient loan conditions for key geopolitical allies, including Zaire, Egypt, Mexico, Argentina, and South Korea [@vreeland2009; @copelovitch2010a; @lipscy2018].
-Closer allies of the United States enjoy IMF loans with less fiscal conditions[@dreher2003].
 This pattern of influence is consistent with the United States’ broader role across international financial institutions, particularly the World Bank, where American dominance is widely acknowledged [@wade2002; @woods2003; @clark2020].
@@
 ```{r percentage_variance}
-load('save/PCA.Rdata')
+load('save/PCA.RData')
 US_PC1_percentage <- US$sdev[1]^2 / sum(US$sdev^2) * 100
@@
 This appendix extends our analysis of IMF conditionality by examining alternative measures that focus specifically on fiscal policy requirements.
-Fiscal conditionality, under a comprehensive review of both theoretical approaches and empirical evidence, is "the practice of giving financial assistance contingent on the implementation of specific policies", and it continues to rise over the years[@dreher2009a].
 Following @guimaraes2021, we consider not only the total number of conditions (as in our main analysis), but also the number of *fiscal* conditions and the required fiscal adjustment (RFA)[^14]—the extent of fiscal tightening mandated by IMF programmes, expressed as a percentage of GDP.

### 2025-09-10 manuscript/manuscript.qmd
Fix scree.RData file name in manuscript
revert the redundant `dir: project` in manuscript

@@
   message: false
-  dir: project
 ---
@@
 #| fig-pos: H
-load('save/scree.Rdata')
+load('save/scree.RData')
 scree_EU

### 2025-09-11 manuscript/manuscript.qmd
Improved keywords and added comprehensive conclusion section
Co-authored-by: kv9898 <105025148+kv9898@users.noreply.github.com>

@@
 In doing so, the paper contributes to debates on hegemony, institutional legitimacy, and intra-Western dynamics in global economic governance."
-keywords: IMF, US hegemony, Europe
+keywords: International Monetary Fund, transatlantic influence, global economic governance, conditionality, principal component analysis, institutional power, financial diplomacy
 execute:
@@
 
-# Case Study
+# Conclusion
 
-# Conclusion
+This paper has examined a fundamental yet underexplored question in the study of global economic governance: who wields greater influence over the International Monetary Fund—the United States or Europe?
+By disaggregating "Western" influence into its constituent components and employing principal component analysis to construct separate measures of US and European alignment, we have uncovered important asymmetries in how transatlantic powers exercise their influence within the Fund.
+
+Our empirical analysis yields several key findings that challenge conventional wisdom about IMF decision-making.
+First, European influence demonstrates a broader and more consistent association with favourable lending outcomes than previously recognised.
+Countries with stronger ties to major European powers—the United Kingdom, France, and Germany—receive significantly larger loans relative to GDP, participate more frequently in IMF programmes, and enjoy higher approval rates for new lending arrangements.
+These effects are statistically significant and substantively meaningful, suggesting that European actors wield considerable institutional influence over the Fund's core lending functions.
+
+Second, contrary to much existing literature that emphasises American dominance, US influence shows a more selective pattern of impact.
+While American ties do not significantly affect loan volumes, participation rates, or approval decisions, they are strongly associated with more lenient conditionality—fewer performance criteria, benchmarks, and prior actions attached to lending arrangements.
+This finding suggests that US power manifests not through broad institutional control over access to resources, but through targeted influence over the terms and conditions of assistance.
+
+Third, our robustness checks reveal that European influence operates on a global rather than regional or colonial basis.
+Neither European borrowers nor African recipients benefit disproportionately from their ties to European powers.
+If anything, the evidence suggests these groups may receive less favourable treatment than other countries with comparable alignment profiles.
+This pattern indicates that European influence reflects genuine institutional power within the IMF rather than geographic or historical favouritism.
+
+These findings have several important implications for our understanding of international financial institutions and global economic governance more broadly.
+Theoretically, they complicate simple narratives of American hegemony in international organisations.
+While the United States undoubtedly remains the most powerful single actor within the Bretton Woods system, our results suggest that European powers collectively exercise substantial influence over key institutional functions.
+This finding aligns with observations about the IMF's historical leadership structure—epitomised by the "gentlemen's agreement" reserving the Managing Director position for Europeans—and suggests that informal power arrangements may be more consequential than formal voting shares or veto rights.
+
+The asymmetric nature of US and European influence also speaks to broader debates about the sources and exercise of institutional power.
+European influence appears to operate through channels that affect aggregate resource allocation—determining who receives assistance and in what amounts.
+American influence, by contrast, manifests through mechanisms that shape the policy content of lending arrangements—affecting what borrowers must do to receive and maintain access to funds.
+This division of labour may reflect different institutional strengths, strategic priorities, or coordination mechanisms between the two actors.
+
+From a policy perspective, these findings have implications for ongoing debates about IMF reform and representation.
+Calls for enhanced emerging market voice and reduced Western dominance must grapple with the reality that "the West" is not a monolithic bloc.
+European and American interests may diverge in important ways, creating opportunities for coalition-building and institutional change that would not exist under unified Western control.
+Understanding these internal dynamics may be crucial for reform advocates seeking to reshape global financial governance.
+
+Our analysis also sheds light on the nature of conditionality, one of the most controversial aspects of IMF operations.
+The finding that American influence is associated with fewer conditions, while European influence shows no such effect, suggests that critiques of overly intrusive conditionality may need to be more precisely targeted.
+European powers, despite their historical association with social democratic governance models, do not appear to advocate for more lenient policy requirements.
+This pattern may reflect the influence of German-style fiscal conservatism within European policy circles or the adoption of neoliberal economic orthodoxy by European technocratic elites.
+
+Several limitations of our analysis should be acknowledged.
+First, our sample period ends in 2010, just as the European sovereign debt crisis was beginning to reshape transatlantic dynamics within the IMF.
+The Fund's involvement in European bailouts through the Troika arrangements may have altered the balance of influence in ways not captured by our data.
+Future research should examine whether the patterns we document persist in the post-crisis period.
+
+Second, our measures of influence, while grounded in the existing literature, capture alignment rather than direct lobbying or pressure.
+We cannot observe the specific mechanisms through which influence is exercised or the content of behind-the-scenes negotiations.
+Qualitative case studies examining specific lending decisions could provide valuable complementary insights into the micro-foundations of the patterns we document.
+
+Third, our focus on the three largest European powers—the UK, France, and Germany—may understate the collective influence of the European Union as a whole.
+As EU coordination mechanisms develop and potentially consolidate European representation within international financial institutions, the dynamics we analyse may continue to evolve.
+Research examining the role of EU institutional actors, such as the European Central Bank or European Commission, in IMF decision-making would provide a useful extension of our work.
+
+Fourth, our analysis focuses exclusively on lending outcomes and does not examine other dimensions of IMF operations, such as surveillance, technical assistance, or institutional governance.
+European and American influence may manifest differently across these various functions.
+Similarly, our study does not address the effectiveness or appropriateness of the lending arrangements influenced by great power politics—a question that would require additional data on economic outcomes and counterfactual analysis.
+
+Looking ahead, several promising avenues for future research emerge from our findings.
+First, examining the mechanisms through which European coordination occurs within the IMF would help illuminate the institutional foundations of the influence we document.
+Studies of EURIMF and SCIMF activities, European Central Bank participation in IMF decisions, or coordination within G7 and G20 forums could provide valuable insights into the operational aspects of European influence.
+
+Second, extending our analysis to include other potential great power influencers—particularly China, whose economic weight and IMF membership have grown substantially since our sample period—would help assess whether the patterns we document reflect specifically transatlantic dynamics or broader features of great power competition within international organisations.
+
+Third, examining variation in influence across different types of IMF programmes—such as Stand-By Arrangements, Extended Fund Facilities, or newer instruments like the Flexible Credit Line—could reveal whether European and American influence operates differently depending on the specific lending modality or the severity of the borrower's economic difficulties.
+
+Finally, our findings suggest the value of disaggregating "Western" influence in studies of other international organisations.
+The World Bank, regional development banks, and newer institutions like the Asian Infrastructure Investment Bank may exhibit similar asymmetries in great power influence that are masked by treating Western actors as a unified bloc.
+
+In conclusion, this paper demonstrates that the International Monetary Fund, despite its technocratic veneer, remains deeply shaped by geopolitical considerations.
+However, the nature of great power influence is more complex and differentiated than often assumed.
+European powers exercise broad institutional influence over access to IMF resources, while American power manifests more selectively through conditionality.
+This division of influence has important implications for understanding both the current operations of the international financial system and the prospects for its future reform.
+As emerging economies continue to challenge Western dominance in global governance, recognising the internal dynamics and potential fractures within the Western coalition becomes increasingly important for scholars, policymakers, and reform advocates alike.
 

### 2025-09-11 manuscript/manuscript.qmd
update keywords

@@
 In doing so, the paper contributes to debates on hegemony, institutional legitimacy, and intra-Western dynamics in global economic governance."
-keywords: International Monetary Fund, transatlantic influence, global economic governance, conditionality, principal component analysis, institutional power, financial diplomacy
+keywords: International Monetary Fund, hegemony, global economic governance, US influence, European influence, conditionality
 execute:

### 2025-09-11 manuscript/manuscript.qmd
fix 2nd para

@@
 First, European influence demonstrates a broader and more consistent association with favourable lending outcomes than previously recognised.
-Countries with stronger ties to major European powers—the United Kingdom, France, and Germany—receive significantly larger loans relative to GDP, participate more frequently in IMF programmes, and enjoy higher approval rates for new lending arrangements.
+Countries with stronger ties to major European powers，namely, the United Kingdom, France, and Germany, receive significantly larger loans relative to GDP, participate more frequently in IMF programmes, and enjoy higher approval rates for new lending arrangements.
 These effects are statistically significant and substantively meaningful, suggesting that European actors wield considerable institutional influence over the Fund's core lending functions.

### 2025-09-11 manuscript/manuscript.qmd
Major: fix 3rd para to avoid misinterpretation

@@
 Second, contrary to much existing literature that emphasises American dominance, US influence shows a more selective pattern of impact.
-While American ties do not significantly affect loan volumes, participation rates, or approval decisions, they are strongly associated with more lenient conditionality—fewer performance criteria, benchmarks, and prior actions attached to lending arrangements.
-This finding suggests that US power manifests not through broad institutional control over access to resources, but through targeted influence over the terms and conditions of assistance.
+While American ties do not significantly affect loan volumes, participation rates, or approval decisions, they are strongly associated with *fewer conditions*, that is, a reduced number of performance criteria, benchmarks, and prior actions attached to lending arrangements.
+This finding suggests that US power manifests not through broad institutional control over access to resources, but through targeted influence over the design of conditionality.
+However, this result should not be interpreted as evidence of systematically “more lenient” conditionality: our additional analyses find no evidence that American ties reduce the required fiscal adjustment imposed by IMF programmes.
 

### 2025-09-11 manuscript/manuscript.qmd
Major: fix 9th para to soften the tone

@@
 Several limitations of our analysis should be acknowledged.
-First, our sample period ends in 2010, just as the European sovereign debt crisis was beginning to reshape transatlantic dynamics within the IMF.
-The Fund's involvement in European bailouts through the Troika arrangements may have altered the balance of influence in ways not captured by our data.
-Future research should examine whether the patterns we document persist in the post-crisis period.
+First, our sample period ends in 2010, reflecting the coverage of the underlying datasets we rely on.
+While this prevents us from capturing the European sovereign debt crisis and the subsequent Troika arrangements, we expect that including the post-2010 period would, if anything, reinforce our findings.
+The IMF’s close collaboration with European institutions during the crisis suggests that European influence over lending decisions likely became even more pronounced after our period of observation.
 

### 2025-09-11 manuscript/manuscript.qmd
fix 10th para
remove the para on 3 european powers

@@
 Qualitative case studies examining specific lending decisions could provide valuable complementary insights into the micro-foundations of the patterns we document.
+However, such approaches often rely on extensive insider information and interviews with IMF officials, which are difficult to obtain given the Fund’s highly consensus-based and closed-door decision-making culture.
 
-Third, our focus on the three largest European powers—the UK, France, and Germany—may understate the collective influence of the European Union as a whole.
-As EU coordination mechanisms develop and potentially consolidate European representation within international financial institutions, the dynamics we analyse may continue to evolve.
-Research examining the role of EU institutional actors, such as the European Central Bank or European Commission, in IMF decision-making would provide a useful extension of our work.
-
-Fourth, our analysis focuses exclusively on lending outcomes and does not examine other dimensions of IMF operations, such as surveillance, technical assistance, or institutional governance.
+Third, our analysis focuses exclusively on lending outcomes and does not examine other dimensions of IMF operations, such as surveillance, technical assistance, or institutional governance.
 European and American influence may manifest differently across these various functions.

### 2025-09-12 manuscript/manuscript.qmd
Major: fix 3rd limitation

@@
 
-Third, our analysis focuses exclusively on lending outcomes and does not examine other dimensions of IMF operations, such as surveillance, technical assistance, or institutional governance.
-European and American influence may manifest differently across these various functions.
-Similarly, our study does not address the effectiveness or appropriateness of the lending arrangements influenced by great power politics—a question that would require additional data on economic outcomes and counterfactual analysis.
+Third, in terms of scope, three clarifications are in order.
+Our analysis focuses exclusively on lending outcomes and does not examine other dimensions of IMF operations, such as surveillance, technical assistance, or institutional governance.
+European and American influence may well manifest differently across these functions.
+Second, the effectiveness of IMF policies is also beyond the scope of our analysis.
+A large body of literature already addresses this question, and its findings can be combined with ours to inform policy debates.
+Third, our approach is deliberately empirical and positivist, aimed at identifying systematic patterns of influence rather than assessing the *appropriateness* of IMF lending arrangements.
+Questions of legitimacy and normative evaluation, long central to debates about the Fund, are not the direct focus of this paper, though we recognise that our results have implications for these debates and welcome further normative interpretations.
 

### 2025-09-12 manuscript/manuscript.qmd
future research: 1st

@@
 Looking ahead, several promising avenues for future research emerge from our findings.
-First, examining the mechanisms through which European coordination occurs within the IMF would help illuminate the institutional foundations of the influence we document.
-Studies of EURIMF and SCIMF activities, European Central Bank participation in IMF decisions, or coordination within G7 and G20 forums could provide valuable insights into the operational aspects of European influence.
+First, it is worth re-examining the determinants of IMF conditionality by moving beyond simple counts of conditions.
+In addition to the required fiscal adjustment, future work could measure conditionality across specific domains—such as trade liberalisation, privatisation, or structural reforms—to assess whether transatlantic influence shapes not only the quantity but also the substance of IMF policy requirements.
 

### 2025-09-12 manuscript/manuscript.qmd
future research: 2nd

@@
 
-Second, extending our analysis to include other potential great power influencers—particularly China, whose economic weight and IMF membership have grown substantially since our sample period—would help assess whether the patterns we document reflect specifically transatlantic dynamics or broader features of great power competition within international organisations.
+A second line of inquiry would be to examine the domestic sources of IMF influence.
+Our study measures international alignment, but the national actors transmitting influence remain a black box.
+Case studies focusing on finance ministries, central banks, foreign ministries, or private banking interests could shed light on which domestic constituencies mobilise influence within the IMF, and whether these channels differ between Europe and the United States.
 

### 2025-09-12 manuscript/manuscript.qmd
future research: 3rd

@@
 
-Third, examining variation in influence across different types of IMF programmes—such as Stand-By Arrangements, Extended Fund Facilities, or newer instruments like the Flexible Credit Line—could reveal whether European and American influence operates differently depending on the specific lending modality or the severity of the borrower's economic difficulties.
-
-Finally, our findings suggest the value of disaggregating "Western" influence in studies of other international organisations.
-The World Bank, regional development banks, and newer institutions like the Asian Infrastructure Investment Bank may exhibit similar asymmetries in great power influence that are masked by treating Western actors as a unified bloc.
+Finally, our findings suggest the value of applying a similar disaggregation of Western influence to other international organisations.
+Beyond the World Bank, promising candidates include the OECD, the Bank for International Settlements (BIS), the WTO, and the European Bank for Reconstruction and Development (EBRD).
+Each involves overlapping Western leadership but potentially divergent US and European priorities, offering a fertile testing ground for whether the asymmetries we document are specific to the IMF or reflect broader patterns of transatlantic governance.
 

### 2025-09-12 manuscript/manuscript.qmd
Better keywords and a draft for conclusion (#62)
* Initial plan

* Improved keywords and added comprehensive conclusion section

Co-authored-by: kv9898 <105025148+kv9898@users.noreply.github.com>

* update keywords

* fix 2nd para

* Major: fix 3rd para to avoid misinterpretation

* Major: fix 9th para to soften the tone

* fix 10th para

* remove the para on 3 european powers

* Major: fix 3rd limitation

* future research: 1st

* future research: 2nd

* future research: 3rd

* Fix closing

@@
 
-In conclusion, this paper demonstrates that the International Monetary Fund, despite its technocratic veneer, remains deeply shaped by geopolitical considerations.
-However, the nature of great power influence is more complex and differentiated than often assumed.
-European powers exercise broad institutional influence over access to IMF resources, while American power manifests more selectively through conditionality.
-This division of influence has important implications for understanding both the current operations of the international financial system and the prospects for its future reform.
-As emerging economies continue to challenge Western dominance in global governance, recognising the internal dynamics and potential fractures within the Western coalition becomes increasingly important for scholars, policymakers, and reform advocates alike.
+To summarise, our analysis underscores that Western influence within the IMF is not monolithic but internally differentiated, with Europe shaping access to resources and the United States shaping the terms of their use.
+Recognising this asymmetry complicates prevailing narratives of IMF governance and highlights the need to take transatlantic politics seriously in studies of global financial institutions.
+More broadly, the paper demonstrates the value of disaggregating great power influence rather than treating blocs as unified actors.
+Doing so not only clarifies the institutional dynamics of the IMF, but also provides a framework for examining other organisations where overlapping alliances mask divergent interests.
+In this way, the findings contribute to a more nuanced understanding of how power operates in international economic governance and point toward the importance of looking inside, as well as outside, the Western coalition when considering the future of multilateralism.
 

### 2025-09-18 manuscript/manuscript.qmd

--- 2025-09-12 manuscript/manuscript.qmd
@@
 In doing so, the paper contributes to debates on hegemony, institutional legitimacy, and intra-Western dynamics in global economic governance."
-keywords: International Monetary Fund, hegemony, global economic governance, US influence, European influence, conditionality
+keywords: IMF, US hegemony, Europe
 execute:
@@
 [@stone2011, p.57].
-Numerous case studies further demonstrate instances of US political pressure resulting in lenient loan conditions for key geopolitical allies, including Zaire, Egypt, Mexico, Argentina, and South Korea [@vreeland2009; @copelovitch2010a; @lipscy2018].
+Numerous case studies further demonstrate instances of US political pressure resulting in lenient loan conditions for key geopolitical allies, including Zaire, Egypt, Mexico, Argentina, South Korea, and MENA countries [@vreeland2009; @copelovitch2010a; @lipscy2018; @harrigan2006].
+Closer allies of the United States enjoy IMF loans with less fiscal conditions [@dreher2003].
 This pattern of influence is consistent with the United States’ broader role across international financial institutions, particularly the World Bank, where American dominance is widely acknowledged [@wade2002; @woods2003; @clark2020].
@@
 These studies typically proxy the strategic interests of major powers using indicators such as UN voting alignment [@thacker1999; @dreher2009], bilateral trade intensity [@barro2005], and financial exposure [@oatley2004].
-Yet, much of this literature has remained US-centric, often using only American alignment as the independent variable of interest [@thacker1999; @oatley2004; @dreher2009].
-Other studies pool European indicators with US ones under the broader category of “Western” influence, thus masking potential differences between the two [@lipscy2018; @barro2005].
-These ties have been shown to correlate with more favourable IMF lending outcomes—larger loans, higher approval rates, and fewer or less stringent conditions [@lipscy2018; @barro2005; @dreher2009].
+Yet, much of this literature has remained US-centric, often using only American alignment as the independent variable of interest [@thacker1999; @oatley2004; @dreher2009; @guimaraes2021].
+Other studies pool European indicators with US ones under the broader category of “Western” influence, thus masking potential differences between the two [@lipscy2018; @barro2005; @youssef2023; @guimaraes2021][^2].
+These ties have been shown to correlate with more favourable IMF lending outcomes—larger loans, higher approval rates, and fewer or less stringent conditions [@lipscy2018; @barro2005; @dreher2009; @youssef2023; @guimaraes2021].
 Yet, despite the availability of these tools, comparative assessments between the US and Europe remain largely absent from this literature.
+
+[^2]: @youssef2023 also include China as a major shareholder when calculating political influence.
 
@@
 
--   UN General Assembly voting similarity[^4]
--   Bilateral trade intensity[^5]
--   Banking sector exposure[^6]
+-   UN General Assembly voting similarity[^5]
+-   Bilateral trade intensity[^6]
+-   Banking sector exposure[^7]
 
-[^4]: Originallly used in @thacker1999.
+[^5]: Originally used in @thacker1999.
 
-[^5]: From @barro2005.
+[^6]: From @barro2005.
 
-[^6]: See @broz2006; @lipscy2017; @oatley2004.
+[^7]: See @broz2006; @lipscy2017; @oatley2004.
 
 ```{r percentage_variance}
-load('save/PCA.RData')
+load('save/PCA.Rdata')
 US_PC1_percentage <- US$sdev[1]^2 / sum(US$sdev^2) * 100
@@
 
+We include a range of economic performance indicators that have been shown to systematically affect IMF lending decisions [@barro2005; @knight1997; @youssef2023].
+To capture these dynamics, we include GDP, per capita GDP, and lagged GDP growth as proxies for macroeconomic performance.
+International reserves as a share of imports are used to measure a country’s external liquidity and ability to manage balance-of-payments pressures without IMF support.
+Finally, a binary variable for OECD membership serves as a proxy for broader institutional and economic differences among countries.
+
 The selection of covariates, including their polynomial transformations[^10], closely follows @lipscy2018 and @barro2005 to ensure comparability with previous findings.
@@
 [^10]: Quadratic terms are included for GDP and per capita GDP to capture potential non-linear effects. See [Model Specification](#sec-model) for details.
-
 ## Summary Statistics
@@
 
+# Case Study
+
 # Conclusion
-
-This paper has examined a fundamental yet underexplored question in the study of global economic governance: who wields greater influence over the International Monetary Fund—the United States or Europe?
-By disaggregating "Western" influence into its constituent components and employing principal component analysis to construct separate measures of US and European alignment, we have uncovered important asymmetries in how transatlantic powers exercise their influence within the Fund.
-
-Our empirical analysis yields several key findings that challenge conventional wisdom about IMF decision-making.
-First, European influence demonstrates a broader and more consistent association with favourable lending outcomes than previously recognised.
-Countries with stronger ties to major European powers，namely, the United Kingdom, France, and Germany, receive significantly larger loans relative to GDP, participate more frequently in IMF programmes, and enjoy higher approval rates for new lending arrangements.
-These effects are statistically significant and substantively meaningful, suggesting that European actors wield considerable institutional influence over the Fund's core lending functions.
-
-Second, contrary to much existing literature that emphasises American dominance, US influence shows a more selective pattern of impact.
-While American ties do not significantly affect loan volumes, participation rates, or approval decisions, they are strongly associated with *fewer conditions*, that is, a reduced number of performance criteria, benchmarks, and prior actions attached to lending arrangements.
-This finding suggests that US power manifests not through broad institutional control over access to resources, but through targeted influence over the design of conditionality.
-However, this result should not be interpreted as evidence of systematically “more lenient” conditionality: our additional analyses find no evidence that American ties reduce the required fiscal adjustment imposed by IMF programmes.
-
-Third, our robustness checks reveal that European influence operates on a global rather than regional or colonial basis.
-Neither European borrowers nor African recipients benefit disproportionately from their ties to European powers.
-If anything, the evidence suggests these groups may receive less favourable treatment than other countries with comparable alignment profiles.
-This pattern indicates that European influence reflects genuine institutional power within the IMF rather than geographic or historical favouritism.
-
-These findings have several important implications for our understanding of international financial institutions and global economic governance more broadly.
-Theoretically, they complicate simple narratives of American hegemony in international organisations.
-While the United States undoubtedly remains the most powerful single actor within the Bretton Woods system, our results suggest that European powers collectively exercise substantial influence over key institutional functions.
-This finding aligns with observations about the IMF's historical leadership structure—epitomised by the "gentlemen's agreement" reserving the Managing Director position for Europeans—and suggests that informal power arrangements may be more consequential than formal voting shares or veto rights.
-
-The asymmetric nature of US and European influence also speaks to broader debates about the sources and exercise of institutional power.
-European influence appears to operate through channels that affect aggregate resource allocation—determining who receives assistance and in what amounts.
-American influence, by contrast, manifests through mechanisms that shape the policy content of lending arrangements—affecting what borrowers must do to receive and maintain access to funds.
-This division of labour may reflect different institutional strengths, strategic priorities, or coordination mechanisms between the two actors.
-
-From a policy perspective, these findings have implications for ongoing debates about IMF reform and representation.
-Calls for enhanced emerging market voice and reduced Western dominance must grapple with the reality that "the West" is not a monolithic bloc.
-European and American interests may diverge in important ways, creating opportunities for coalition-building and institutional change that would not exist under unified Western control.
-Understanding these internal dynamics may be crucial for reform advocates seeking to reshape global financial governance.
-
-Our analysis also sheds light on the nature of conditionality, one of the most controversial aspects of IMF operations.
-The finding that American influence is associated with fewer conditions, while European influence shows no such effect, suggests that critiques of overly intrusive conditionality may need to be more precisely targeted.
-European powers, despite their historical association with social democratic governance models, do not appear to advocate for more lenient policy requirements.
-This pattern may reflect the influence of German-style fiscal conservatism within European policy circles or the adoption of neoliberal economic orthodoxy by European technocratic elites.
-
-Several limitations of our analysis should be acknowledged.
-First, our sample period ends in 2010, reflecting the coverage of the underlying datasets we rely on.
-While this prevents us from capturing the European sovereign debt crisis and the subsequent Troika arrangements, we expect that including the post-2010 period would, if anything, reinforce our findings.
-The IMF’s close collaboration with European institutions during the crisis suggests that European influence over lending decisions likely became even more pronounced after our period of observation.
-
-Second, our measures of influence, while grounded in the existing literature, capture alignment rather than direct lobbying or pressure.
-We cannot observe the specific mechanisms through which influence is exercised or the content of behind-the-scenes negotiations.
-Qualitative case studies examining specific lending decisions could provide valuable complementary insights into the micro-foundations of the patterns we document.
-However, such approaches often rely on extensive insider information and interviews with IMF officials, which are difficult to obtain given the Fund’s highly consensus-based and closed-door decision-making culture.
-
-Third, in terms of scope, three clarifications are in order.
-Our analysis focuses exclusively on lending outcomes and does not examine other dimensions of IMF operations, such as surveillance, technical assistance, or institutional governance.
-European and American influence may well manifest differently across these functions.
-Second, the effectiveness of IMF policies is also beyond the scope of our analysis.
-A large body of literature already addresses this question, and its findings can be combined with ours to inform policy debates.
-Third, our approach is deliberately empirical and positivist, aimed at identifying systematic patterns of influence rather than assessing the *appropriateness* of IMF lending arrangements.
-Questions of legitimacy and normative evaluation, long central to debates about the Fund, are not the direct focus of this paper, though we recognise that our results have implications for these debates and welcome further normative interpretations.
-
-Looking ahead, several promising avenues for future research emerge from our findings.
-First, it is worth re-examining the determinants of IMF conditionality by moving beyond simple counts of conditions.
-In addition to the required fiscal adjustment, future work could measure conditionality across specific domains—such as trade liberalisation, privatisation, or structural reforms—to assess whether transatlantic influence shapes not only the quantity but also the substance of IMF policy requirements.
-
-A second line of inquiry would be to examine the domestic sources of IMF influence.
-Our study measures international alignment, but the national actors transmitting influence remain a black box.
-Case studies focusing on finance ministries, central banks, foreign ministries, or private banking interests could shed light on which domestic constituencies mobilise influence within the IMF, and whether these channels differ between Europe and the United States.
-
-Finally, our findings suggest the value of applying a similar disaggregation of Western influence to other international organisations.
-Beyond the World Bank, promising candidates include the OECD, the Bank for International Settlements (BIS), the WTO, and the European Bank for Reconstruction and Development (EBRD).
-Each involves overlapping Western leadership but potentially divergent US and European priorities, offering a fertile testing ground for whether the asymmetries we document are specific to the IMF or reflect broader patterns of transatlantic governance.
-
-In conclusion, this paper demonstrates that the International Monetary Fund, despite its technocratic veneer, remains deeply shaped by geopolitical considerations.
-However, the nature of great power influence is more complex and differentiated than often assumed.
-European powers exercise broad institutional influence over access to IMF resources, while American power manifests more selectively through conditionality.
-This division of influence has important implications for understanding both the current operations of the international financial system and the prospects for its future reform.
-As emerging economies continue to challenge Western dominance in global governance, recognising the internal dynamics and potential fractures within the Western coalition becomes increasingly important for scholars, policymakers, and reform advocates alike.
 
@@
 #| fig-pos: H
-load('save/scree.RData')
+load('save/scree.Rdata')
 scree_EU

### 2025-09-18 manuscript/manuscript.qmd
Add back blank line
Move the definition of conditionality

@@
 [@stone2011, p.57].
+An abundant body of literature highlights American influence over the *conditionality* attached to IMF loans, defined as the practice of providing financial assistance contingent on the implementation of specific policies [@dreher2009a].
 Numerous case studies further demonstrate instances of US political pressure resulting in lenient loan conditions for key geopolitical allies, including Zaire, Egypt, Mexico, Argentina, South Korea, and MENA countries [@vreeland2009; @copelovitch2010a; @lipscy2018; @harrigan2006].
@@
 [^10]: Quadratic terms are included for GDP and per capita GDP to capture potential non-linear effects. See [Model Specification](#sec-model) for details.
+
 ## Summary Statistics
@@
 This appendix extends our analysis of IMF conditionality by examining alternative measures that focus specifically on fiscal policy requirements.
-Fiscal conditionality, under a comprehensive review of both theoretical approaches and empirical evidence, is "the practice of giving financial assistance contingent on the implementation of specific policies", and it continues to rise over the years [@dreher2009a].
 Following @guimaraes2021, we consider not only the total number of conditions (as in our main analysis), but also the number of *fiscal* conditions and the required fiscal adjustment (RFA)[^14]—the extent of fiscal tightening mandated by IMF programmes, expressed as a percentage of GDP.

### 2025-09-18 manuscript/manuscript.qmd
update footnote counters

@@
 
-[^9]: The first principal components capture the majority of the variance (`r format(EU_PC1_percentage, digits = 1)`% for Europe and `r format(US_PC1_percentage, digits = 1)`% for the United States).
-See Appendix [-@sec-scree] for scree plots and additional details.
+[^9]: GDP, before log transformation, is measured in millions of constant 2002 US dollars, while GDP per capita is measured in thousands of constant 2002 US dollars.
 
@@
 We include a range of economic performance indicators that have been shown to systematically affect IMF lending decisions [@barro2005; @knight1997].
-To capture these dynamics, we include log GDP, per capita GDP, and lagged GDP growth as proxies for macroeconomic performance[^9].
+To capture these dynamics, we include log GDP, per capita GDP, and lagged GDP growth as proxies for macroeconomic performance[^10].
 International reserves measured in months of imports are used to measure a country’s external liquidity and ability to manage balance-of-payments pressures without IMF support.
@@
 
-[^9]: GDP, before log transformation, is measured in millions of constant 2002 US dollars, while GDP per capita is measured in thousands of constant 2002 US dollars.
+[^10]: GDP, before log transformation, is measured in millions of constant 2002 US dollars, while GDP per capita is measured in thousands of constant 2002 US dollars.
 
@@
 
-The selection of covariates, including their polynomial transformations[^10], closely follows @lipscy2018 and @barro2005 to ensure comparability with previous findings.
+The selection of covariates, including their polynomial transformations[^11], closely follows @lipscy2018 and @barro2005 to ensure comparability with previous findings.
 
-[^10]: Quadratic terms are included for GDP and per capita GDP to capture potential non-linear effects. See [Model Specification](#sec-model) for details.
+[^11]: Quadratic terms are included for GDP and per capita GDP to capture potential non-linear effects. See [Model Specification](#sec-model) for details.
 
@@
 -   $\lambda_t$ denotes five-year period fixed effects.
--   $\epsilon_{it}$ is the error term, assumed to be independently and identically distributed as $\mathcal{N}(0, \sigma^2)$[^11].
+-   $\epsilon_{it}$ is the error term, assumed to be independently and identically distributed as $\mathcal{N}(0, \sigma^2)$[^12].
 
-[^11]: Country fixed effects are not included because the number of time periods ($T$) is small relative to the number of countries ($N$).
+[^12]: Country fixed effects are not included because the number of time periods ($T$) is small relative to the number of countries ($N$).
 In such cases, including country fixed effects would lead to inconsistent estimates of the disturbance variance $\sigma$ [@greene2004].
@@
 This allows us to conveniently reuse the same latent variable specification $Y^*_{it}$ as in the Tobit model.
-The key differences are that the error term is now assumed to follow a standard normal distribution $\mathcal{N}(0, 1)$, and that the latent outcome maps to a binary observed outcome as follows[^12]:
+The key differences are that the error term is now assumed to follow a standard normal distribution $\mathcal{N}(0, 1)$, and that the latent outcome maps to a binary observed outcome as follows[^13]:
 
-[^12]: Alternatively, the Probit model can be expressed in terms of probabilities as $\Pr(Y_{it}=1)=\Phi(\alpha + \beta_1 \text{USA}_{it} + \beta_2 \text{EUP}_{it} + \gamma^\top \mathbf{P}_{it} + \delta^\top \mathbf{Z}_{it} + \lambda_t)$, where $\Phi(\cdot)$ is the standard normal cumulative distribution function.
+[^13]: Alternatively, the Probit model can be expressed in terms of probabilities as $\Pr(Y_{it}=1)=\Phi(\alpha + \beta_1 \text{USA}_{it} + \beta_2 \text{EUP}_{it} + \gamma^\top \mathbf{P}_{it} + \delta^\top \mathbf{Z}_{it} + \lambda_t)$, where $\Phi(\cdot)$ is the standard normal cumulative distribution function.
 
@@
 In contrast, similar ties with the United States do not yield statistically significant effects on these outcomes.
-The differences between American and European influence on the first two outcomes are statistically significant at the 5% level[^13].
+The differences between American and European influence on the first two outcomes are statistically significant at the 5% level[^14].
 
-[^13]: Based on tests of equality for the null hypothesis that $\text{USA} = \text{EUP}$.
+[^14]: Based on tests of equality for the null hypothesis that $\text{USA} = \text{EUP}$.
 $F$-tests are used for the linear models; for the Probit model, the test corresponds to an asymptotic Chi-squared statistic, which is appropriate given the large sample size ($N = 522$).
@@
 This appendix extends our analysis of IMF conditionality by examining alternative measures that focus specifically on fiscal policy requirements.
-Following @guimaraes2021, we consider not only the total number of conditions (as in our main analysis), but also the number of *fiscal* conditions and the required fiscal adjustment (RFA)[^14]—the extent of fiscal tightening mandated by IMF programmes, expressed as a percentage of GDP.
+Following @guimaraes2021, we consider not only the total number of conditions (as in our main analysis), but also the number of *fiscal* conditions and the required fiscal adjustment (RFA)[^15]—the extent of fiscal tightening mandated by IMF programmes, expressed as a percentage of GDP.
 We report RFA at two time horizons: one year after programme approval ($T+1$) and at the programme’s end year.
@@
 
-[^14]: To merge these variables with our dataset, aggregation is necessary: our data are at the five-year country level, whereas theirs are at the programme level.
+[^15]: To merge these variables with our dataset, aggregation is necessary: our data are at the five-year country level, whereas theirs are at the programme level.
 We sum the number of fiscal conditions and average the RFA variables.

### 2025-09-18 manuscript/manuscript.qmd
Fix conditionality literature

@@
 [@stone2011, p.57].
-An abundant body of literature highlights American influence over the *conditionality* attached to IMF loans, defined as the practice of providing financial assistance contingent on the implementation of specific policies [@dreher2009a].
-Numerous case studies further demonstrate instances of US political pressure resulting in lenient loan conditions for key geopolitical allies, including Zaire, Egypt, Mexico, Argentina, South Korea, and MENA countries [@vreeland2009; @copelovitch2010a; @lipscy2018; @harrigan2006].
-Closer allies of the United States enjoy IMF loans with less fiscal conditions [@dreher2003].
+An abundant body of literature highlights American influence over the *conditionality* attached to IMF loans, defined as "the practice of providing financial assistance contingent on the implementation of specific policies" [@dreher2009a].
+Numerous case studies further demonstrate US political pressure over conditionality, including in Zaire, Egypt, Mexico, Argentina, South Korea, and several MENA countries [@vreeland2009; @copelovitch2010a; @lipscy2018; @harrigan2006].
 This pattern of influence is consistent with the United States’ broader role across international financial institutions, particularly the World Bank, where American dominance is widely acknowledged [@wade2002; @woods2003; @clark2020].
@@
 Yet, much of this literature has remained US-centric, often using only American alignment as the independent variable of interest [@thacker1999; @oatley2004; @dreher2009; @guimaraes2021].
-Other studies pool European indicators with US ones under the broader category of “Western” influence, thus masking potential differences between the two [@lipscy2018; @barro2005; @youssef2023; @guimaraes2021][^2].
+Other studies pool European indicators with US ones under the broader category of “Western” influence, thus masking potential differences between the two [@lipscy2018; @barro2005; @youssef2023; @guimaraes2021; @dreher2003][^2].
 These ties have been shown to correlate with more favourable IMF lending outcomes—larger loans, higher approval rates, and fewer or less stringent conditions [@lipscy2018; @barro2005; @dreher2009; @youssef2023; @guimaraes2021].

### 2025-09-18 manuscript/manuscript.qmd
add 2 literatures to the current version (#59)
* add 2 literatures to the current version

* Fix escaping

* spaces before brackets

* Add Recent determinants (#63)

* Add footnote explaining youssef2023

* Add youssef2023 to economic indicators

* Add guimaraes (2021) to lit review

* Add harrigan to qualitative studies

* Apply GPT review suggestions

* Merge branch 'add-literature' into recent-determinants

* Add back blank line

* Move the definition of conditionality

* update footnote counters

* Fix conditionality literature

@@
 In doing so, the paper contributes to debates on hegemony, institutional legitimacy, and intra-Western dynamics in global economic governance."
-keywords: IMF, US hegemony, Europe
+keywords: International Monetary Fund, hegemony, global economic governance, US influence, European influence, conditionality
 execute:
@@
 
-# Case Study
+# Conclusion
 
-# Conclusion
+This paper has examined a fundamental yet underexplored question in the study of global economic governance: who wields greater influence over the International Monetary Fund—the United States or Europe?
+By disaggregating "Western" influence into its constituent components and employing principal component analysis to construct separate measures of US and European alignment, we have uncovered important asymmetries in how transatlantic powers exercise their influence within the Fund.
+
+Our empirical analysis yields several key findings that challenge conventional wisdom about IMF decision-making.
+First, European influence demonstrates a broader and more consistent association with favourable lending outcomes than previously recognised.
+Countries with stronger ties to major European powers，namely, the United Kingdom, France, and Germany, receive significantly larger loans relative to GDP, participate more frequently in IMF programmes, and enjoy higher approval rates for new lending arrangements.
+These effects are statistically significant and substantively meaningful, suggesting that European actors wield considerable institutional influence over the Fund's core lending functions.
+
+Second, contrary to much existing literature that emphasises American dominance, US influence shows a more selective pattern of impact.
+While American ties do not significantly affect loan volumes, participation rates, or approval decisions, they are strongly associated with *fewer conditions*, that is, a reduced number of performance criteria, benchmarks, and prior actions attached to lending arrangements.
+This finding suggests that US power manifests not through broad institutional control over access to resources, but through targeted influence over the design of conditionality.
+However, this result should not be interpreted as evidence of systematically “more lenient” conditionality: our additional analyses find no evidence that American ties reduce the required fiscal adjustment imposed by IMF programmes.
+
+Third, our robustness checks reveal that European influence operates on a global rather than regional or colonial basis.
+Neither European borrowers nor African recipients benefit disproportionately from their ties to European powers.
+If anything, the evidence suggests these groups may receive less favourable treatment than other countries with comparable alignment profiles.
+This pattern indicates that European influence reflects genuine institutional power within the IMF rather than geographic or historical favouritism.
+
+These findings have several important implications for our understanding of international financial institutions and global economic governance more broadly.
+Theoretically, they complicate simple narratives of American hegemony in international organisations.
+While the United States undoubtedly remains the most powerful single actor within the Bretton Woods system, our results suggest that European powers collectively exercise substantial influence over key institutional functions.
+This finding aligns with observations about the IMF's historical leadership structure—epitomised by the "gentlemen's agreement" reserving the Managing Director position for Europeans—and suggests that informal power arrangements may be more consequential than formal voting shares or veto rights.
+
+The asymmetric nature of US and European influence also speaks to broader debates about the sources and exercise of institutional power.
+European influence appears to operate through channels that affect aggregate resource allocation—determining who receives assistance and in what amounts.
+American influence, by contrast, manifests through mechanisms that shape the policy content of lending arrangements—affecting what borrowers must do to receive and maintain access to funds.
+This division of labour may reflect different institutional strengths, strategic priorities, or coordination mechanisms between the two actors.
+
+From a policy perspective, these findings have implications for ongoing debates about IMF reform and representation.
+Calls for enhanced emerging market voice and reduced Western dominance must grapple with the reality that "the West" is not a monolithic bloc.
+European and American interests may diverge in important ways, creating opportunities for coalition-building and institutional change that would not exist under unified Western control.
+Understanding these internal dynamics may be crucial for reform advocates seeking to reshape global financial governance.
+
+Our analysis also sheds light on the nature of conditionality, one of the most controversial aspects of IMF operations.
+The finding that American influence is associated with fewer conditions, while European influence shows no such effect, suggests that critiques of overly intrusive conditionality may need to be more precisely targeted.
+European powers, despite their historical association with social democratic governance models, do not appear to advocate for more lenient policy requirements.
+This pattern may reflect the influence of German-style fiscal conservatism within European policy circles or the adoption of neoliberal economic orthodoxy by European technocratic elites.
+
+Several limitations of our analysis should be acknowledged.
+First, our sample period ends in 2010, reflecting the coverage of the underlying datasets we rely on.
+While this prevents us from capturing the European sovereign debt crisis and the subsequent Troika arrangements, we expect that including the post-2010 period would, if anything, reinforce our findings.
+The IMF’s close collaboration with European institutions during the crisis suggests that European influence over lending decisions likely became even more pronounced after our period of observation.
+
+Second, our measures of influence, while grounded in the existing literature, capture alignment rather than direct lobbying or pressure.
+We cannot observe the specific mechanisms through which influence is exercised or the content of behind-the-scenes negotiations.
+Qualitative case studies examining specific lending decisions could provide valuable complementary insights into the micro-foundations of the patterns we document.
+However, such approaches often rely on extensive insider information and interviews with IMF officials, which are difficult to obtain given the Fund’s highly consensus-based and closed-door decision-making culture.
+
+Third, in terms of scope, three clarifications are in order.
+Our analysis focuses exclusively on lending outcomes and does not examine other dimensions of IMF operations, such as surveillance, technical assistance, or institutional governance.
+European and American influence may well manifest differently across these functions.
+Second, the effectiveness of IMF policies is also beyond the scope of our analysis.
+A large body of literature already addresses this question, and its findings can be combined with ours to inform policy debates.
+Third, our approach is deliberately empirical and positivist, aimed at identifying systematic patterns of influence rather than assessing the *appropriateness* of IMF lending arrangements.
+Questions of legitimacy and normative evaluation, long central to debates about the Fund, are not the direct focus of this paper, though we recognise that our results have implications for these debates and welcome further normative interpretations.
+
+Looking ahead, several promising avenues for future research emerge from our findings.
+First, it is worth re-examining the determinants of IMF conditionality by moving beyond simple counts of conditions.
+In addition to the required fiscal adjustment, future work could measure conditionality across specific domains—such as trade liberalisation, privatisation, or structural reforms—to assess whether transatlantic influence shapes not only the quantity but also the substance of IMF policy requirements.
+
+A second line of inquiry would be to examine the domestic sources of IMF influence.
+Our study measures international alignment, but the national actors transmitting influence remain a black box.
+Case studies focusing on finance ministries, central banks, foreign ministries, or private banking interests could shed light on which domestic constituencies mobilise influence within the IMF, and whether these channels differ between Europe and the United States.
+
+Finally, our findings suggest the value of applying a similar disaggregation of Western influence to other international organisations.
+Beyond the World Bank, promising candidates include the OECD, the Bank for International Settlements (BIS), the WTO, and the European Bank for Reconstruction and Development (EBRD).
+Each involves overlapping Western leadership but potentially divergent US and European priorities, offering a fertile testing ground for whether the asymmetries we document are specific to the IMF or reflect broader patterns of transatlantic governance.
+
+To summarise, our analysis underscores that Western influence within the IMF is not monolithic but internally differentiated, with Europe shaping access to resources and the United States shaping the terms of their use.
+Recognising this asymmetry complicates prevailing narratives of IMF governance and highlights the need to take transatlantic politics seriously in studies of global financial institutions.
+More broadly, the paper demonstrates the value of disaggregating great power influence rather than treating blocs as unified actors.
+Doing so not only clarifies the institutional dynamics of the IMF, but also provides a framework for examining other organisations where overlapping alliances mask divergent interests.
+In this way, the findings contribute to a more nuanced understanding of how power operates in international economic governance and point toward the importance of looking inside, as well as outside, the Western coalition when considering the future of multilateralism.
 

### 2025-09-21 manuscript/manuscript.qmd
Add footnote for loan types (#64)

@@
 
-In line with @lipscy2018 and @barro2005, we use the following four measures of IMF lending outcomes as our dependent variables:
+In line with @lipscy2018 and @barro2005, we use the following four measures of IMF lending outcomes as our dependent variables[^4]:
 
@@
 
+[^4]: We only include Stand-By Arrangements (SBA) and Extended Fund Facility (EFF) programmes for comparability with prior studies [@lipscy2018; @barro2005; @dreher2006].
+These non-concessional programmes are the IMF’s main channels of financial assistance and are most subject to political influence, especially relative to concessional programmes [@youssef2023].
+
 ## Independent Variables
@@
 These propensities are based on widely used proxies from the existing literature, which reflect the degree of a country’s political and economic alignment with the US or Europe [@lipscy2018].
-This approach assumes that both the US and European powers[^4] are more likely to advocate for lending decisions that benefit their close partners.
+This approach assumes that both the US and European powers[^5] are more likely to advocate for lending decisions that benefit their close partners.
 
-[^4]: For operationalisation, we focus only on ties to the three largest Western European powers—namely, the United Kingdom, France, and Germany—and use their average values.
+[^5]: For operationalisation, we focus only on ties to the three largest Western European powers—namely, the United Kingdom, France, and Germany—and use their average values.
 This approach reflects the assumption that smaller European countries are unlikely to exert meaningful influence over IMF decisions on behalf of their partners.
@@
 
--   UN General Assembly voting similarity[^5]
--   Bilateral trade intensity[^6]
--   Banking sector exposure[^7]
+-   UN General Assembly voting similarity[^6]
+-   Bilateral trade intensity[^7]
+-   Banking sector exposure[^8]
 
-[^5]: Originally used in @thacker1999.
+[^6]: Originally used in @thacker1999.
 
-[^6]: From @barro2005.
+[^7]: From @barro2005.
 
-[^7]: See @broz2006; @lipscy2017; @oatley2004.
+[^8]: See @broz2006; @lipscy2017; @oatley2004.
 
@@
 
-However, as noted by @lipscy2018 and @belsley2004, including multiple proxies for the same latent construct in a single regression model leads to multicollinearity, which reduces statistical power and inflates standard errors, as the variables are highly correlated in our data[^8].
+However, as noted by @lipscy2018 and @belsley2004, including multiple proxies for the same latent construct in a single regression model leads to multicollinearity, which reduces statistical power and inflates standard errors, as the variables are highly correlated in our data[^9].
 To address this issue, we apply principal component analysis (PCA) to reduce the dimensionality of the data [@dunteman2008].
-Specifically, we extract the first principal component — a linear combination of the original variables that captures the greatest share of variance — to serve as a single measure of influence[^9].
+Specifically, we extract the first principal component — a linear combination of the original variables that captures the greatest share of variance — to serve as a single measure of influence[^10].
 
-[^8]: See Appendix [-@sec-corr].
+[^9]: See Appendix [-@sec-corr].
 
-[^9]: GDP, before log transformation, is measured in millions of constant 2002 US dollars, while GDP per capita is measured in thousands of constant 2002 US dollars.
+[^10]: GDP, before log transformation, is measured in millions of constant 2002 US dollars, while GDP per capita is measured in thousands of constant 2002 US dollars.
 
@@
 We include a range of economic performance indicators that have been shown to systematically affect IMF lending decisions [@barro2005; @knight1997].
-To capture these dynamics, we include log GDP, per capita GDP, and lagged GDP growth as proxies for macroeconomic performance[^10].
+To capture these dynamics, we include log GDP, per capita GDP, and lagged GDP growth as proxies for macroeconomic performance[^11].
 International reserves measured in months of imports are used to measure a country’s external liquidity and ability to manage balance-of-payments pressures without IMF support.
@@
 
-[^10]: GDP, before log transformation, is measured in millions of constant 2002 US dollars, while GDP per capita is measured in thousands of constant 2002 US dollars.
+[^11]: GDP, before log transformation, is measured in millions of constant 2002 US dollars, while GDP per capita is measured in thousands of constant 2002 US dollars.
 
@@
 
-The selection of covariates, including their polynomial transformations[^11], closely follows @lipscy2018 and @barro2005 to ensure comparability with previous findings.
+The selection of covariates, including their polynomial transformations[^12], closely follows @lipscy2018 and @barro2005 to ensure comparability with previous findings.
 
-[^11]: Quadratic terms are included for GDP and per capita GDP to capture potential non-linear effects. See [Model Specification](#sec-model) for details.
+[^12]: Quadratic terms are included for GDP and per capita GDP to capture potential non-linear effects. See [Model Specification](#sec-model) for details.
 
@@
 -   $\lambda_t$ denotes five-year period fixed effects.
--   $\epsilon_{it}$ is the error term, assumed to be independently and identically distributed as $\mathcal{N}(0, \sigma^2)$[^12].
+-   $\epsilon_{it}$ is the error term, assumed to be independently and identically distributed as $\mathcal{N}(0, \sigma^2)$[^13].
 
-[^12]: Country fixed effects are not included because the number of time periods ($T$) is small relative to the number of countries ($N$).
+[^13]: Country fixed effects are not included because the number of time periods ($T$) is small relative to the number of countries ($N$).
 In such cases, including country fixed effects would lead to inconsistent estimates of the disturbance variance $\sigma$ [@greene2004].
@@
 This allows us to conveniently reuse the same latent variable specification $Y^*_{it}$ as in the Tobit model.
-The key differences are that the error term is now assumed to follow a standard normal distribution $\mathcal{N}(0, 1)$, and that the latent outcome maps to a binary observed outcome as follows[^13]:
+The key differences are that the error term is now assumed to follow a standard normal distribution $\mathcal{N}(0, 1)$, and that the latent outcome maps to a binary observed outcome as follows[^14]:
 
-[^13]: Alternatively, the Probit model can be expressed in terms of probabilities as $\Pr(Y_{it}=1)=\Phi(\alpha + \beta_1 \text{USA}_{it} + \beta_2 \text{EUP}_{it} + \gamma^\top \mathbf{P}_{it} + \delta^\top \mathbf{Z}_{it} + \lambda_t)$, where $\Phi(\cdot)$ is the standard normal cumulative distribution function.
+[^14]: Alternatively, the Probit model can be expressed in terms of probabilities as $\Pr(Y_{it}=1)=\Phi(\alpha + \beta_1 \text{USA}_{it} + \beta_2 \text{EUP}_{it} + \gamma^\top \mathbf{P}_{it} + \delta^\top \mathbf{Z}_{it} + \lambda_t)$, where $\Phi(\cdot)$ is the standard normal cumulative distribution function.
 
@@
 In contrast, similar ties with the United States do not yield statistically significant effects on these outcomes.
-The differences between American and European influence on the first two outcomes are statistically significant at the 5% level[^14].
+The differences between American and European influence on the first two outcomes are statistically significant at the 5% level[^15].
 
-[^14]: Based on tests of equality for the null hypothesis that $\text{USA} = \text{EUP}$.
+[^15]: Based on tests of equality for the null hypothesis that $\text{USA} = \text{EUP}$.
 $F$-tests are used for the linear models; for the Probit model, the test corresponds to an asymptotic Chi-squared statistic, which is appropriate given the large sample size ($N = 522$).
@@
 This appendix extends our analysis of IMF conditionality by examining alternative measures that focus specifically on fiscal policy requirements.
-Following @guimaraes2021, we consider not only the total number of conditions (as in our main analysis), but also the number of *fiscal* conditions and the required fiscal adjustment (RFA)[^15]—the extent of fiscal tightening mandated by IMF programmes, expressed as a percentage of GDP.
+Following @guimaraes2021, we consider not only the total number of conditions (as in our main analysis), but also the number of *fiscal* conditions and the required fiscal adjustment (RFA)[^16]—the extent of fiscal tightening mandated by IMF programmes, expressed as a percentage of GDP.
 We report RFA at two time horizons: one year after programme approval ($T+1$) and at the programme’s end year.
@@
 
-[^15]: To merge these variables with our dataset, aggregation is necessary: our data are at the five-year country level, whereas theirs are at the programme level.
+[^16]: To merge these variables with our dataset, aggregation is necessary: our data are at the five-year country level, whereas theirs are at the programme level.
 We sum the number of fiscal conditions and average the RFA variables.

### 2025-09-24 manuscript/manuscript.qmd
drop legitimacy from intro

@@
 Third, it challenges the assumption of geographic bias, showing that European powers do not disproportionately favour European borrowers.
-These findings shed new light on the internal politics of the IMF and speak to wider debates on institutional legitimacy, great power contestation, and the geopolitics of global finance.
+These findings shed new light on the internal politics of the IMF and speak to wider debates on great power contestation and the geopolitics of global finance.
 
@@
 
+::: {.data-availability}
+**Data Availability Statement**
+
+The replication files for this paper are available at [https://github.com/IMFPaper/IMF](https://github.com/IMFPaper/IMF).
+:::
+
+::: {.funding}
+**Funding Statement**
+
+This research is not funded.
+:::
+
+::: {.competing-interests}
+**Competing Interests**
+
+The authors declare none.
+:::
+
 \newpage

### 2025-09-24 manuscript/manuscript.qmd
abstract: change legitimacy to power

@@
 They also suggest that Europe wields broader institutional influence over access to IMF resources, while US power manifests more selectively through conditionality.
-In doing so, the paper contributes to debates on hegemony, institutional legitimacy, and intra-Western dynamics in global economic governance."
+In doing so, the paper contributes to debates on hegemony, institutional power, and intra-Western dynamics in global economic governance."
 keywords: International Monetary Fund, hegemony, global economic governance, US influence, European influence, conditionality

### 2025-09-24 manuscript/manuscript.qmd
conclusion: update para on hegemony

@@
 These findings have several important implications for our understanding of international financial institutions and global economic governance more broadly.
-Theoretically, they complicate simple narratives of American hegemony in international organisations.
-While the United States undoubtedly remains the most powerful single actor within the Bretton Woods system, our results suggest that European powers collectively exercise substantial influence over key institutional functions.
-This finding aligns with observations about the IMF's historical leadership structure—epitomised by the "gentlemen's agreement" reserving the Managing Director position for Europeans—and suggests that informal power arrangements may be more consequential than formal voting shares or veto rights.
+Theoretically, these results complicate conventional narratives of uncontested US hegemony in the IMF.
+While the United States remains the most powerful single actor within the Bretton Woods system, its influence is concentrated in shaping conditionality rather than in determining access to IMF resources.
+By contrast, European powers collectively wield substantial institutional authority over lending volumes and approvals, reflecting the Fund’s historical leadership structure epitomised by the “gentlemen’s agreement.” Hegemony within the IMF, therefore, is fragmented rather than monolithic, with American primacy tempered by European institutional strength.
 

### 2025-09-24 manuscript/manuscript.qmd
Add para on broad picture of hegemony

@@
 
-The asymmetric nature of US and European influence also speaks to broader debates about the sources and exercise of institutional power.
-European influence appears to operate through channels that affect aggregate resource allocation—determining who receives assistance and in what amounts.
-American influence, by contrast, manifests through mechanisms that shape the policy content of lending arrangements—affecting what borrowers must do to receive and maintain access to funds.
-This division of labour may reflect different institutional strengths, strategic priorities, or coordination mechanisms between the two actors.
+More broadly, the practice of IMF governance underscores that even the United States, as the system’s hegemon, operates under constraints.
+Securing outcomes typically requires consensus from other major powers, most notably European states, and compromises are often struck to maintain institutional cohesion.
+US influence also depends on a degree of legitimation, whether through the appearance of multilateral power-sharing, as in the transatlantic balance within the IMF, or through the promotion of ideological frameworks such as neoliberal economic orthodoxy.
+Yet such strategies can produce unintended consequences: Europe has not only embraced but also entrenched neoliberal doctrines, allowing it to leverage them in ways that enhance its own institutional authority.
+This dynamic illustrates a broader pattern in which American-founded institutions, originally designed to project US preferences, may evolve into arenas that constrain or even oppose US objectives, as exemplified by the UN Security Council’s opposition to the Iraq War.
+Hegemony, in this light, is both exercised and contested within the very structures the United States helped to create.
 

### 2025-09-24 manuscript/manuscript.qmd
Refine the para above

@@
 
-These findings have several important implications for our understanding of international financial institutions and global economic governance more broadly.
-Theoretically, these results complicate conventional narratives of uncontested US hegemony in the IMF.
-While the United States remains the most powerful single actor within the Bretton Woods system, its influence is concentrated in shaping conditionality rather than in determining access to IMF resources.
-By contrast, European powers collectively wield substantial institutional authority over lending volumes and approvals, reflecting the Fund’s historical leadership structure epitomised by the “gentlemen’s agreement.” Hegemony within the IMF, therefore, is fragmented rather than monolithic, with American primacy tempered by European institutional strength.
+These findings have several important implications for our understanding of international financial institutions and global economic governance.
+Theoretically, they complicate conventional narratives of uncontested US hegemony in the IMF.
+While the United States remains the most powerful single actor within the Bretton Woods system, its influence is concentrated in shaping the *policy content* of lending arrangements—determining what borrowers must do to receive and maintain access to funds—rather than in deciding who receives assistance.
+By contrast, European powers collectively wield substantial institutional authority over *resource allocation*, influencing which countries obtain loans and in what amounts.
+This pattern, rooted in the Fund’s historical leadership structure epitomised by the “gentlemen’s agreement,” suggests that hegemony within the IMF is fragmented rather than monolithic, with American primacy tempered by European institutional strength.
 

### 2025-09-24 manuscript/manuscript.qmd
Add clarification on personal networks

@@
 
-Third, in terms of scope, three clarifications are in order.
+Third, in terms of scope, four clarifications are in order.
 Our analysis focuses exclusively on lending outcomes and does not examine other dimensions of IMF operations, such as surveillance, technical assistance, or institutional governance.
@@
 Questions of legitimacy and normative evaluation, long central to debates about the Fund, are not the direct focus of this paper, though we recognise that our results have implications for these debates and welcome further normative interpretations.
+Fourth, while our analysis focuses on formal indicators of alignment, we also recognise that informal personal networks can play an important role in IMF politics.
+The relationships that borrowing-country officials cultivate within international financial institutions may shape loan negotiations and outcomes in ways that are not captured by our framework.
+Assessing the impact of such networks, however, falls outside the scope of this study.
 

### 2025-09-24 manuscript/manuscript.qmd
Add declarations

--- 2025-09-24 manuscript/manuscript.qmd
@@
 They also suggest that Europe wields broader institutional influence over access to IMF resources, while US power manifests more selectively through conditionality.
-In doing so, the paper contributes to debates on hegemony, institutional power, and intra-Western dynamics in global economic governance."
+In doing so, the paper contributes to debates on hegemony, institutional legitimacy, and intra-Western dynamics in global economic governance."
 keywords: International Monetary Fund, hegemony, global economic governance, US influence, European influence, conditionality
@@
 Third, it challenges the assumption of geographic bias, showing that European powers do not disproportionately favour European borrowers.
-These findings shed new light on the internal politics of the IMF and speak to wider debates on great power contestation and the geopolitics of global finance.
+These findings shed new light on the internal politics of the IMF and speak to wider debates on institutional legitimacy, great power contestation, and the geopolitics of global finance.
 
@@
 These findings have several important implications for our understanding of international financial institutions and global economic governance more broadly.
-Theoretically, these results complicate conventional narratives of uncontested US hegemony in the IMF.
-While the United States remains the most powerful single actor within the Bretton Woods system, its influence is concentrated in shaping conditionality rather than in determining access to IMF resources.
-By contrast, European powers collectively wield substantial institutional authority over lending volumes and approvals, reflecting the Fund’s historical leadership structure epitomised by the “gentlemen’s agreement.” Hegemony within the IMF, therefore, is fragmented rather than monolithic, with American primacy tempered by European institutional strength.
+Theoretically, they complicate simple narratives of American hegemony in international organisations.
+While the United States undoubtedly remains the most powerful single actor within the Bretton Woods system, our results suggest that European powers collectively exercise substantial influence over key institutional functions.
+This finding aligns with observations about the IMF's historical leadership structure—epitomised by the "gentlemen's agreement" reserving the Managing Director position for Europeans—and suggests that informal power arrangements may be more consequential than formal voting shares or veto rights.
 
-More broadly, the practice of IMF governance underscores that even the United States, as the system’s hegemon, operates under constraints.
-Securing outcomes typically requires consensus from other major powers, most notably European states, and compromises are often struck to maintain institutional cohesion.
-US influence also depends on a degree of legitimation, whether through the appearance of multilateral power-sharing, as in the transatlantic balance within the IMF, or through the promotion of ideological frameworks such as neoliberal economic orthodoxy.
-Yet such strategies can produce unintended consequences: Europe has not only embraced but also entrenched neoliberal doctrines, allowing it to leverage them in ways that enhance its own institutional authority.
-This dynamic illustrates a broader pattern in which American-founded institutions, originally designed to project US preferences, may evolve into arenas that constrain or even oppose US objectives, as exemplified by the UN Security Council’s opposition to the Iraq War.
-Hegemony, in this light, is both exercised and contested within the very structures the United States helped to create.
+The asymmetric nature of US and European influence also speaks to broader debates about the sources and exercise of institutional power.
+European influence appears to operate through channels that affect aggregate resource allocation—determining who receives assistance and in what amounts.
+American influence, by contrast, manifests through mechanisms that shape the policy content of lending arrangements—affecting what borrowers must do to receive and maintain access to funds.
+This division of labour may reflect different institutional strengths, strategic priorities, or coordination mechanisms between the two actors.
 
@@
 ::: {.data-availability}
-**Data Availability Statement**
-
-The replication files for this paper are available at [https://github.com/IMFPaper/IMF](https://github.com/IMFPaper/IMF).
+**Data Availability Statement**  The replication files for this paper are available at [https://github.com/IMFPaper/IMF](https://github.com/IMFPaper/IMF).
 :::
@@
 ::: {.funding}
-**Funding Statement**
-
-This research is not funded.
+**Funding Statement**  This research is not funded.
 :::
@@
 ::: {.competing-interests}
-**Competing Interests**
-
-The authors declare none.
+**Competing Interests**  The authors declare none.
 :::

### 2025-09-24 manuscript/manuscript.qmd
Add declarations (#68)

@@
 ::: {.data-availability}
-**Data Availability Statement**  The replication files for this paper are available at [https://github.com/IMFPaper/IMF](https://github.com/IMFPaper/IMF).
+**Data Availability Statement**
+
+The replication files for this paper are available at [https://github.com/IMFPaper/IMF](https://github.com/IMFPaper/IMF).
 :::
@@
 ::: {.funding}
-**Funding Statement**  This research is not funded.
+**Funding Statement**
+
+This research is not funded.
 :::
@@
 ::: {.competing-interests}
-**Competing Interests**  The authors declare none.
+**Competing Interests**
+
+The authors declare none.
 :::

### 2025-10-04 tests/testthat/test-manuscript.R
change expected length to 400

@@ -8,7 +8,7 @@ test_that("manuscript.qmd file exists and is readable", {
   
   # Check that file can be read
   manuscript_content <- readLines(here("manuscript/manuscript.qmd"), warn = FALSE)
-  expect_gt(length(manuscript_content), 100,
+  expect_gt(length(manuscript_content), 400,
             label = "manuscript.qmd should have substantial content")
 })
 

### 2025-10-04 tests/testthat/test-manuscript.R
remove non-dependencies

@@ -18,13 +18,9 @@ test_that("all required data dependencies exist", {
     "save/corr.RData",
     "save/PCA.RData", 
     "save/scree.RData",
-    "save/regModels.RData",
     "save/regTable.RData",
-    "save/regModels_africa.RData",
     "save/regTable_africa.RData",
-    "save/regModels_europe.RData",
     "save/regTable_europe.RData",
-    "save/regModels_fiscal.RData",
     "save/regTable_fiscal.RData",
     "data/panel_data_pca.rds"
   )

### 2025-10-04 tests/testthat/test-manuscript.R
Fix that part that checks loading

@@ -68,7 +68,8 @@ test_that("manuscript.qmd loads all required data files correctly", {
     "save/regTable.RData",
     "save/regTable_africa.RData",
     "save/regTable_europe.RData",
-    "save/regTable_fiscal.RData"
+    "save/regTable_fiscal.RData",
+    "data/panel_data_pca.rds"
   )
   
   for (file in expected_loads) {

### 2025-10-04 tests/testthat/test-manuscript.R
Remove final parts

@@ -79,175 +79,3 @@ test_that("manuscript.qmd loads all required data files correctly", {
   }
 })
 
-test_that("manuscript.qmd uses tinytable correctly", {
-  # Read the manuscript file
-  manuscript_content <- readLines(here("manuscript/manuscript.qmd"), warn = FALSE)
-  manuscript_text <- paste(manuscript_content, collapse = "\n")
-  
-  # Check that tinytable is used
-  expect_true(grepl("tinytable", manuscript_text, fixed = TRUE),
-              info = "tinytable package not used in manuscript")
-  
-  # Check that theme_latex is used (for LaTeX/PDF output)
-  expect_true(grepl("theme_latex", manuscript_text, fixed = TRUE),
-              info = "tinytable::theme_latex not used in manuscript")
-  
-  # Check for proper tinytable usage pattern
-  expect_true(grepl("tinytable::theme_latex\\(resize_width", manuscript_text),
-              info = "Expected tinytable::theme_latex pattern not found")
-})
-
-test_that("all loaded data files can be read without errors", {
-  # Test that all data files can actually be loaded
-  data_files <- list(
-    corr = "save/corr.RData",
-    PCA = "save/PCA.RData",
-    scree = "save/scree.RData",
-    regTable = "save/regTable.RData",
-    regTable_africa = "save/regTable_africa.RData",
-    regTable_europe = "save/regTable_europe.RData",
-    regTable_fiscal = "save/regTable_fiscal.RData",
-    panel_data_pca = "data/panel_data_pca.rds"
-  )
-  
-  for (file_name in names(data_files)) {
-    file_path <- here(data_files[[file_name]])
-    
-    expect_no_error({
-      if (grepl("\\.rds$", file_path)) {
-        readRDS(file_path)
-      } else {
-        load(file_path)
-      }
-    }, message = paste("Error loading", file_name, "from", data_files[[file_name]]))
-  }
-})
-
-test_that("loaded regression tables are tinytable objects", {
-  # Load each regression table and check it's a tinytable object
-  table_files <- c(
-    "save/regTable.RData",
-    "save/regTable_africa.RData",
-    "save/regTable_europe.RData",
-    "save/regTable_fiscal.RData"
-  )
-  
-  for (table_file in table_files) {
-    # Create a new environment to load into
-    env <- new.env()
-    load(here(table_file), envir = env)
-    
-    # Check that resultsTable was loaded
[…]

### 2025-10-04 tests/testthat/helper-r-chunk.R
Add helper for extracting r chunks

@@ -1,0 +1,44 @@
+  # Helper function to extract R code chunks
+  extract_r_chunks <- function(lines) {
+    chunks <- list()
+    in_chunk <- FALSE
+    chunk_lines <- c()
+    chunk_label <- NULL
+    chunk_num <- 0
+    
+    for (i in seq_along(lines)) {
+      line <- lines[i]
+      
+      # Start of R chunk
+      if (grepl("^```\\{r", line)) {
+        in_chunk <- TRUE
+        chunk_num <- chunk_num + 1
+        
+        # Extract label if present
+        label_match <- regmatches(line, regexec("#\\|\\s*label:\\s*([^\\s}]+)", lines[i + 1]))
+        if (length(label_match[[1]]) > 0) {
+          chunk_label <- label_match[[1]][2]
+        } else {
+          chunk_label <- paste0("chunk_", chunk_num)
+        }
+        
+        chunk_lines <- c()
+        next
+      }
+      
+      # End of chunk
+      if (in_chunk && grepl("^```\\s*$", line)) {
+        in_chunk <- FALSE
+        chunks[[chunk_label]] <- chunk_lines
+        chunk_label <- NULL
+        next
+      }
+      
+      # Inside chunk - collect lines (skip chunk options)
+      if (in_chunk && !grepl("^#\\|", line)) {
+        chunk_lines <- c(chunk_lines, line)
+      }
+    }
+    
+    return(chunks)
+  }

### 2025-10-04 tests/testthat/test-manuscript.R
Add script for testing R chunks in manuscript.qmd by claude

@@ -79,3 +79,74 @@ test_that("manuscript.qmd loads all required data files correctly", {
   }
 })
 
+test_that("R code chunks in manuscript.qmd execute correctly and match snapshots", {
+  # Read manuscript file
+  manuscript_path <- here("manuscript/manuscript.qmd")
+  manuscript_content <- readLines(manuscript_path, warn = FALSE)
+  
+  source(here("tests/testthat/helper-r-chunk.R"))
+  
+  # Extract all R chunks
+  chunks <- extract_r_chunks(manuscript_content)
+  expect_gt(length(chunks), 0, label = "Should find at least one R chunk")
+  
+  # Set working directory to project root for chunk execution
+  old_wd <- getwd()
+  setwd(here())
+  on.exit(setwd(old_wd), add = TRUE)
+  
+  # Create a clean environment for chunk execution
+  chunk_env <- new.env(parent = .GlobalEnv)
+  
+  # Execute each chunk and snapshot outputs
+  for (chunk_name in names(chunks)) {
+    chunk_code <- chunks[[chunk_name]]
+    
+    # Skip empty chunks
+    if (length(chunk_code) == 0 || all(trimws(chunk_code) == "")) {
+      next
+    }
+    
+    # Capture output and any side effects
+    output <- capture.output({
+      result <- tryCatch({
+        eval(parse(text = chunk_code), envir = chunk_env)
+      }, error = function(e) {
+        list(error = e$message)
+      })
+    })
+    
+    # Create snapshot data
+    snapshot_data <- list(
+      chunk_name = chunk_name,
+      code = chunk_code,
+      output = output,
+      result = if (exists("result")) {
+        if (inherits(result, "list") && !is.null(result$error)) {
+          result
+        } else if (is.null(result)) {
+          NULL
+        } else {
+          # Capture structure for non-null results
+          list(
+            class = class(result),
+            summary = if (is.data.frame(result)) {
+              list(
+                nrow = nrow(result),
+                ncol = ncol(result),
+                colnames = colnames(result)
+              )
+            } else {
+              capture.output(str(result, max.level = 2))
+            }
+          )
+        }
+      } else {
+        NULL
+      }
+    )
+    
+    # Snapshot the chunk execution
+    expect_snapshot(snapshot_data, cran = FALSE, variant = chunk_name)
+  }
+})

### 2025-10-04 tests/testthat/test-manuscript.R
Refactor wd and reformat

@@ -3,20 +3,28 @@ library(here)
 i_am("tests/testthat/test-manuscript.R")
 
 test_that("manuscript.qmd file exists and is readable", {
-  expect_true(file.exists(here("manuscript/manuscript.qmd")),
-              info = "manuscript.qmd not found")
-  
+  expect_true(
+    file.exists(here("manuscript/manuscript.qmd")),
+    info = "manuscript.qmd not found"
+  )
+
   # Check that file can be read
-  manuscript_content <- readLines(here("manuscript/manuscript.qmd"), warn = FALSE)
-  expect_gt(length(manuscript_content), 400,
-            label = "manuscript.qmd should have substantial content")
+  manuscript_content <- readLines(
+    here("manuscript/manuscript.qmd"),
+    warn = FALSE
+  )
+  expect_gt(
+    length(manuscript_content),
+    400,
+    label = "manuscript.qmd should have substantial content"
+  )
 })
 
 test_that("all required data dependencies exist", {
   # Check that all required saved data files exist
   required_files <- c(
     "save/corr.RData",
-    "save/PCA.RData", 
+    "save/PCA.RData",
     "save/scree.RData",
     "save/regTable.RData",
     "save/regTable_africa.RData",
@@ -24,42 +32,66 @@ test_that("all required data dependencies exist", {
     "save/regTable_fiscal.RData",
     "data/panel_data_pca.rds"
   )
-  
+
   for (file in required_files) {
-    expect_true(file.exists(here(file)),
-                info = paste(file, "not found - run analysis scripts first"))
+    expect_true(
+      file.exists(here(file)),
+      info = paste(file, "not found - run analysis scripts first")
+    )
   }
 })
 
 test_that("manuscript.qmd contains expected sections", {
   # Read the manuscript file
-  manuscript_content <- readLines(here("manuscript/manuscript.qmd"), warn = FALSE)
+  manuscript_content <- readLines(
+    here("manuscript/manuscript.qmd"),
+    warn = FALSE
+  )
   manuscript_text <- paste(manuscript_content, collapse = "\n")
-  
+
   # Check for key sections
-  expect_true(grepl("# Introduction", manuscript_text, fixed = TRUE),
-              info = "Introduction section not found")
-  expect_true(grepl("# Literature Review", manuscript_text, fixed = FALSE),
-              info = "Literature Review section not found")
-  expect_true(grepl("# Hypotheses", manuscript_text, fixed = TRUE),
-              info = "Hypotheses section not found")
[…]

### 2025-10-04 tests/testthat/test-manuscript.R
bind lines together

@@ -138,11 +138,14 @@ test_that("R code chunks in manuscript.qmd execute correctly and match snapshots
         next
       }
 
+      # Concatenate lines of the chunk into a single string
+      chunk_code_combined <- paste(chunk_code, collapse = "\n")
+
       # Capture output and any side effects
       output <- capture.output({
         result <- tryCatch(
           {
-            eval(parse(text = chunk_code), envir = chunk_env)
+            eval(parse(text = chunk_code_combined), envir = chunk_env)
           },
           error = function(e) {
             list(error = e$message)
@@ -153,7 +156,7 @@ test_that("R code chunks in manuscript.qmd execute correctly and match snapshots
       # Create snapshot data
       snapshot_data <- list(
         chunk_name = chunk_name,
-        code = chunk_code,
+        code = chunk_code_combined,
         output = output,
         result = if (exists("result")) {
           if (inherits(result, "list") && !is.null(result$error)) {

### 2025-10-04 manuscript/manuscript.qmd
manuscript: fix 2nd chunk name

--- 2025-09-24 manuscript/manuscript.qmd
@@
 
-```{r percentage_variance}
+```{r}
+#| label: percentage_variance
 load('save/PCA.RData')

### 2025-10-04 tests/testthat/helper-r-chunk.R
improved snapshot making

@@ -1,44 +1,51 @@
-  # Helper function to extract R code chunks
-  extract_r_chunks <- function(lines) {
-    chunks <- list()
-    in_chunk <- FALSE
-    chunk_lines <- c()
-    chunk_label <- NULL
-    chunk_num <- 0
-    
-    for (i in seq_along(lines)) {
-      line <- lines[i]
-      
-      # Start of R chunk
-      if (grepl("^```\\{r", line)) {
-        in_chunk <- TRUE
-        chunk_num <- chunk_num + 1
-        
-        # Extract label if present
-        label_match <- regmatches(line, regexec("#\\|\\s*label:\\s*([^\\s}]+)", lines[i + 1]))
-        if (length(label_match[[1]]) > 0) {
-          chunk_label <- label_match[[1]][2]
+# Helper function to extract R code chunks
+extract_r_chunks <- function(lines) {
+  chunks <- list()
+  in_chunk <- FALSE
+  chunk_lines <- c()
+  chunk_label <- NULL
+  chunk_num <- 0
+
+  for (i in seq_along(lines)) {
+    line <- lines[i]
+
+    # Start of R chunk
+    if (grepl("^```\\{r", line)) {
+      in_chunk <- TRUE
+      chunk_num <- chunk_num + 1
+
+      # Extract label if present
+      if (i + 1 <= length(lines) && grepl("^#\\|\\s*label:", lines[i + 1])) {
+        label_match <- regmatches(
+          lines[i + 1],
+          regexec("^#\\|\\s*label:\\s*(.+)$", lines[i + 1])
+        )
+        if (length(label_match[[1]]) > 1) {
+          chunk_label <- trimws(label_match[[1]][2]) # Trim whitespace around the label
         } else {
           chunk_label <- paste0("chunk_", chunk_num)
         }
-        
-        chunk_lines <- c()
-        next
+      } else {
+        chunk_label <- paste0("chunk_", chunk_num)
       }
-      
-      # End of chunk
-      if (in_chunk && grepl("^```\\s*$", line)) {
-        in_chunk <- FALSE
-        chunks[[chunk_label]] <- chunk_lines
-        chunk_label <- NULL
-        next
-      }
-      
-      # Inside chunk - collect lines (skip chunk options)
-      if (in_chunk && !grepl("^#\\|", line)) {
-        chunk_lines <- c(chunk_lines, line)
-      }
+
+      chunk_lines <- c()
[…]

### 2025-10-04 tests/testthat/test-manuscript.R
Use LaTeX output for tables

@@ -138,6 +138,9 @@ test_that("R code chunks in manuscript.qmd execute correctly and match snapshots
         next
       }
 
+      # Check if this is a table chunk (starts with tbl-)
+      is_table_chunk <- grepl("^tbl-", chunk_name)
+
       # Concatenate lines of the chunk into a single string
       chunk_code_combined <- paste(chunk_code, collapse = "\n")
 
@@ -153,35 +156,60 @@ test_that("R code chunks in manuscript.qmd execute correctly and match snapshots
         )
       })
 
-      # Create snapshot data
-      snapshot_data <- list(
-        chunk_name = chunk_name,
-        code = chunk_code_combined,
-        output = output,
-        result = if (exists("result")) {
+      # Create snapshot data based on chunk type
+      if (is_table_chunk) {
+        # For table chunks, capture the LaTeX output
+        latex_output <- if (exists("result") && !is.null(result)) {
           if (inherits(result, "list") && !is.null(result$error)) {
             result
-          } else if (is.null(result)) {
-            NULL
+          } else if (inherits(result, "tinytable")) {
+            # Capture LaTeX representation
+            capture.output(print(result, output = "latex"))
           } else {
-            # Capture structure for non-null results
-            list(
-              class = class(result),
-              summary = if (is.data.frame(result)) {
-                list(
-                  nrow = nrow(result),
-                  ncol = ncol(result),
-                  colnames = colnames(result)
-                )
-              } else {
-                capture.output(str(result, max.level = 2))
-              }
-            )
+            # Try to convert to character/print for other table types
+            capture.output(print(result))
           }
         } else {
           NULL
         }
-      )
+        
+        snapshot_data <- list(
+          chunk_name = chunk_name,
+          code = chunk_code_combined,
+          latex_output = latex_output
+        )
+      } else {
+        # For non-table chunks, use the summary approach
+        snapshot_data <- list(
+          chunk_name = chunk_name,
+          code = chunk_code_combined,
+          output = output,
+          result = if (exists("result")) {
+            if (inherits(result, "list") && !is.null(result$error)) {
+              result
+            } else if (is.null(result)) {
+              NULL
+            } else {
+              # Capture structure for non-null results
+              list(
+                class = class(result),
+                summary = if (is.data.frame(result)) {
+                  list(
+                    nrow = nrow(result),
+                    ncol = ncol(result),
+                    colnames = colnames(result)
+                  )
+                } else {
[…]

### 2025-10-04 tests/testthat/test-manuscript.R
Prevent writing to PDF

@@ -129,6 +129,9 @@ test_that("R code chunks in manuscript.qmd execute correctly and match snapshots
     # Create a clean environment for chunk execution
     chunk_env <- new.env(parent = .GlobalEnv)
 
+    # Suppress Rplots.pdf creation
+    pdf(NULL)
+
     # Execute each chunk and snapshot outputs
     for (chunk_name in names(chunks)) {
       chunk_code <- chunks[[chunk_name]]
@@ -172,7 +175,7 @@ test_that("R code chunks in manuscript.qmd execute correctly and match snapshots
         } else {
           NULL
         }
-        
+
         snapshot_data <- list(
           chunk_name = chunk_name,
           code = chunk_code_combined,
@@ -209,9 +212,12 @@ test_that("R code chunks in manuscript.qmd execute correctly and match snapshots
           }
         )
       }
-      
+
       # Snapshot the chunk execution
       expect_snapshot(snapshot_data, cran = FALSE, variant = chunk_name)
     }
+
+    # Close the PDF device
+    dev.off()
   })
 })

### 2025-10-04 tests/testthat/test-manuscript.R
escape -0.00 for ARM MacOS

@@ -176,6 +176,15 @@ test_that("R code chunks in manuscript.qmd execute correctly and match snapshots
           NULL
         }
 
+        # Avoid floating point issues in ARM MacOS
+        if (!is.null(latex_output)) {
+          latex_output <- gsub(
+            "\\\\num\\{-0\\.00\\}",
+            "\\\\num{0.00}",
+            latex_output
+          )
+        }
+        
         snapshot_data <- list(
           chunk_name = chunk_name,
           code = chunk_code_combined,

### 2025-10-04 manuscript/manuscript.qmd
Add tests for manuscript (#88)
* Initial plan

* Add comprehensive test suite for manuscript.qmd

Co-authored-by: kv9898 <105025148+kv9898@users.noreply.github.com>

* Update README to document new manuscript test

Co-authored-by: kv9898 <105025148+kv9898@users.noreply.github.com>

* change expected length to 400

* remove non-dependencies

* Fix that part that checks loading

* Remove final parts

* Add helper for extracting r chunks

* Add script for testing R chunks in manuscript.qmd by claude

* Refactor wd and reformat

* bind lines together

* manuscript: fix 2nd chunk name

* improved s

@@
     Mean + SD + Min + (Q1 = P25) + Median + (Q3 = P75) + Max + N,
+  output = "latex",
   data = dta

### 2025-10-04 tests/testthat/helper-r-chunk.R

@@ -1,0 +1,51 @@
+# Helper function to extract R code chunks
+extract_r_chunks <- function(lines) {
+  chunks <- list()
+  in_chunk <- FALSE
+  chunk_lines <- c()
+  chunk_label <- NULL
+  chunk_num <- 0
+
+  for (i in seq_along(lines)) {
+    line <- lines[i]
+
+    # Start of R chunk
+    if (grepl("^```\\{r", line)) {
+      in_chunk <- TRUE
+      chunk_num <- chunk_num + 1
+
+      # Extract label if present
+      if (i + 1 <= length(lines) && grepl("^#\\|\\s*label:", lines[i + 1])) {
+        label_match <- regmatches(
+          lines[i + 1],
+          regexec("^#\\|\\s*label:\\s*(.+)$", lines[i + 1])
+        )
+        if (length(label_match[[1]]) > 1) {
+          chunk_label <- trimws(label_match[[1]][2]) # Trim whitespace around the label
+        } else {
+          chunk_label <- paste0("chunk_", chunk_num)
+        }
+      } else {
+        chunk_label <- paste0("chunk_", chunk_num)
+      }
+
+      chunk_lines <- c()
+      next
+    }
+
+    # End of chunk
+    if (in_chunk && grepl("^```\\s*$", line)) {
+      in_chunk <- FALSE
+      chunks[[chunk_label]] <- chunk_lines
+      chunk_label <- NULL
+      next
+    }
+
+    # Inside chunk - collect lines (skip chunk options)
+    if (in_chunk && !grepl("^#\\|", line)) {
+      chunk_lines <- c(chunk_lines, line)
+    }
+  }
+
+  return(chunks)
+}

### 2025-10-04 tests/testthat/test-manuscript.R

@@ -1,0 +1,232 @@
+library(testthat)
+library(here)
+i_am("tests/testthat/test-manuscript.R")
+
+test_that("manuscript.qmd file exists and is readable", {
+  expect_true(
+    file.exists(here("manuscript/manuscript.qmd")),
+    info = "manuscript.qmd not found"
+  )
+
+  # Check that file can be read
+  manuscript_content <- readLines(
+    here("manuscript/manuscript.qmd"),
+    warn = FALSE
+  )
+  expect_gt(
+    length(manuscript_content),
+    400,
+    label = "manuscript.qmd should have substantial content"
+  )
+})
+
+test_that("all required data dependencies exist", {
+  # Check that all required saved data files exist
+  required_files <- c(
+    "save/corr.RData",
+    "save/PCA.RData",
+    "save/scree.RData",
+    "save/regTable.RData",
+    "save/regTable_africa.RData",
+    "save/regTable_europe.RData",
+    "save/regTable_fiscal.RData",
+    "data/panel_data_pca.rds"
+  )
+
+  for (file in required_files) {
+    expect_true(
+      file.exists(here(file)),
+      info = paste(file, "not found - run analysis scripts first")
+    )
+  }
+})
+
+test_that("manuscript.qmd contains expected sections", {
+  # Read the manuscript file
+  manuscript_content <- readLines(
+    here("manuscript/manuscript.qmd"),
+    warn = FALSE
+  )
+  manuscript_text <- paste(manuscript_content, collapse = "\n")
+
+  # Check for key sections
+  expect_true(
+    grepl("# Introduction", manuscript_text, fixed = TRUE),
+    info = "Introduction section not found"
+  )
+  expect_true(
+    grepl("# Literature Review", manuscript_text, fixed = FALSE),
+    info = "Literature Review section not found"
+  )
+  expect_true(
+    grepl("# Hypotheses", manuscript_text, fixed = TRUE),
+    info = "Hypotheses section not found"
+  )
+  expect_true(
+    grepl("# Data", manuscript_text, fixed = TRUE),
+    info = "Data section not found"
+  )
+  expect_true(
+    grepl("# Model Specification", manuscript_text, fixed = FALSE),
+    info = "Model Specification section not found"
+  )
+  expect_true(
+    grepl("# Results", manuscript_text, fixed = TRUE),
+    info = "Results section not found"
+  )
+  expect_true(
[…]

### 2025-10-04 manuscript/manuscript.qmd
Update packages and disable DESCRIPTION (#84)
* Update packages

* Update tinytable outputs

* Remove DESCRIPTION

* Downgrade Matrix (later version not usable)

* Add preamble for new tinytable

* update scree plot

* README: use test-manuscript as file example

* test-manuscript: use local edition 3

* Update snapshots

@@
           \usepackage{amsthm, thm-restate}
+          \usepackage{}
           \newtheorem{hyp}{Hypothesis}
@@
           \AtAppendix{\counterwithin{table}{section}}
+      - file: extra/preamble.tex
 abstract: "This paper investigates whether the United States or Europe holds greater influence over the International Monetary Fund (IMF)—a longstanding yet underexamined question in the study of global economic governance.

### 2025-10-04 tests/testthat/test-manuscript.R

@@ -1,6 +1,7 @@
 library(testthat)
 library(here)
 i_am("tests/testthat/test-manuscript.R")
+testthat::local_edition(3)
 
 test_that("manuscript.qmd file exists and is readable", {
   expect_true(
